---
title: "International Student Presence and Dutch Graduate Labor Market Outcomes"
subtitle: "A Shift-Share Instrumental Variables Approach — Analysis"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages}
pacman::p_load(`pacman`)
p_load(tidyverse); p_load(fixest); p_load(kableExtra)
```

```{r load-data}
# -- Load the estimation dataset saved by ThesisManualH4_prep.Rmd -----------
est_data <- read.csv('./est_data.csv', stringsAsFactors = FALSE)
cat(sprintf('Loaded %d observations\n', nrow(est_data)))
```

```{r regressions, results='asis'}
# -- Main regression results ---------------------------------------------------
# Three panels: (A) Full sample, (B) HBO only, (C) WO only.
# Within each panel: OLS and IV (2SLS) for Full, STEM, and Non-STEM.
# All models use aggregate share_int_agg ~ Z_visa_field_agg.
# Controls: age, age², education_years, hours_worked, hours², partner, geslacht.
# Fixed effect: field. SEs clustered at field level.

# --- Panel A: All onderwijssoort (pooled) ------------------------------------

a_ols_full  <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = est_data, cluster = ~ field)
a_iv_full   <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                       share_int_agg ~ Z_visa_field_agg,
                      data = est_data, cluster = ~ field)
a_ols_stem  <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = est_data, cluster = ~ field,
                      subset = ~ stem)
a_iv_stem   <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                       share_int_agg ~ Z_visa_field_agg,
                      data = est_data, cluster = ~ field,
                      subset = ~ stem)
a_ols_nstem <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = est_data, cluster = ~ field,
                      subset = ~ !stem)
a_iv_nstem  <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                       share_int_agg ~ Z_visa_field_agg,
                      data = est_data, cluster = ~ field,
                      subset = ~ !stem)

# First-stage F-stats (Panel A)
fs_a_full <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                    data = est_data, cluster = ~ field)
fs_a_stem <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                    data = est_data, cluster = ~ field, subset = ~ stem)
fs_a_nstem <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                     data = est_data, cluster = ~ field, subset = ~ !stem)

f_a_full  <- round(wald(fs_a_full,  'Z_visa_field_agg', print = FALSE)$stat, 1)
f_a_stem  <- round(wald(fs_a_stem,  'Z_visa_field_agg', print = FALSE)$stat, 1)
f_a_nstem <- round(wald(fs_a_nstem, 'Z_visa_field_agg', print = FALSE)$stat, 1)

cat('\n\\textbf{Panel A: All Education Levels}\

')

etable(a_ols_full, a_iv_full, a_ols_stem, a_iv_stem, a_ols_nstem, a_iv_nstem,
       se.below = TRUE,
       dict = c(share_int_agg = 'Intl. Share'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('Full', '', 'STEM', '', 'Non-STEM', ''),
       extralines = list('F-stat (1st stage)' =
                          c(f_a_full, f_a_full, f_a_stem, f_a_stem, f_a_nstem, f_a_nstem)))

# --- Panel B: HBO only -------------------------------------------------------

hbo <- est_data %>% filter(onderwijssoort == 'Hoger beroepsonderwijs')

b_ols_full  <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = hbo, cluster = ~ field)
b_iv_full   <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                       share_int_agg ~ Z_visa_field_agg,
                      data = hbo, cluster = ~ field)
b_ols_stem  <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = hbo, cluster = ~ field, subset = ~ stem)
b_iv_stem   <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                       share_int_agg ~ Z_visa_field_agg,
                      data = hbo, cluster = ~ field, subset = ~ stem)
b_ols_nstem <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = hbo, cluster = ~ field, subset = ~ !stem)
b_iv_nstem  <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                       share_int_agg ~ Z_visa_field_agg,
                      data = hbo, cluster = ~ field, subset = ~ !stem)

fs_b_full <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                    data = hbo, cluster = ~ field)
fs_b_stem <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                    data = hbo, cluster = ~ field, subset = ~ stem)
fs_b_nstem <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                     data = hbo, cluster = ~ field, subset = ~ !stem)

f_b_full  <- round(wald(fs_b_full,  'Z_visa_field_agg', print = FALSE)$stat, 1)
f_b_stem  <- round(wald(fs_b_stem,  'Z_visa_field_agg', print = FALSE)$stat, 1)
f_b_nstem <- round(wald(fs_b_nstem, 'Z_visa_field_agg', print = FALSE)$stat, 1)

cat('\n\\textbf{Panel B: HBO (Higher Professional Education)}\

')

etable(b_ols_full, b_iv_full, b_ols_stem, b_iv_stem, b_ols_nstem, b_iv_nstem,
       se.below = TRUE,
       dict = c(share_int_agg = 'Intl. Share'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('Full', '', 'STEM', '', 'Non-STEM', ''),
       extralines = list('F-stat (1st stage)' =
                          c(f_b_full, f_b_full, f_b_stem, f_b_stem, f_b_nstem, f_b_nstem)))

# --- Panel C: WO only --------------------------------------------------------

wo <- est_data %>% filter(onderwijssoort == 'Wetenschappelijk onderwijs')

c_ols_full  <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = wo, cluster = ~ field)
c_iv_full   <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                       share_int_agg ~ Z_visa_field_agg,
                      data = wo, cluster = ~ field)
c_ols_stem  <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = wo, cluster = ~ field, subset = ~ stem)
c_iv_stem   <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                       share_int_agg ~ Z_visa_field_agg,
                      data = wo, cluster = ~ field, subset = ~ stem)
c_ols_nstem <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = wo, cluster = ~ field, subset = ~ !stem)
c_iv_nstem  <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                       share_int_agg ~ Z_visa_field_agg,
                      data = wo, cluster = ~ field, subset = ~ !stem)

fs_c_full <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                    data = wo, cluster = ~ field)
fs_c_stem <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                    data = wo, cluster = ~ field, subset = ~ stem)
fs_c_nstem <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                     data = wo, cluster = ~ field, subset = ~ !stem)

f_c_full  <- round(wald(fs_c_full,  'Z_visa_field_agg', print = FALSE)$stat, 1)
f_c_stem  <- round(wald(fs_c_stem,  'Z_visa_field_agg', print = FALSE)$stat, 1)
f_c_nstem <- round(wald(fs_c_nstem, 'Z_visa_field_agg', print = FALSE)$stat, 1)

cat('\n\\textbf{Panel C: WO (Research University)}\

')

etable(c_ols_full, c_iv_full, c_ols_stem, c_iv_stem, c_ols_nstem, c_iv_nstem,
       se.below = TRUE,
       dict = c(share_int_agg = 'Intl. Share'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('Full', '', 'STEM', '', 'Non-STEM', ''),
       extralines = list('F-stat (1st stage)' =
                          c(f_c_full, f_c_full, f_c_stem, f_c_stem, f_c_nstem, f_c_nstem)))

# --- Panel D: Full sample with education FE --------------------------------

d_ols_full  <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field + onderwijssoort,
                      data = est_data, cluster = ~ field)
d_iv_full   <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field + onderwijssoort |
                       share_int_agg ~ Z_visa_field_agg,
                      data = est_data, cluster = ~ field)
d_ols_stem  <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field + onderwijssoort,
                      data = est_data, cluster = ~ field, subset = ~ stem)
d_iv_stem   <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field + onderwijssoort |
                       share_int_agg ~ Z_visa_field_agg,
                      data = est_data, cluster = ~ field, subset = ~ stem)
d_ols_nstem <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field + onderwijssoort,
                      data = est_data, cluster = ~ field, subset = ~ !stem)
d_iv_nstem  <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field + onderwijssoort |
                       share_int_agg ~ Z_visa_field_agg,
                      data = est_data, cluster = ~ field, subset = ~ !stem)

fs_d_full <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field + onderwijssoort,
                    data = est_data, cluster = ~ field)
fs_d_stem <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field + onderwijssoort,
                     data = est_data, cluster = ~ field, subset = ~ stem)
fs_d_nstem <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field + onderwijssoort,
                     data = est_data, cluster = ~ field, subset = ~ !stem)

f_d_full  <- round(wald(fs_d_full,  'Z_visa_field_agg', print = FALSE)$stat, 1)
f_d_stem  <- round(wald(fs_d_stem,  'Z_visa_field_agg', print = FALSE)$stat, 1)
f_d_nstem <- round(wald(fs_d_nstem, 'Z_visa_field_agg', print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel D: Full Sample with Education FE}\\\n\n')

etable(d_ols_full, d_iv_full, d_ols_stem, d_iv_stem, d_ols_nstem, d_iv_nstem,
       se.below = TRUE,
       dict = c(share_int_agg = 'Intl. Share'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       extralines = list('F-stat (1st stage)' =
                          c(f_d_full, f_d_full, f_d_stem, f_d_stem, f_d_nstem, f_d_nstem)))

# --- Panel F: Gender Effects -------------------------------------------------

female <- est_data %>% filter(geslacht == 'female')
male   <- est_data %>% filter(geslacht == 'male')

# OLS with gender interaction
f_ols_interact <- feols(log_income ~ share_int_agg * geslacht + age + age_sq + education_years + hours_worked + hours_sq + partner | field,
                        data = est_data, cluster = ~ field)

# IV split by gender
f_iv_female <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner | field |
                      share_int_agg ~ Z_visa_field_agg,
                     data = female, cluster = ~ field)
f_iv_male <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner | field |
                    share_int_agg ~ Z_visa_field_agg,
                   data = male, cluster = ~ field)

# First-stage F-stats (split by gender)
fs_female <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner | field,
                   data = female, cluster = ~ field)
fs_male <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner | field,
                 data = male, cluster = ~ field)

f_female <- round(wald(fs_female, 'Z_visa_field_agg', print = FALSE)$stat, 1)
f_male   <- round(wald(fs_male,   'Z_visa_field_agg', print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel F: Gender Effects}\\
\\textit{OLS with interaction; IV estimated separately by gender.}\\
\\textit{Gender subsample F-stats: Females = ', f_female, ', Males = ', f_male, '.}\\

')

etable(f_ols_interact, f_iv_female, f_iv_male,
       se.below = TRUE,
       dict = c(share_int_agg = 'Intl. Share',
                'share_int_agg:geslachtmale' = 'Intl. Share × Male'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('OLS (interact)', 'IV: Female', 'IV: Male'),
       extralines = list('F-stat (1st stage)' = c('', f_female, f_male)))

# --- Panel G: Migration Background Effects -----------------------------------

dutch_sample <- est_data %>% filter(dutch == 'dutch')
non_dutch    <- est_data %>% filter(dutch == 'non_dutch')

# OLS with migration interaction
g_ols_interact <- feols(log_income ~ share_int_agg * dutch + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                        data = est_data, cluster = ~ field)

# IV split by migration background
g_iv_dutch <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                     share_int_agg ~ Z_visa_field_agg,
                    data = dutch_sample, cluster = ~ field)
g_iv_non_dutch <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                         share_int_agg ~ Z_visa_field_agg,
                        data = non_dutch, cluster = ~ field)

# First-stage F-stats
fs_dutch <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                  data = dutch_sample, cluster = ~ field)
fs_non_dutch <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = non_dutch, cluster = ~ field)

f_dutch <- round(wald(fs_dutch, 'Z_visa_field_agg', print = FALSE)$stat, 1)
f_non_dutch <- round(wald(fs_non_dutch, 'Z_visa_field_agg', print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel G: Migration Background Effects}\\
\\textit{OLS with interaction; IV estimated separately by background.}\\
\\textit{F-stats: Dutch = ', f_dutch, ', Non-Dutch = ', f_non_dutch, '.}\\

')

etable(g_ols_interact, g_iv_dutch, g_iv_non_dutch,
       se.below = TRUE,
       dict = c(share_int_agg = 'Intl. Share',
                'share_int_agg:dutchnon_dutch' = 'Intl. Share × Non-Dutch'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('OLS (interact)', 'IV: Dutch', 'IV: Non-Dutch'),
       extralines = list('F-stat (1st stage)' = c('', f_dutch, f_non_dutch)))

# --- Panel H: Age Group Effects (20-34 vs 35+) --------------------------------
# Younger workers (20-34) face more immediate competition from new graduates,
# potentially depressing wages (negative IV coefficient), while older workers
# (35+) may benefit from complementarities (positive coefficient).
# Split-sample IV because single instrument cannot identify interaction terms.

# Split into age groups
young <- est_data %>% filter(age >= 20 & age <= 34)
older <- est_data %>% filter(age >= 35)

# OLS with age-group interaction
h_ols_interact <- feols(
  log_income ~ share_int_agg * age_young + age + age_sq + education_years +
    hours_worked + hours_sq + partner + geslacht | field,
  data = est_data %>% mutate(age_young = age >= 20 & age <= 34),
  cluster = ~ field
)

# IV split by age group
h_iv_young <- feols(
  log_income ~ age + age_sq + education_years + hours_worked + hours_sq +
    partner + geslacht | field | share_int_agg ~ Z_visa_field_agg,
  data = young, cluster = ~ field
)
h_iv_older <- feols(
  log_income ~ age + age_sq + education_years + hours_worked + hours_sq +
    partner + geslacht | field | share_int_agg ~ Z_visa_field_agg,
  data = older, cluster = ~ field
)

# First-stage F-stats (split by age group)
fs_young <- feols(
  share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years +
    hours_worked + hours_sq + partner + geslacht | field,
  data = young, cluster = ~ field
)
fs_older <- feols(
  share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years +
    hours_worked + hours_sq + partner + geslacht | field,
  data = older, cluster = ~ field
)

f_young <- round(wald(fs_young, 'Z_visa_field_agg', print = FALSE)$stat, 1)
f_older <- round(wald(fs_older, 'Z_visa_field_agg', print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel H: Age Group Effects (20-34 vs 35+)}\\
\\textit{OLS with interaction; IV estimated separately by age group.}\\
\\textit{Age subsample F-stats: 20-34 = ', f_young, ', 35+ = ', f_older, '.}\\
\n')

etable(h_ols_interact, h_iv_young, h_iv_older,
       se.below = TRUE,
       dict = c(share_int_agg = 'Intl. Share',
                'share_int_agg:age_youngTRUE' = 'Intl. Share × Age 20-34'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('OLS (interact)', 'IV: Age 20-34', 'IV: Age 35+'),
       extralines = list('F-stat (1st stage)' = c('', f_young, f_older)))
```

```{r regressions-lag}
# -- Panel E: Lagged IV specifications --------------------------------------
# OLS + IV pairs for contemporaneous and lagged instruments.
# 'Long-term' = all three lags jointly instrumenting share_int_agg.
# Full sample only (not split by STEM).

# Build lagged instruments
lags <- est_data %>%
  group_by(field, survey_year) %>%
  summarise(Z_visa = first(Z_visa_field_agg), .groups = 'drop')

est_data_lag <- est_data
for (L in 1:3) {
  lag_df <- lags
  lag_df$survey_year <- lag_df$survey_year + L
  names(lag_df)[3] <- paste0('Z_visa_lag', L)
  est_data_lag <- est_data_lag %>% left_join(lag_df, by = c('field', 'survey_year'))
}

# OLS + IV pairs for each specification
ols_contemp <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                      data = est_data_lag, cluster = ~ field)
iv_contemp  <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                       share_int_agg ~ Z_visa_field_agg,
                      data = est_data_lag, cluster = ~ field)

ols_lag1 <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                   data = est_data_lag, cluster = ~ field)
iv_lag1  <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                    share_int_agg ~ Z_visa_lag1,
                   data = est_data_lag, cluster = ~ field)

ols_lag2 <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                   data = est_data_lag, cluster = ~ field)
iv_lag2  <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                    share_int_agg ~ Z_visa_lag2,
                   data = est_data_lag, cluster = ~ field)

ols_lag3 <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                   data = est_data_lag, cluster = ~ field)
iv_lag3  <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                    share_int_agg ~ Z_visa_lag3,
                   data = est_data_lag, cluster = ~ field)

ols_all <- feols(log_income ~ share_int_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                  data = est_data_lag, cluster = ~ field)
iv_all  <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field |
                   share_int_agg ~ Z_visa_lag1 + Z_visa_lag2 + Z_visa_lag3,
                  data = est_data_lag, cluster = ~ field)

# First-stage F-stats
fs_contemp <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                     data = est_data_lag, cluster = ~ field)
fs_lag1 <- feols(share_int_agg ~ Z_visa_lag1 + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                  data = est_data_lag, cluster = ~ field)
fs_lag2 <- feols(share_int_agg ~ Z_visa_lag2 + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                  data = est_data_lag, cluster = ~ field)
fs_lag3 <- feols(share_int_agg ~ Z_visa_lag3 + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                  data = est_data_lag, cluster = ~ field)
fs_all <- feols(share_int_agg ~ Z_visa_lag1 + Z_visa_lag2 + Z_visa_lag3 + age + age_sq + education_years + hours_worked + hours_sq + partner + geslacht | field,
                 data = est_data_lag, cluster = ~ field)

f_contemp <- round(wald(fs_contemp, 'Z_visa_field_agg', print = FALSE)$stat, 1)
f_lag1 <- round(wald(fs_lag1, 'Z_visa_lag1', print = FALSE)$stat, 1)
f_lag2 <- round(wald(fs_lag2, 'Z_visa_lag2', print = FALSE)$stat, 1)
f_lag3 <- round(wald(fs_lag3, 'Z_visa_lag3', print = FALSE)$stat, 1)
f_all <- round(wald(fs_all, c('Z_visa_lag1', 'Z_visa_lag2', 'Z_visa_lag3'), print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel E: Lagged IV Specifications}\\n\\textit{Each pair: OLS | IV. Long-term = lags 1-3 jointly.}\\n
')

etable(ols_contemp, iv_contemp, ols_lag1, iv_lag1, ols_lag2, iv_lag2, ols_lag3, iv_lag3, ols_all, iv_all,
       se.below = TRUE,
       dict = c(share_int_agg = 'Intl. Share'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('Contemp', '', 'Lag 1', '', 'Lag 2', '', 'Lag 3', '', 'Long-term', ''),
       subtitles = c('OLS', 'IV', 'OLS', 'IV', 'OLS', 'IV', 'OLS', 'IV', 'OLS', 'IV'),
       extralines = list('F-stat (1st stage)' =
                          c('', f_contemp, '', f_lag1, '', f_lag2, '', f_lag3, '', f_all)))
```

```{r regressions-z, results='asis'}
# -- Standardized regressions ---------------------------------------------------
# Same 4 panels as above but with z-score standardized continuous variables
# so coefficients represent the effect of a one-SD change.

est_data_z <- est_data %>%
  mutate(across(c(share_int_agg, age, education_years, hours_worked,
                   log_income, Z_visa_field_agg),
                ~ ( . - mean(., na.rm = TRUE)) / sd(., na.rm = TRUE),
                .names = "{.col}_z")) %>%
  mutate(age_sq_z = age_z^2,
         hours_sq_z = hours_worked_z^2)

cat('\n\n\\textbf{Standardized Regressions}\\n\n')
cat('All continuous variables normalized to z-scores (mean 0, SD 1).\n\n')

# --- Panel A: All onderwijssoort (pooled) ------------------------------------

az_ols_full  <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field,
                      data = est_data_z, cluster = ~ field)
az_iv_full   <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = est_data_z, cluster = ~ field)
az_ols_stem  <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field,
                      data = est_data_z, cluster = ~ field, subset = ~ stem)
az_iv_stem   <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = est_data_z, cluster = ~ field, subset = ~ stem)
az_ols_nstem <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field,
                      data = est_data_z, cluster = ~ field, subset = ~ !stem)
az_iv_nstem  <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = est_data_z, cluster = ~ field, subset = ~ !stem)

fs_az_full <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field,
                    data = est_data_z, cluster = ~ field)
fs_az_stem <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field,
                    data = est_data_z, cluster = ~ field, subset = ~ stem)
fs_az_nstem <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                       education_years_z + hours_worked_z + hours_sq_z +
                       partner + geslacht | field,
                     data = est_data_z, cluster = ~ field, subset = ~ !stem)

f_az_full  <- round(wald(fs_az_full,  'Z_visa_field_agg_z', print = FALSE)$stat, 1)
f_az_stem  <- round(wald(fs_az_stem,  'Z_visa_field_agg_z', print = FALSE)$stat, 1)
f_az_nstem <- round(wald(fs_az_nstem, 'Z_visa_field_agg_z', print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel A: All Education Levels (Standardized)}\\n\n')

etable(az_ols_full, az_iv_full, az_ols_stem, az_iv_stem, az_ols_nstem, az_iv_nstem,
       se.below = TRUE,
       dict = c(share_int_agg_z = 'Intl. Share (z)', Z_visa_field_agg_z = 'Visa Instr. (z)'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('Full', '', 'STEM', '', 'Non-STEM', ''),
       extralines = list('F-stat (1st stage)' =
                          c(f_az_full, f_az_full, f_az_stem, f_az_stem,
                            f_az_nstem, f_az_nstem)))

# --- Panel B: HBO only -------------------------------------------------------

hbo_z <- est_data_z %>% filter(onderwijssoort == 'Hoger beroepsonderwijs')

bz_ols_full  <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field,
                      data = hbo_z, cluster = ~ field)
bz_iv_full   <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = hbo_z, cluster = ~ field)
bz_ols_stem  <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field,
                      data = hbo_z, cluster = ~ field, subset = ~ stem)
bz_iv_stem   <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = hbo_z, cluster = ~ field, subset = ~ stem)
bz_ols_nstem <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field,
                      data = hbo_z, cluster = ~ field, subset = ~ !stem)
bz_iv_nstem  <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = hbo_z, cluster = ~ field, subset = ~ !stem)

fs_bz_full <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field,
                    data = hbo_z, cluster = ~ field)
fs_bz_stem <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field,
                    data = hbo_z, cluster = ~ field, subset = ~ stem)
fs_bz_nstem <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                       education_years_z + hours_worked_z + hours_sq_z +
                       partner + geslacht | field,
                     data = hbo_z, cluster = ~ field, subset = ~ !stem)

f_bz_full  <- round(wald(fs_bz_full,  'Z_visa_field_agg_z', print = FALSE)$stat, 1)
f_bz_stem  <- round(wald(fs_bz_stem,  'Z_visa_field_agg_z', print = FALSE)$stat, 1)
f_bz_nstem <- round(wald(fs_bz_nstem, 'Z_visa_field_agg_z', print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel B: HBO (Standardized)}\\n\n')

etable(bz_ols_full, bz_iv_full, bz_ols_stem, bz_iv_stem, bz_ols_nstem, bz_iv_nstem,
       se.below = TRUE,
       dict = c(share_int_agg_z = 'Intl. Share (z)', Z_visa_field_agg_z = 'Visa Instr. (z)'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('Full', '', 'STEM', '', 'Non-STEM', ''),
       extralines = list('F-stat (1st stage)' =
                          c(f_bz_full, f_bz_full, f_bz_stem, f_bz_stem,
                            f_bz_nstem, f_bz_nstem)))

# --- Panel C: WO only --------------------------------------------------------

wo_z <- est_data_z %>% filter(onderwijssoort == 'Wetenschappelijk onderwijs')

cz_ols_full  <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field,
                      data = wo_z, cluster = ~ field)
cz_iv_full   <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = wo_z, cluster = ~ field)
cz_ols_stem  <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field,
                      data = wo_z, cluster = ~ field, subset = ~ stem)
cz_iv_stem   <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = wo_z, cluster = ~ field, subset = ~ stem)
cz_ols_nstem <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field,
                      data = wo_z, cluster = ~ field, subset = ~ !stem)
cz_iv_nstem  <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = wo_z, cluster = ~ field, subset = ~ !stem)

fs_cz_full <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field,
                    data = wo_z, cluster = ~ field)
fs_cz_stem <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field,
                    data = wo_z, cluster = ~ field, subset = ~ stem)
fs_cz_nstem <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                       education_years_z + hours_worked_z + hours_sq_z +
                       partner + geslacht | field,
                     data = wo_z, cluster = ~ field, subset = ~ !stem)

f_cz_full  <- round(wald(fs_cz_full,  'Z_visa_field_agg_z', print = FALSE)$stat, 1)
f_cz_stem  <- round(wald(fs_cz_stem,  'Z_visa_field_agg_z', print = FALSE)$stat, 1)
f_cz_nstem <- round(wald(fs_cz_nstem, 'Z_visa_field_agg_z', print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel C: WO (Standardized)}\\n\n')

etable(cz_ols_full, cz_iv_full, cz_ols_stem, cz_iv_stem, cz_ols_nstem, cz_iv_nstem,
       se.below = TRUE,
       dict = c(share_int_agg_z = 'Intl. Share (z)', Z_visa_field_agg_z = 'Visa Instr. (z)'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('Full', '', 'STEM', '', 'Non-STEM', ''),
       extralines = list('F-stat (1st stage)' =
                          c(f_cz_full, f_cz_full, f_cz_stem, f_cz_stem,
                            f_cz_nstem, f_cz_nstem)))

# --- Panel D-z: Full sample with education FE (standardized) -----------------

dz_ols_full  <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field + onderwijssoort,
                      data = est_data_z, cluster = ~ field)
dz_iv_full   <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field + onderwijssoort |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = est_data_z, cluster = ~ field)
dz_ols_stem  <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field + onderwijssoort,
                      data = est_data_z, cluster = ~ field, subset = ~ stem)
dz_iv_stem   <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field + onderwijssoort |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = est_data_z, cluster = ~ field, subset = ~ stem)
dz_ols_nstem <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                        education_years_z + hours_worked_z + hours_sq_z +
                        partner + geslacht | field + onderwijssoort,
                      data = est_data_z, cluster = ~ field, subset = ~ !stem)
dz_iv_nstem  <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                        hours_worked_z + hours_sq_z + partner + geslacht | field + onderwijssoort |
                        share_int_agg_z ~ Z_visa_field_agg_z,
                      data = est_data_z, cluster = ~ field, subset = ~ !stem)

fs_dz_full <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field + onderwijssoort,
                    data = est_data_z, cluster = ~ field)
fs_dz_stem <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field + onderwijssoort,
                    data = est_data_z, cluster = ~ field, subset = ~ stem)
fs_dz_nstem <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z +
                       education_years_z + hours_worked_z + hours_sq_z +
                       partner + geslacht | field + onderwijssoort,
                     data = est_data_z, cluster = ~ field, subset = ~ !stem)

f_dz_full  <- round(wald(fs_dz_full,  'Z_visa_field_agg_z', print = FALSE)$stat, 1)
f_dz_stem  <- round(wald(fs_dz_stem,  'Z_visa_field_agg_z', print = FALSE)$stat, 1)
f_dz_nstem <- round(wald(fs_dz_nstem, 'Z_visa_field_agg_z', print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel D: Full Sample with Education FE (Standardized)}\\\n\n')

etable(dz_ols_full, dz_iv_full, dz_ols_stem, dz_iv_stem, dz_ols_nstem, dz_iv_nstem,
       se.below = TRUE,
       dict = c(share_int_agg_z = 'Intl. Share (z)', Z_visa_field_agg_z = 'Visa Instr. (z)'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('Full', '', 'STEM', '', 'Non-STEM', ''),
       extralines = list('F-stat (1st stage)' =
                          c(f_dz_full, f_dz_full, f_dz_stem, f_dz_stem,
                            f_dz_nstem, f_dz_nstem)))

# --- Panel H-z: Age Group Effects (20-34 vs 35+) — Standardized -------------

# Split into age groups (using already-standardized data)
young_z <- est_data_z %>% filter(age >= 20 & age <= 34)
older_z <- est_data_z %>% filter(age >= 35)

# OLS with age-group interaction
hz_ols_interact <- feols(
  log_income_z ~ share_int_agg_z * age_young + age_z + age_sq_z +
    education_years_z + hours_worked_z + hours_sq_z + partner + geslacht | field,
  data = est_data_z %>% mutate(age_young = age >= 20 & age <= 34),
  cluster = ~ field
)

# IV split by age group
hz_iv_young <- feols(
  log_income_z ~ age_z + age_sq_z + education_years_z + hours_worked_z +
    hours_sq_z + partner + geslacht | field |
    share_int_agg_z ~ Z_visa_field_agg_z,
  data = young_z, cluster = ~ field
)
hz_iv_older <- feols(
  log_income_z ~ age_z + age_sq_z + education_years_z + hours_worked_z +
    hours_sq_z + partner + geslacht | field |
    share_int_agg_z ~ Z_visa_field_agg_z,
  data = older_z, cluster = ~ field
)

# First-stage F-stats (split by age group)
fs_young_z <- feols(
  share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z + education_years_z +
    hours_worked_z + hours_sq_z + partner + geslacht | field,
  data = young_z, cluster = ~ field
)
fs_older_z <- feols(
  share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z + education_years_z +
    hours_worked_z + hours_sq_z + partner + geslacht | field,
  data = older_z, cluster = ~ field
)

f_young_z <- round(wald(fs_young_z, 'Z_visa_field_agg_z', print = FALSE)$stat, 1)
f_older_z <- round(wald(fs_older_z, 'Z_visa_field_agg_z', print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel H: Age Group Effects (20-34 vs 35+) — Standardized}\\
\\textit{OLS with interaction; IV estimated separately by age group.}\\
\\textit{Age subsample F-stats: 20-34 = ', f_young_z, ', 35+ = ', f_older_z, '.}\\
\n')

etable(hz_ols_interact, hz_iv_young, hz_iv_older,
       se.below = TRUE,
       dict = c(share_int_agg_z = 'Intl. Share (z)',
                'share_int_agg_z:age_youngTRUE' = 'Intl. Share (z) × Age 20-34'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('OLS (interact)', 'IV: Age 20-34', 'IV: Age 35+'),
       extralines = list('F-stat (1st stage)' = c('', f_young_z, f_older_z)))
```

```{r regressions-z-lag}
# -- Panel E-z: Lagged IV specifications (standardized) ---------------------

lags_z <- est_data_z %>%
  group_by(field, survey_year) %>%
  summarise(Z_visa_z = first(Z_visa_field_agg_z), .groups = 'drop')

est_data_lag_z <- est_data_z
for (L in 1:3) {
  lag_df <- lags_z
  lag_df$survey_year <- lag_df$survey_year + L
  names(lag_df)[3] <- paste0('Z_visa_lag', L, '_z')
  est_data_lag_z <- est_data_lag_z %>% left_join(lag_df, by = c('field', 'survey_year'))
}

# OLS + IV pairs
ols_contemp_z <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                         education_years_z + hours_worked_z + hours_sq_z +
                         partner + geslacht | field,
                       data = est_data_lag_z, cluster = ~ field)
iv_contemp_z  <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                         hours_worked_z + hours_sq_z + partner + geslacht | field |
                         share_int_agg_z ~ Z_visa_field_agg_z,
                       data = est_data_lag_z, cluster = ~ field)

ols_lag1_z <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field,
                    data = est_data_lag_z, cluster = ~ field)
iv_lag1_z  <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                      hours_worked_z + hours_sq_z + partner + geslacht | field |
                      share_int_agg_z ~ Z_visa_lag1_z,
                    data = est_data_lag_z, cluster = ~ field)

ols_lag2_z <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field,
                    data = est_data_lag_z, cluster = ~ field)
iv_lag2_z  <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                      hours_worked_z + hours_sq_z + partner + geslacht | field |
                      share_int_agg_z ~ Z_visa_lag2_z,
                    data = est_data_lag_z, cluster = ~ field)

ols_lag3_z <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + geslacht | field,
                    data = est_data_lag_z, cluster = ~ field)
iv_lag3_z  <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                      hours_worked_z + hours_sq_z + partner + geslacht | field |
                      share_int_agg_z ~ Z_visa_lag3_z,
                    data = est_data_lag_z, cluster = ~ field)

ols_all_z <- feols(log_income_z ~ share_int_agg_z + age_z + age_sq_z +
                     education_years_z + hours_worked_z + hours_sq_z +
                     partner + geslacht | field,
                   data = est_data_lag_z, cluster = ~ field)
iv_all_z  <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                     hours_worked_z + hours_sq_z + partner + geslacht | field |
                     share_int_agg_z ~ Z_visa_lag1_z + Z_visa_lag2_z + Z_visa_lag3_z,
                   data = est_data_lag_z, cluster = ~ field)

fs_contemp_z <- feols(share_int_agg_z ~ Z_visa_field_agg_z + age_z + age_sq_z + education_years_z + hours_worked_z + hours_sq_z + partner + geslacht | field,
                       data = est_data_lag_z, cluster = ~ field)
fs_lag1_z <- feols(share_int_agg_z ~ Z_visa_lag1_z + age_z + age_sq_z + education_years_z + hours_worked_z + hours_sq_z + partner + geslacht | field,
                    data = est_data_lag_z, cluster = ~ field)
fs_lag2_z <- feols(share_int_agg_z ~ Z_visa_lag2_z + age_z + age_sq_z + education_years_z + hours_worked_z + hours_sq_z + partner + geslacht | field,
                    data = est_data_lag_z, cluster = ~ field)
fs_lag3_z <- feols(share_int_agg_z ~ Z_visa_lag3_z + age_z + age_sq_z + education_years_z + hours_worked_z + hours_sq_z + partner + geslacht | field,
                    data = est_data_lag_z, cluster = ~ field)
fs_all_z <- feols(share_int_agg_z ~ Z_visa_lag1_z + Z_visa_lag2_z + Z_visa_lag3_z + age_z + age_sq_z + education_years_z + hours_worked_z + hours_sq_z + partner + geslacht | field,
                   data = est_data_lag_z, cluster = ~ field)

f_contemp_z <- round(wald(fs_contemp_z, 'Z_visa_field_agg_z', print = FALSE)$stat, 1)
f_lag1_z <- round(wald(fs_lag1_z, 'Z_visa_lag1_z', print = FALSE)$stat, 1)
f_lag2_z <- round(wald(fs_lag2_z, 'Z_visa_lag2_z', print = FALSE)$stat, 1)
f_lag3_z <- round(wald(fs_lag3_z, 'Z_visa_lag3_z', print = FALSE)$stat, 1)
f_all_z <- round(wald(fs_all_z, c('Z_visa_lag1_z', 'Z_visa_lag2_z', 'Z_visa_lag3_z'), print = FALSE)$stat, 1)

cat('\n\n\\textbf{Panel E: Lagged IV Specifications — Standardized}\\n\\textit{Each pair: OLS | IV. Long-term = lags 1-3 jointly.}\\n
')

etable(ols_contemp_z, iv_contemp_z, ols_lag1_z, iv_lag1_z, ols_lag2_z, iv_lag2_z, ols_lag3_z, iv_lag3_z, ols_all_z, iv_all_z,
       se.below = TRUE,
       dict = c(share_int_agg_z = 'Intl. Share (z)'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('Contemp', '', 'Lag 1', '', 'Lag 2', '', 'Lag 3', '', 'Long-term', ''),
       subtitles = c('OLS', 'IV', 'OLS', 'IV', 'OLS', 'IV', 'OLS', 'IV', 'OLS', 'IV'),
       extralines = list('F-stat (1st stage)' =
                          c('', f_contemp_z, '', f_lag1_z, '', f_lag2_z, '', f_lag3_z, '', f_all_z)))
```

```{r placebo-low-share, results='asis'}
# -- Exclusion Restriction: Placebo test on bottom two fields ------------------
# If the IV effect operates through international graduate labour supply,
# it should be concentrated in fields with above-median international shares.
# The two fields with the lowest average international share serve as a placebo:
# the first stage should be weaker and the reduced-form / IV should be near zero.

# Identify the two fields with the lowest mean share_int_agg
field_means <- est_data %>%
  group_by(field) %>%
  summarise(mean_share = mean(share_int_agg, na.rm = TRUE), .groups = 'drop') %>%
  arrange(mean_share)
bottom2_fields <- field_means$field[1:2]

d_low  <- est_data %>% filter(field %in% bottom2_fields)
d_high <- est_data %>% filter(!field %in% bottom2_fields)

cat(sprintf('Bottom 2 fields by mean share_int_agg: %s\n',
            paste(bottom2_fields, collapse = ', ')))
cat(sprintf('Bottom-two-field subsample: %d obs\n', nrow(d_low)))
cat(sprintf('Remaining fields: %d obs\n', nrow(d_high)))

# First stages
fs_placebo_low  <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq +
                           education_years + hours_worked + hours_sq +
                           partner + geslacht | field,
                         data = d_low, cluster = ~ field)
fs_placebo_high <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq +
                           education_years + hours_worked + hours_sq +
                           partner + geslacht | field,
                         data = d_high, cluster = ~ field)

f_placebo_low  <- round(wald(fs_placebo_low,  'Z_visa_field_agg', print = FALSE)$stat, 1)
f_placebo_high <- round(wald(fs_placebo_high, 'Z_visa_field_agg', print = FALSE)$stat, 1)

# IV
iv_placebo_low  <- feols(log_income ~ age + age_sq + education_years +
                           hours_worked + hours_sq + partner + geslacht | field |
                           share_int_agg ~ Z_visa_field_agg,
                         data = d_low, cluster = ~ field)
iv_placebo_high <- feols(log_income ~ age + age_sq + education_years +
                           hours_worked + hours_sq + partner + geslacht | field |
                           share_int_agg ~ Z_visa_field_agg,
                         data = d_high, cluster = ~ field)

# Reduced form (Z → wages directly)
rf_placebo_low  <- feols(log_income ~ Z_visa_field_agg + age + age_sq +
                           education_years + hours_worked + hours_sq +
                           partner + geslacht | field,
                         data = d_low, cluster = ~ field)
rf_placebo_high <- feols(log_income ~ Z_visa_field_agg + age + age_sq +
                           education_years + hours_worked + hours_sq +
                           partner + geslacht | field,
                         data = d_high, cluster = ~ field)

# Anderson-Rubin test (low-share subsample)
# The AR test is robust to weak instruments. Under H0: β_IV = 0, the
# reduced-form F-stat on Z follows an F(1, N - K - 1) distribution,
# where K is the number of regressors (controls + FE dummies).
# We use wald() on the reduced form to get the AR statistic.
ar_low <- wald(rf_placebo_low, 'Z_visa_field_agg', print = FALSE)
ar_low_stat <- ar_low$stat
ar_low_df1  <- ar_low$df1
ar_low_df2  <- ar_low$df2
ar_low_pval <- pf(ar_low_stat, ar_low_df1, ar_low_df2, lower.tail = FALSE)

# Wild cluster bootstrap AR test: robust with few clusters (bottom two fields)
# The Webb (2014) six-point distribution (wcv13) is recommended for < 10 clusters.
# We implement a manual wild cluster bootstrap: under H0 (beta_Z = 0), generate
# y* from the restricted model, multiply residuals by Webb (2014) six-point weights
# at the cluster level, re-estimate the unrestricted model, and collect the F-stat.
rf_restricted <- feols(log_income ~ age + age_sq + education_years +
                         hours_worked + hours_sq + partner + geslacht | field,
                       data = d_low, cluster = ~ field)
e_resid <- resid(rf_restricted)
clusters <- unique(d_low$field)
n_clusters <- length(clusters)
set.seed(20240617)
B <- 9999
boot_F <- numeric(B)
for (b in 1:B) {
  # Webb (2014) six-point distribution: ±sqrt(3/2), ±1, ±sqrt(1/2) with equal prob
  w <- sample(c(-sqrt(3/2), -1, -sqrt(1/2), sqrt(1/2), 1, sqrt(3/2)),
              n_clusters, replace = TRUE)
  names(w) <- clusters
  y_star <- fitted(rf_restricted) + e_resid * w[as.character(d_low$field)]
  d_boot <- d_low
  d_boot$y_star <- y_star
  rf_boot <- feols(y_star ~ Z_visa_field_agg + age + age_sq + education_years +
                     hours_worked + hours_sq + partner + geslacht | field,
                   data = d_boot, cluster = ~ field)
  boot_F[b] <- wald(rf_boot, "Z_visa_field_agg", print = FALSE)$stat
}
ar_boot_pval <- mean(boot_F >= ar_low_stat)

cat('\n\n\\textbf{Exclusion Restriction Defence 1: Placebo by International Share}\\
\\textit{Bottom two fields / remaining fields of share\\_int\\_agg. If the instrument works through
the international graduate channel, effects should concentrate in high-share fields.}\\
\\textit{Bottom-two-field F = ', f_placebo_low, ', Upper F = ', f_placebo_high, '.}\\
\\textit{AR test (bottom-two-field subsample): asymptotic F(', ar_low_df1, ', ', ar_low_df2, ') = ',
      round(ar_low_stat, 3), ', p < 0.001;
      wild cluster bootstrap p = ', round(ar_boot_pval, 4), ' (B = ', B, ').}\\
\n')
etable(rf_placebo_low, rf_placebo_high, iv_placebo_low, iv_placebo_high,
       se.below = TRUE,
       dict = c(share_int_agg = 'Intl. Share', Z_visa_field_agg = 'Visa Instr.'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('RF: Bottom 2', 'RF: Rest', 'IV: Bottom 2', 'IV: Rest'),
       extralines = list('F-stat (1st stage)' =
                          c('', '', f_placebo_low, f_placebo_high)))

```{r field-trends, results='asis'}
# -- Exclusion Restriction: Field-specific linear trends -----------------------
# If the exclusion restriction is violated because the instrument correlates
# with field-level trade or FDI trends, adding field-specific linear trends
# should absorb that channel. If the IV estimate survives, the concern is
# attenuated.

# Contemporaneous IV with field-specific trends (full sample)
iv_trend <- feols(log_income ~ age + age_sq + education_years +
                    hours_worked + hours_sq + partner + geslacht |
                    field + field[survey_year] |
                    share_int_agg ~ Z_visa_field_agg,
                  data = est_data, cluster = ~ field)

# First stage with trends
fs_trend <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq +
                    education_years + hours_worked + hours_sq +
                    partner + geslacht | field + field[survey_year],
                  data = est_data, cluster = ~ field)

f_trend <- round(wald(fs_trend, 'Z_visa_field_agg', print = FALSE)$stat, 1)

# STEM / Non-STEM with trends
iv_trend_stem  <- feols(log_income ~ age + age_sq + education_years +
                          hours_worked + hours_sq + partner + geslacht |
                          field + field[survey_year] |
                          share_int_agg ~ Z_visa_field_agg,
                        data = est_data, cluster = ~ field, subset = ~ stem)
iv_trend_nstem <- feols(log_income ~ age + age_sq + education_years +
                          hours_worked + hours_sq + partner + geslacht |
                          field + field[survey_year] |
                          share_int_agg ~ Z_visa_field_agg,
                        data = est_data, cluster = ~ field, subset = ~ !stem)

fs_trend_stem  <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq +
                          education_years + hours_worked + hours_sq +
                          partner + geslacht | field + field[survey_year],
                        data = est_data, cluster = ~ field, subset = ~ stem)
fs_trend_nstem <- feols(share_int_agg ~ Z_visa_field_agg + age + age_sq +
                          education_years + hours_worked + hours_sq +
                          partner + geslacht | field + field[survey_year],
                        data = est_data, cluster = ~ field, subset = ~ !stem)

f_trend_stem  <- round(wald(fs_trend_stem,  'Z_visa_field_agg', print = FALSE)$stat, 1)
f_trend_nstem <- round(wald(fs_trend_nstem, 'Z_visa_field_agg', print = FALSE)$stat, 1)

cat('\n\n\\textbf{Exclusion Restriction Defence 2: Field-Specific Linear Trends}\\
\\textit{Field-specific linear trends on survey\\_year absorb differential trends
that could correlate with trade or FDI linkages.}\\
\\textit{F-stats: Full = ', f_trend, ', STEM = ', f_trend_stem,
', Non-STEM = ', f_trend_nstem, '.}\\
\n')

etable(iv_trend, iv_trend_stem, iv_trend_nstem,
       se.below = TRUE,
       dict = c(share_int_agg = 'Intl. Share'),
       digits = 3, digits.stats = 3,
       fitstat = c('n', 'r2'),
       headers = c('Full', 'STEM', 'Non-STEM'),
       extralines = list('F-stat (1st stage)' =
                          c(f_trend, f_trend_stem, f_trend_nstem)))
```
