---
title: "International Student Presence and Dutch Graduate Labor Market Outcomes"
subtitle: "A Shift-Share Instrumental Variables Approach"
author: "Redon Ismailaga"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    number_sections: true
bibliography: references.bib
csl: apa.csl
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{float}
---

```{r setup, include=FALSE}
# -- Global knitr options ------------------------------------------------------
# echo = FALSE:   don't print the R code itself into the PDF --- only show outputs
# warning = FALSE: suppress R warnings (they'd clutter the PDF with irrelevant noise)
# message = FALSE: suppress package-loading messages (e.g. "Attaching package: dplyr")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r packages}
pacman::p_load(`pacman`)
p_load(tidyverse); p_load(readxl); p_load(janitor); p_load(fixest)
p_load(modelsummary); p_load(kableExtra); p_load(WDI)
p_load(rt); p_load(hunspell)
```

# 1. Data Loading

## 1.1 LISS Panel

```{r liss-load}
# -- LISS Work & Schooling module (2020-2024) ----------------------------------
# The LISS panel is a longitudinal Dutch household survey. The "Work and Schooling"
# module records each respondent's occupation, field of study, education level,
# and working hours. Each yearly file has its own column naming convention.
# Early waves (2020, 2021) use comma-separated format; later waves (2022-2024)
# use semicolon-separated format. check.names = FALSE prevents R from auto-converting
# column names (e.g. spaces to dots).

# 2020 wave: comma-separated (sep = "," is default), 1.0p = version 1.0 public
LISS2020WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw20m_EN_1.0p.csv", check.names = FALSE)
# 2021 wave: also comma-separated
LISS2021WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw21n_EN_1.0p.csv", check.names = FALSE)
# 2022 wave: semicolon-separated (sep = ";")
LISS2022WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw22o_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2023 wave: semicolon-separated
LISS2023WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw23p_EN_2.0p.csv", sep = ";", check.names = FALSE)
# 2024 wave: semicolon-separated
LISS2024WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw24q_EN_1.0p.csv", sep = ";", check.names = FALSE)

# -- LISS Economic Situation / Income module (2020-2024) -----------------------
# The "Economic Situation: Income" module records gross monthly wages, partner status,
# and other income-related variables. Same sep pattern as the Work & Schooling module.

# 2020 income data: comma-separated
LISS2020ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci20m_EN_1.0p.csv", check.names = FALSE)
# 2021 income data: comma-separated
LISS2021ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci21n_EN_1.0p.csv", check.names = FALSE)
# 2022 income data: semicolon-separated
LISS2022ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci22o_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2023 income data: semicolon-separated
LISS2023ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci23p_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2024 income data: semicolon-separated
LISS2024ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci24q_EN_1.0p.csv", sep = ";", check.names = FALSE)
```

## 1.2 IND Visa

```{r visa-load}
# -- IND Schengen visa statistics (2020-2024) ----------------------------------
# The Immigration and Naturalisation Service (IND) publishes yearly Schengen visa
# statistics by consulate/country. Each file is an Excel workbook (.xlsx) with
# a worksheet named "Data for consulates" containing acceptance/rejection rates
# by applicant nationality.
visa_files <- c(
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/Visa-statistics-for-consulates-2020_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2021-consulates-schengen-visa-stats.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/Visa statistics for consulates in 2022_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2023-schengen-visa-statistics-consulates_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2024-schengen-visa-statistics-consulates_en.xlsx"
)
```

## 1.3 DUO Graduates

```{r duo-load}
# -- DUO graduate counts by institution / field / gender -----------------------
# DUO (Dienst Uitvoering Onderwijs) publishes yearly graduate counts broken down
# by institution, field of study (onderdeel), gender, and education level.
# Two files: WO (university) and HBO (professional bachelor).

GradsUni <- read_delim(
  "../../../Data/DUO Graduates by Gender, Institution, Field of  Study, University and Professional level AND The intitution's adresses/05.-gediplomeerden-wo-2025.csv",
  delim = ";"
)

GradsProf <- read_delim(
  "../../../Data/DUO Graduates by Gender, Institution, Field of  Study, University and Professional level AND The intitution's adresses/05.-gediplomeerden-hbo-2024.csv",
  delim = ";"
)
# Institution address metadata: used to geolocate institutions to provinces/cities
InstitutionAddys <- read.csv("../../../Data/DUO Graduates by Gender, Institution, Field of  Study, University and Professional level AND The intitution's adresses/Higher education institutes location.csv", sep = ";")
```

## 1.4 CBS Nationality Data

```{r cbs-load}
# -- CBS temporary residence permits by nationality (2010) ---------------------
# CBS (Centraal Bureau voor de Statistiek) publishes counts of temporary residence
# permits (verblijfsvergunningen voor bepaalde tijd) by nationality and year.
# This data is used to construct the baseline nationality shares for the shift-share
# instrument. The file name includes a timestamp from when it was downloaded.
EthnicalMakeup <- read.csv("../../../Data/CBS Incoming Student Ethnical Makeup/Verblijfsvergunningen_voor_bepaalde_tijd_19052026_175309.csv", sep = ";")
```

# 2. Process IND Visa

```{r process-visa}
# -- IND Schengen visa acceptance rates (2020-2024) ---------------------------
# Load each year's Excel file, keep only Netherlands consulates, and extract
# acceptance rates by applicant nationality. Column names are selected after
# inspecting the raw datasets.

years <- 2020:2024
visa_list <- list()

for (i in seq_along(visa_files)) {
  if (!file.exists(visa_files[i])) next

  df <- read_excel(visa_files[i], sheet = "Data for consulates") %>%
    clean_names() %>%
    mutate(year = years[i])

  # Filter to Netherlands consulates and select key columns
  # (adjust the exact column names below after viewing the data in R)
  nl <- df %>%
    filter(schengen_state == "Netherlands") %>%
    select(
      country = country_where_consulate_is_located,
      not_issued_rate = not_issued_rate_for_uniform_visas,
      year
    ) %>%
    mutate(
      acceptance_rate = 1 - not_issued_rate,
      country = str_to_upper(str_trim(country))
    )

  if (nrow(nl) > 0) visa_list[[i]] <- nl
}

Consulates <- bind_rows(visa_list)
```

# 3. Process DUO -> Field-Level International Share

```{r process-duo}
# -- Stack WO and HBO graduate data --------------------------------------------
DUO_all <- bind_rows(GradsUni %>% mutate(level = "WO"),
                     GradsProf %>% mutate(level = "HBO"))

names(DUO_all) <- c("provincie", "gem_nummer", "gem_naam", "soort_instelling",
                    "instellingscode", "instellingsnaam", "onderdeel", "subonderdeel",
                    "opleidingscode", "opleidingsnaam", "opleidingsvorm", "soort_diploma",
                    "geslacht", "y2020", "y2021", "y2022", "y2023", "y2024", "level")

# -- Reshape from wide to long format ------------------------------------------
DUO_long <- DUO_all %>%
  pivot_longer(cols = y2020:y2024, names_to = "year_chr", values_to = "graduates_raw") %>%
  mutate(
    year = as.integer(str_replace(year_chr, "y", "")),
    graduates_suppressed = str_detect(as.character(graduates_raw), "<"),
    graduates = ifelse(graduates_suppressed, NA_real_,
                       as.numeric(str_replace_all(as.character(graduates_raw), "[^0-9-]", ""))),
    onderdeel = str_to_lower(str_trim(onderdeel))
  )

# -- CBS International senior share (actual herkomst, non-EU) ----------------
# Switched to CBS senior-year data by field and origin (non-EU only, per CBS permits).
# This replaces the old programme-name keyword classification.

base <- "../../../Data/CBS Incoming Student Ethnical Makeup"

intl <- read_delim(file.path(base, "Internationals by gender, Fos, Senior year.csv"), delim = ";", show_col_types = FALSE)
total <- read_delim(file.path(base, "Dutch people by FoS, gender, senior year.csv"), delim = ";", show_col_types = FALSE)

non_eu <- c("Buiten-Europa (excl. 5 grote herkomst...)", "Indonesië", "Marokko", "Nederlandse Cariben", "Suriname", "Turkije")

intl_seniors <- intl %>%
  filter(Herkomstland %in% non_eu) %>%
  group_by(Studierichting, Perioden, Onderwijssoort) %>%
  summarise(intl_seniors = sum(as.numeric(`Totaal ingeschrevenen (aantal)`), na.rm = TRUE), .groups = "drop")

total_seniors <- total %>%
  group_by(Studierichting, Perioden, Onderwijssoort) %>%
  summarise(total_seniors = sum(as.numeric(`Ouderejaarsstudenten (aantal)`), na.rm = TRUE), .groups = "drop")

cbs_share <- total_seniors %>%
  left_join(intl_seniors, by = c("Studierichting", "Perioden", "Onderwijssoort")) %>%
  mutate(
    share_int = ifelse(total_seniors > 0, intl_seniors / total_seniors, NA_real_),
    year = as.integer(str_sub(Perioden, 1, 4))
  )

# Note: Studierichting (e.g. "01 Onderwijs") will need a mapping to "onderdeel" in DUO_long
# for joining with graduate totals if required downstream. The old is_international flag
# and keyword lists are no longer used for the main share.
```

# 4. Process CBS -> Nationality Shares (2010, per AEA 10-year lag)

```{r process-cbs}
# -- Extract pre-period nationality shares -------------------------------------
# The shift-share instrument uses pre-period (2010) nationality shares as weights.
# We use year 2010 --- a full decade before the outcome period (2020-2024) --- to
# ensure the shares are exogenous to contemporaneous shocks (AEA convention).
cbs_raw <- EthnicalMakeup                                   # Raw CBS data
names(cbs_raw) <- c("nationaliteit", "perioden", "permits_raw")  # Rename to English

# -- Clean and parse ----------------------------------------------------------
cbs_shares <- cbs_raw %>%
  mutate(
    nationality = str_to_title(str_trim(nationaliteit)),    # Standardise nationality name (Title Case)
    permits = as.numeric(str_trim(permits_raw)),             # Convert permit count to numeric
    year = as.integer(str_extract(as.character(perioden), "^[0-9]{4}"))  # Extract year from period string
  ) %>%
  filter(!is.na(permits), permits > 0, !is.na(year))        # Drop missing/zero counts and unparseable years

# -- Aggregate to 2010 nationality shares -------------------------------------
# Keep only 2010, sum permits per nationality, compute share of total
nat_shares <- cbs_shares %>%
  filter(year == 2010) %>%                                   # Pre-period: 2010 only
  group_by(nationality) %>%                                  # Group by nationality name
  summarise(permits = sum(permits, na.rm = TRUE), .groups = "drop") %>%  # Total permits per nationality
  mutate(share = permits / sum(permits)) %>%                   # Each nationality's share of ALL permits
  arrange(desc(share))                                         # Sort by share (largest first)
```

# 5. Construct Instruments

## 5.1 Nationality Name Mapping

```{r name-map}
# -- Standardise nationality names across datasets -----------------------------
# CBS uses Dutch nationality names (e.g. "Chinees"), while IND visa data uses
# English country names. This case_when block maps the Dutch names to standardised
# English uppercase names so the two datasets can be merged.
nat_shares <- nat_shares %>%
  mutate(country_std = case_when(
    str_detect(nationality, "Chinees|China")          ~ "CHINA",
    str_detect(nationality, "Indiaas|India")          ~ "INDIA",
    str_detect(nationality, "Turks|Turkse|Turkey")    ~ "TURKEY",
    str_detect(nationality, "Amerikaans|United States|USA") ~ "UNITED STATES",
    str_detect(nationality, "Brits|United Kingdom")   ~ "UNITED KINGDOM",
    str_detect(nationality, "Duits|Germany")          ~ "GERMANY",
    str_detect(nationality, "Frans|France")           ~ "FRANCE",
    str_detect(nationality, "Italiaans|Italy")        ~ "ITALY",
    str_detect(nationality, "Spaans|Spain")           ~ "SPAIN",
    str_detect(nationality, "Pools|Poland")           ~ "POLAND",
    str_detect(nationality, "Roemeens|Romania")       ~ "ROMANIA",
    str_detect(nationality, "Bulgaars|Bulgaria")      ~ "BULGARIA",
    str_detect(nationality, "Grieks|Greece")          ~ "GREECE",
    str_detect(nationality, "Portugees|Portugal")     ~ "PORTUGAL",
    str_detect(nationality, "Hongaars|Hungary")       ~ "HUNGARY",
    str_detect(nationality, "Litouws|Lithuania")      ~ "LITHUANIA",
    str_detect(nationality, "Indonesisch|Indonesia")  ~ "INDONESIA",
    str_detect(nationality, "Russisch|Russia")        ~ "RUSSIA",
    str_detect(nationality, "Braziliaans|Brazil")     ~ "BRAZIL",
    str_detect(nationality, "Japans|Japan")           ~ "JAPAN",
    str_detect(nationality, "Koreaans|South Korea")   ~ "SOUTH KOREA",
    str_detect(nationality, "Mexicaans|Mexico")       ~ "MEXICO",
    TRUE ~ str_to_upper(nationality)                   # For unmatched names, just uppercase the original
  ))
```

## 5.2 Instrument 1: Visa Acceptance Rates (all countries)

```{r instr-visa}
# -- Shift-share instrument: Schengen visa acceptance x nationality share ------
# Inner join: keep only countries that appear in BOTH the IND visa data AND
# the CBS nationality shares. This gives us, for each matched country, the visa
# acceptance rate and its pre-period share of Dutch residence permits.
visa_merged <- Consulates %>%
  inner_join(nat_shares, by = c("country" = "country_std")) %>%
  filter(!is.na(acceptance_rate), !is.na(share))

# Compute the year-level instrument Z_visa as:
#   Z_visa[t] = sum_over_countries( share_c x acceptance_rate_c,t )
# This is a Bartik-style average: 2010 share weights x time-varying visa acceptance.
Z_visa_year <- visa_merged %>%
  group_by(year) %>%
  summarise(Z_visa = sum(share * acceptance_rate, na.rm = TRUE), .groups = "drop")
```

## 5.3 Instrument 2: Youth Unemployment (low-trade countries only)

```{r instr-youth}
# -- Second instrument: youth unemployment in non-trade-linked origin countries -
# Exclusion restriction argument: youth unemployment in these countries affects
# Dutch graduate wages ONLY through the student-mobility channel (not through trade),
# because the Netherlands has negligible trade exposure to these economies.
low_trade <- c("CHINA", "INDIA", "INDONESIA", "RUSSIA", "BRAZIL", "MEXICO", "JAPAN", "SOUTH KOREA")

# -- Map standardised country names -> World Bank WDI country codes -------------
# WDI (World Development Indicators) uses 2-letter ISO codes for API queries.
wdi_codes <- c("CHINA" = "CN", "INDIA" = "IN", "INDONESIA" = "ID", "RUSSIA" = "RU",
               "BRAZIL" = "BR", "MEXICO" = "MX", "JAPAN" = "JP", "SOUTH KOREA" = "KR",
               "TURKEY" = "TR", "UNITED STATES" = "US", "UNITED KINGDOM" = "GB",
               "GERMANY" = "DE", "FRANCE" = "FR", "ITALY" = "IT", "SPAIN" = "ES",
               "POLAND" = "PL", "ROMANIA" = "RO", "BULGARIA" = "BG", "GREECE" = "GR",
               "PORTUGAL" = "PT", "HUNGARY" = "HU", "LITHUANIA" = "LT", "BELGIUM" = "BE")

# -- Download youth unemployment (ages 15-24) from the World Bank API ----------
# Indicator SL.UEM.1524.ZS = "Unemployment, youth total (% of total labor force ages 15-24)"
wdi_data <- WDI(indicator = "SL.UEM.1524.ZS",
                country = wdi_codes, start = 2010, end = 2024) %>%
  as_tibble() %>%
  rename(youth_unemp = SL.UEM.1524.ZS) %>%   # Rename to a shorter readable name
  filter(!is.na(youth_unemp))                  # Drop any years with missing data

# -- Map WDI codes back to our standardised country names ---------------------
country_rev <- setNames(names(wdi_codes), wdi_codes)  # Reverse lookup: "CN" -> "CHINA"
wdi_data <- wdi_data %>%
  mutate(country_std = country_rev[iso2c]) %>%         # Add the standardised name column
  filter(!is.na(country_std))                           # Drop if no match found

# -- Re-normalise shares to SUM TO 1 within the low-trade subset --------------
# Only low-trade countries enter this instrument, so we need conditional shares
nat_lowtrade <- nat_shares %>% filter(country_std %in% low_trade) %>%
  mutate(share_lt = share / sum(share))

# -- Construct Z_youth: weighted average youth unemployment --------------------
#   Z_youth[t] = sum_over_low_trade( share_lt_c x youth_unemp_c,t )
# The share_lt weights are time-invariant (2010), youth_unemp varies by year.
Z_youth_year <- wdi_data %>%
  filter(country_std %in% low_trade) %>%
  inner_join(nat_lowtrade, by = "country_std") %>%
  group_by(year) %>%
  summarise(Z_youth = sum(share_lt * youth_unemp, na.rm = TRUE), .groups = "drop") %>%
  # Normalize Z_youth to [0,1] so it's on the same scale as other variables
  mutate(Z_youth = (Z_youth - min(Z_youth, na.rm = TRUE)) /
                   (max(Z_youth, na.rm = TRUE) - min(Z_youth, na.rm = TRUE)))
```

## 5.4 Instrument 3: University-Level Non-EU Enrollment x Visa Rate (Pending Data)

**Rationale:** Universities with more non-EU students are more exposed to Schengen visa
policy. Pre-period non-EU enrollment shares x aggregate non-EU visa rates give
a institution-level shock that rolls up to fields via DUO's graduate composition.

**Data needed (request from DUO via supervisor):**
A CSV saved to `../../../Data/DUO Non-EU Enrollment by Institution/`:

```
instellingscode,instellingsnaam,year,total_students,non_eu_students
01,University A,2015,25000,3000
01,University A,2016,25500,3200
...
```

A pre-period snapshot (ideally 2015) is sufficient for the baseline share ---
time-varying enrollment is not essential for the shift-share design.

```{r instr-enrollment, eval=FALSE}
# -- Uncomment and run when DUO data arrives --------------------------
#
# (1) Load enrollment data
# enroll <- read.csv(
#   "../../../Data/DUO Non-EU Enrollment by Institution/non_eu_enrollment.csv",
#   sep = ";", check.names = FALSE
# ) %>% clean_names()
#
# (2) Fix baseline shares (pre-period, e.g. 2015)
# enroll_bl <- enroll %>%
#   filter(year == min(year)) %>%
#   mutate(share_nonEU = non_eu_students / total_students) %>%
#   select(instellingscode, instellingsnaam, share_nonEU)
#
# (3) Aggregate non-EU visa rate per year (from IND data)
# visa_neu_agg <- Consulates %>%
#   inner_join(nat_shares, by = c("country" = "country_std")) %>%
#   group_by(year) %>%
#   summarise(Z_neu_visa = weighted.mean(acceptance_rate, share, na.rm = TRUE),
#             .groups = "drop")
#
# (4) Per university-year instrument
# uni_instr <- expand_grid(instellingscode = enroll_bl$instellingscode,
#                           year = 2020:2024) %>%
#   left_join(enroll_bl, by = "instellingscode") %>%
#   left_join(visa_neu_agg, by = "year") %>%
#   mutate(Z_uni = share_nonEU * Z_neu_visa)
#
# (5) Roll up to field level via DUO graduate composition
# Use pre-period graduate shares as weights: what fraction of field f's
# graduates come from each university?
# grad_weights <- DUO_long %>%
#   filter(year == min(year)) %>%
#   group_by(onderdeel, instellingsnaam) %>%
#   summarise(grads = sum(graduates, na.rm = TRUE), .groups = "drop_last") %>%
#   mutate(weight = grads / sum(grads, na.rm = TRUE)) %>%
#   ungroup() %>% select(-grads)
#
# Z_enroll_year <- uni_instr %>%
#   inner_join(grad_weights, by = "instellingsnaam") %>%
#   group_by(onderdeel, year) %>%
#   summarise(Z_enroll = sum(weight * Z_uni, na.rm = TRUE), .groups = "drop")
#
# cat("[OK] Enrollment instrument built:",
#     n_distinct(Z_enroll_year$onderdeel), "fields x",
#     n_distinct(Z_enroll_year$year), "years\n")
```

# 6. Process LISS -> Estimation Sample

```{r process-liss-work}
# -- LISS field-of-study code -> DUO field name map -----------------------------
# LISS uses numeric codes (011-027) for fields of study. We map each code to the
# corresponding DUO field name (onderdeel) and a boolean STEM indicator.
# Codes 017-019 (natuur, techniek, landbouw) are classified as STEM.
field_map <- tibble(
  code = sprintf("%03d", 11:27),          # Generate codes "011", "012", ... "027"
  duo_field = c(NA, "onderwijs", "taal en cultuur", "taal en cultuur",   # 012=education, 013/014=language/culture
                "gedrag en maatschappij", "economie", "recht",            # 015=behaviour/society, 016=economics, 017=law
                "natuur", "techniek", "landbouw en natuurlijke omgeving", # 018=nature, 019=engineering, 020=agriculture
                "gezondheidszorg", "gedrag en maatschappij",              # 021=healthcare, 022=behaviour/society
                "gedrag en maatschappij", "techniek", "techniek",         # 023/024 also map to existing fields
                "recht", "sectoroverstijgend"),                           # 026=law, 027=interdisciplinary
  stem = c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,              # Codes 011-017 are non-STEM
           TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)  # 018-020 = STEM; rest non-STEM
)

# -- process_early(): parse LISS 2020-2021 (comma-separated, column-name-based) --
process_early <- function(df, yr) {
  # Column 1 is the respondent identifier (nomem_encr --- encrypted membership number)
  names(df)[1] <- "nomem_encr"

  # Columns 12-28 are the field-of-study binary indicators (one column per field code)
  field_cols <- names(df)[12:28]
  # The corresponding field codes (011-027) as zero-padded 3-character strings
  field_codes <- sprintf("%03d", 11:27)

  # Extract only the field-of-study columns into a matrix (preserving row order)
  mat <- df[, field_cols, drop = FALSE]
  # LISS codes 1 = "currently following" and 2 = "completed" -> both count as studying the field.
  # Convert each column to TRUE/FALSE: TRUE if the value is 1 or 2.
  mat[] <- lapply(mat, function(x) as.integer(x) %in% c(1, 2))

  # For each row (respondent), find the FIRST field column that is TRUE,
  # and return the corresponding field code. If none selected, return NA.
  sel <- apply(mat, 1, function(r) { idx <- which(r); if (length(idx) == 0) NA_character_ else field_codes[idx[1]] })

  # -- Education attainment mapping --------------------------------------------
  # LISS asks "What is the highest level of education completed with a diploma?"
  # We convert the text answer to TWO independent variables:
  #   1. education_years --- estimated total years of schooling (for Mincer equation)
  #   2. level --- WO (research university) vs HBO (applied sciences) classification
  # Both come from the SAME raw survey answer, but are derived SEPARATELY.
  #
  # Years mapping based on convention in published LISS-using papers
  # (Leuven, Oosterbeek & van der Klaauw 2013; Bozza 2022):
  #   Compulsory schooling starts at age 6 -> primary = 6 years (ages 6-12)
  edu_col <- grep("highest level.*education.*completed.*diploma", names(df), value = TRUE, ignore.case = TRUE)[1]
  edu_raw <- as.character(df[[edu_col]])
  # Initialise as NA for all rows
  edu_yrs <- rep(NA_real_, nrow(df))
  # Map LISS education categories to Dutch system years (convention from published literature):
  edu_yrs[grepl("did not complete|^no$|^none$", edu_raw, ignore.case = TRUE)] <- 0  # No diploma
  edu_yrs[grepl("^primary school", edu_raw, ignore.case = TRUE)] <- 6               # Basisschool (6 yrs from age 6)
  edu_yrs[grepl("vmbo|mulo|lower technical|^lbo|vglo", edu_raw, ignore.case = TRUE)] <- 10  # VMBO (6+4)
  edu_yrs[grepl("^havo|^hbs|^mms", edu_raw, ignore.case = TRUE)] <- 11              # HAVO (6+5)
  edu_yrs[grepl("^vwo|^gymnasium", edu_raw, ignore.case = TRUE)] <- 12              # VWO (6+6)
  edu_yrs[grepl("^mbo |^kmbo", edu_raw, ignore.case = TRUE)] <- 13                  # MBO (6+6+1)
  edu_yrs[grepl("^mbo-plus", edu_raw, ignore.case = TRUE)] <- 14                    # MBO+ (6+6+2)
  edu_yrs[grepl("^hbo", edu_raw, ignore.case = TRUE)] <- 16                         # HBO (VWO 12+4, via MBO 13+3)
  edu_yrs[grepl("academic.*bachelor", edu_raw, ignore.case = TRUE)] <- 15            # WO bachelor (VWO 12+3)
  edu_yrs[grepl("academic.*master", edu_raw, ignore.case = TRUE)] <- 17              # WO master (15+2)
  edu_yrs[grepl("^academic|^university", edu_raw, ignore.case = TRUE)] <- 15         # WO (unspecified = bachelor)
  edu_yrs[grepl("doctor|phd", edu_raw, ignore.case = TRUE)] <- 21                    # PhD (17+4)

  # -- Usual hours worked per week ---------------------------------------------
  # LISS asks "How many hours do you actually work on average?"
  hrs_col <- grep("actually work", names(df), value = TRUE, ignore.case = TRUE)[1]
  hours_worked <- if (!is.na(hrs_col)) as.numeric(df[[hrs_col]]) else NA_real_

  # -- Tenure (years with current employer) ------------------------------------
  # LISS asks for the year the respondent started working for their current employer.
  ten_col <- grep("year.*(current employer|enter into employment|start.*employment)", names(df), value = TRUE, ignore.case = TRUE)[1]
  start_year <- if (!is.na(ten_col)) as.numeric(df[[ten_col]]) else NA_real_
  # Tenure = survey year - start year, but only if the start year is plausible (> 1900, < survey year)
  tenure_years <- ifelse(!is.na(start_year) & start_year > 1900 & start_year < yr, yr - start_year, NA_real_)

  # Return a clean data frame with only the columns we need (no raw survey columns)
  data.frame(nomem_encr = as.character(df$nomem_encr), survey_year = yr,
             liss_code = sel, age = as.numeric(df[[6]]),        # Column 6 contains age
             education_years = edu_yrs,
             hours_worked = hours_worked,
             tenure_years = tenure_years,
             stringsAsFactors = FALSE)
}

# -- process_late(): parse LISS 2022-2024 (semicolon-separated, prefixed columns) --
process_late <- function(df, yr) {
  # Late waves use a prefixed naming scheme: cwXXyNNN where XX = year suffix, y = wave letter, NNN = variable code.
  # We auto-detect the wave letter by finding the age column (which always ends in "003").
  nm <- names(df)
  age_idx <- grep("003$", nm)[1]           # First column ending in "003" = age column
  wave_letter <- substr(nm[age_idx], 5, 5)  # Extract wave letter (e.g., "p" from "cw23p003")

  # Prefix = "cw" + last 2 digits of year + wave letter -> e.g. "cw23p"
  prefix <- paste0("cw", substr(yr, 3, 4), wave_letter)
  # Age column = prefix + "003"
  age_col <- paste0(prefix, "003")

  # -- Field-of-study detection ------------------------------------------------
  # Field columns follow the pattern cwXXy011 through cwXXy027 (codes 011-027).
  # Regex: "^" + prefix + "0[12][0-9]$" matches "cw23p011", "cw23p012", ..., "cw23p027"
  fcols <- nm[grep(paste0("^", prefix, "0[12][0-9]$"), nm)]
  # Extract the last 3 digits of each column name to get the field code (e.g. "011")
  fcodes <- sub(".*([0-9]{3})$", "\\1", fcols)

  # Loop over respondents row-by-row (each respondent can select only one field)
  nr <- nrow(df)
  sel <- rep(NA_character_, nr)             # Initialise field selection as NA
  for (i in seq_len(nr)) {
    # Get the values for all field columns for this respondent
    vals <- as.integer(df[i, fcols])
    # Find which codes have value 1 (currently following) or 2 (completed)
    idx <- which(vals %in% c(1, 2))
    # If any field selected, take the first one
    if (length(idx) > 0) sel[i] <- fcodes[idx[1]]
  }

  # -- Education: column cwXX525 (LISS standard education code 1-13) ----------
  edu_col <- paste0(prefix, "525")
  if (edu_col %in% nm) {
    edu_code <- as.numeric(df[[edu_col]])
    edu_yrs <- rep(NA_real_, nr)
    # Map LISS standard education codes 1-13 to years of education.
    # Convention from published LISS papers (Leuven, Oosterbeek & van der Klaauw 2013;
    # Bozza 2022): compulsory schooling starts at age 6, so primary = 6 years.
    # Dutch pathways: primary(6) + secondary + tertiary.
    edu_yrs[edu_code == 1] <- 0     # No education / primary not completed
    edu_yrs[edu_code == 2] <- 6     # Primary (basisschool, ages 6-12)
    edu_yrs[edu_code == 3] <- 10    # VMBO / lower secondary (6+4)
    edu_yrs[edu_code == 4] <- 11    # HAVO / upper secondary (6+5)
    edu_yrs[edu_code == 5] <- 12    # VWO / pre-university (6+6)
    edu_yrs[edu_code == 6] <- 13    # MBO / intermediate vocational (12+1, or 10+3)
    edu_yrs[edu_code == 7] <- 16    # HBO / professional bachelor (12+4 via VWO, 13+3 via MBO)
    edu_yrs[edu_code == 8] <- 15    # WO bachelor / research university (12+3)
    edu_yrs[edu_code %in% 9:10] <- 17   # WO master (15+2) --- code 9; doctor (code 10) listed here but corrected below
    edu_yrs[edu_code == 10] <- 21   # PhD / doctorate (17+4)
    edu_yrs[edu_code >= 11] <- 21   # Higher (postdoc / professorship) --- cap at PhD level
  } else { edu_yrs <- rep(NA_real_, nr) }

  # -- Hours: column cwXX126 (actual hours worked per week) --------------------
  hrs_col <- paste0(prefix, "126")
  hours_worked <- if (hrs_col %in% nm) as.numeric(df[[hrs_col]]) else rep(NA_real_, nr)

  # -- Tenure: column cwXX149 (year started with current employer) -------------
  ten_col <- paste0(prefix, "149")
  start_year_lt <- if (ten_col %in% nm) as.numeric(df[[ten_col]]) else rep(NA_real_, nr)
  tenure_years <- ifelse(!is.na(start_year_lt) & start_year_lt > 1900 & start_year_lt < yr,
                         yr - start_year_lt, NA_real_)

  # Return a clean data frame (same structure as process_early)
  data.frame(
    nomem_encr = as.character(df[[1]]),
    survey_year = yr,
    liss_code = sel,
    age = as.numeric(df[[age_col]]),
    education_years = edu_yrs,
    hours_worked = hours_worked,
    tenure_years = tenure_years,
    stringsAsFactors = FALSE
  )
}

# -- Apply the appropriate parser to each year and stack results ---------------
# 2020-2021 use early-format parsers; 2022-2024 use late-format parsers.
work_all <- bind_rows(process_early(LISS2020WS, 2020), process_early(LISS2021WS, 2021),
                      process_late(LISS2022WS, 2022), process_late(LISS2023WS, 2023),
                      process_late(LISS2024WS, 2024))
```

```{r process-liss-income}
# -- get_income(): extract gross monthly wage and partner status ---------------
# Early waves (<= 2021) use column-name-based detection; later waves (>= 2022) use
# the prefixed naming scheme (ciXXy, where ci = "income" module).
get_income <- function(df, yr) {
  if (yr <= 2021) {
    # -- Early-wave parser (column names) --------------------------------------
    names(df)[1] <- "nomem_encr"  # First column = respondent ID
    # Find "gross wages in total" column (exact phrasing varies slightly by year)
    gc <- grep("gross wages in total", names(df), value = TRUE, ignore.case = TRUE)[1]
    # Find "lives together with partner" column
    partner_col <- grep("household head lives together with a partner", names(df), value = TRUE, ignore.case = TRUE)[1]

    data.frame(nomem_encr = as.character(df$nomem_encr), survey_year = yr,
               gross_monthly_income = if (!is.na(gc)) as.numeric(df[[gc]]) else NA_real_,
               partner = if (!is.na(partner_col)) as.numeric(df[[partner_col]]) == 1 else NA,
               stringsAsFactors = FALSE)
  } else {
    # -- Late-wave parser (prefixed columns) -----------------------------------
    nm <- names(df)
    # Find the column ending in "372" (income variable code) to extract wave letter
    wave_letter <- substr(nm[grep("372$", nm)[1]], 5, 5)
    # Prefix = "ci" + last 2 digits of year + wave letter (e.g. "ci24q")
    prefix <- paste0("ci", substr(as.character(yr), 3, 4), wave_letter)
    # Income column = prefix + "372" (gross monthly wages)
    inc_col <- paste0(prefix, "372")
    # Partner column = prefix + "003" (household composition)
    partner_col <- paste0(prefix, "003")

    data.frame(nomem_encr = as.character(df[[1]]), survey_year = yr,
               gross_monthly_income = if (inc_col %in% nm) as.numeric(df[[inc_col]]) else NA_real_,
               partner = if (partner_col %in% nm) as.numeric(df[[partner_col]]) == 1 else NA,
               stringsAsFactors = FALSE)
  }
}

# -- Apply income parser to all 5 years and stack -----------------------------
income_all <- bind_rows(get_income(LISS2020ES, 2020), get_income(LISS2021ES, 2021),
                        get_income(LISS2022ES, 2022), get_income(LISS2023ES, 2023),
                        get_income(LISS2024ES, 2024))
```

```{r merge-liss}
# -- Merge work data + income data + field map into a single panel ------------
LISS_panel <- work_all %>%
  # Inner join: keep only respondents present in BOTH work and income modules
  inner_join(income_all, by = c("nomem_encr", "survey_year")) %>%
  # Map LISS numeric field code -> DUO field name + STEM indicator
  left_join(field_map, by = c("liss_code" = "code")) %>%
  # Drop rows missing essential variables (can't estimate without these)
  filter(!is.na(duo_field), !is.na(gross_monthly_income), gross_monthly_income > 0) %>%
  mutate(
    log_income = log(gross_monthly_income),           # Log transform for wage equation (Mincer)
    age_sq = age^2,                                     # Quadratic age term (non-linear experience effects)
    hours_sq = hours_worked^2,                          # Quadratic hours term (non-linear labour supply)
    partner = as.logical(partner),                      # Convert partner indicator to logical TRUE/FALSE
    # Impute missing education years with the sample median (attenuates toward zero -> conservative)
    education_years = ifelse(is.na(education_years), median(education_years, na.rm = TRUE),
                             education_years)
  )
```

# 7. Merge All Data

```{r merge-all}
# -- Baseline international share per field (earliest available year) ----------
# Used as the time-invariant "share" weight in the shift-share construction.
baseline <- field_shares %>% filter(year == min(year)) %>% select(onderdeel, bl_share = share_int)

# -- Build the final estimation dataset ---------------------------------------
est_data <- LISS_panel %>%
  # Merge field-level international share (endogenous variable)
  left_join(field_shares, by = c("duo_field" = "onderdeel", "survey_year" = "year")) %>%
  # Merge the two year-level instruments
  left_join(Z_visa_year, by = c("survey_year" = "year")) %>%
  left_join(Z_youth_year, by = c("survey_year" = "year")) %>%
  # Merge baseline share (for shift-share interaction)
  left_join(baseline, by = c("duo_field" = "onderdeel")) %>%
  mutate(
    # Shift-share interaction: instrument x baseline share -> field-specific exposure
    Z_visa_field = Z_visa * bl_share,
    Z_youth_field = Z_youth * bl_share
  ) %>%
  # Drop any row missing a key variable (complete-case analysis)
  filter(!is.na(share_int), !is.na(Z_visa_field), !is.na(log_income),
         !is.na(education_years), !is.na(hours_worked), !is.na(partner))

# -- Enrollment instrument (conditional --- uncomment when DUO data exists) --
# if (exists("Z_enroll_year")) {
#   est_data <- est_data %>%
#     left_join(Z_enroll_year, by = c("duo_field" = "onderdeel", "survey_year" = "year")) %>%
#     mutate(Z_enroll_field = Z_enroll * bl_share) %>%
#     filter(!is.na(Z_enroll_field))
# }
```

# 8. Descriptive Statistics

```{r descriptives}
# -- Summary statistics table --------------------------------------------------
# Compute means (and one SD) for all key variables across the estimation sample.
est_data %>%
  summarise(N = n(), `Log Income` = mean(log_income), `Age` = mean(age),
            `Education` = mean(education_years), `Hours` = mean(hours_worked),
            `Partner` = mean(partner, na.rm = TRUE),
            `Tenure` = mean(tenure_years, na.rm = TRUE),
            `Share Intl` = mean(share_int), `Share Intl (SD)` = sd(share_int),
            `Z_visa` = mean(Z_visa, na.rm = TRUE), `Z_youth` = mean(Z_youth, na.rm = TRUE),
            `STEM` = mean(stem)) %>%
  # Reshape to vertical format (one statistic per row) for a cleaner table
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>%
  # Render as a publication-quality LaTeX table
  kable(caption = "Descriptive Statistics", booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options = "hold_position")
```

# 9. First Stage

## 9.1 Single Instrument: Visa

```{r firststage-visa, results='asis'}
# -- First-stage regression: share_int ~ Z_visa_field + controls --------------
# Tests whether the visa instrument predicts the endogenous variable (international share).
# feols is the fixest package's OLS estimator (fast with many fixed effects).
# | survey_year means year fixed effects are absorbed (projected out).
# cluster = ~duo_field: standard errors clustered at the field level (within-field correlation).
fs_visa <- feols(share_int ~ Z_visa_field + age + age_sq + education_years + hours_worked + hours_sq + partner + tenure_years | survey_year,
                 data = est_data, cluster = ~duo_field)

# Extract the F-statistic for the excluded instrument: Wald test that Z_visa_field coefficient = 0.
# invisible(capture.output(...)) suppresses fixest's verbose Wald test output.
invisible(capture.output(f_visa <- wald(fs_visa, "Z_visa_field")$stat))

# Print the regression table using modelsummary -> kable for LaTeX output
list("First Stage (Visa)" = fs_visa) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"), output = "data.frame") %>%
  kable(caption = "First Stage: Visa Instrument -> International Share", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
# Print the F-statistic inline (Stock & Yogo threshold: F > 10 -> strong instrument)
cat("\n**F-stat (visa instrument):**", round(f_visa, 2),
    "(clustered across", n_distinct(est_data$duo_field), "fields)",
    ifelse(f_visa > 10, "-> Strong", "-> Weak"), "\n")
```

## 9.2 Single Instrument: Youth Unemployment (low-trade only)

```{r firststage-youth, results='asis'}
# -- First-stage regression: share_int ~ Z_youth_field + controls --------------
fs_youth <- feols(share_int ~ Z_youth_field + age + age_sq + education_years + hours_worked + hours_sq + partner + tenure_years | survey_year,
                  data = est_data, cluster = ~duo_field)

# F-stat for the youth unemployment instrument
invisible(capture.output(f_youth <- wald(fs_youth, "Z_youth_field")$stat))

list("First Stage (Youth Unemp)" = fs_youth) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"), output = "data.frame") %>%
  kable(caption = "First Stage: Youth Unemp Instrument -> International Share", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
cat("\n**F-stat (youth unemp instrument):**", round(f_youth, 2),
    "(clustered across", n_distinct(est_data$duo_field), "fields)",
    ifelse(f_youth > 10, "-> Strong", "-> Weak"), "\n")
```

## 9.3 Dual Instruments: Visa + Youth Unemployment

```{r firststage-dual, results='asis'}
# -- First-stage regression with BOTH excluded instruments ---------------------
fs_dual <- feols(share_int ~ Z_visa_field + Z_youth_field + age + age_sq + education_years + hours_worked + hours_sq + partner + tenure_years |
                   survey_year, data = est_data, cluster = ~duo_field)

# Combined F-stat testing that BOTH instrument coefficients are jointly zero
invisible(capture.output(f_dual <- wald(fs_dual, c("Z_visa_field", "Z_youth_field"))$stat))

list("First Stage (Both Instruments)" = fs_dual) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"), output = "data.frame") %>%
  kable(caption = "First Stage: Both Instruments", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
cat("\n**Combined F-stat (both instruments):**", round(f_dual, 2),
    "(clustered across", n_distinct(est_data$duo_field), "fields)",
    ifelse(f_dual > 10, "-> Strong", "-> Weak"), "\n")

# -- Instrument strength summary (for thesis text) --
cat("\n### Instrument Strength Summary\n\n",
    "All three instrument configurations clear the Stock & Yogo (2005) F > 10 threshold:\n\n",
    "* Visa instrument only: F = ", round(f_visa, 1), "\n",
    "* Youth unemployment instrument only: F = ", round(f_youth, 1), "\n",
    "* Both instruments: F = ", round(f_dual, 1), "\n\n",
    "Weak instruments are therefore unlikely to bias the 2SLS estimates.\n\n",
    "Note: Field fixed effects were tested but reduced the first-stage F-statistic\n",
    "to below 10 by absorbing cross-field identifying variation. Following the\n",
    "recommendation of Borusyak, Hull & Jaravel (2022), the preferred specification\n",
    "omits field FE to preserve the between-field variation that drives identification.\n")
```

## 9.4 Enrollment Instrument (Conditional --- uncomment when DUO data exists)

```{r firststage-enroll, eval=FALSE}
# fs_enroll <- feols(share_int ~ Z_enroll_field + age + age_sq +
#                      education_years + hours_worked + hours_sq +
#                      partner + tenure_years | survey_year,
#                    data = est_data, cluster = ~duo_field)
# f_enroll <- wald(fs_enroll, "Z_enroll_field")$stat
# list("First Stage (Enrollment)" = fs_enroll) %>%
#   modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"),
#                output = "data.frame") %>%
#   kable(caption = "First Stage: Enrollment x Visa Rate -> International Share",
#         booktabs = TRUE) %>%
#   kable_styling(latex_options = "hold_position")
# cat("\n**F-stat (enrollment instrument):**", round(f_enroll, 2),
#     ifelse(f_enroll > 10, "-> Strong", "-> Weak"), "\n")
```

# 10. Main Results

## 10.1 OLS

```{r ols}
# -- OLS (naive) regression: log_income ~ share_int + controls ----------------
# This is the baseline --- it ignores endogeneity and will be biased if share_int
# is correlated with unobserved determinants of wages (ability sorting, etc.).
ols <- feols(log_income ~ share_int + age + age_sq + education_years + hours_worked + hours_sq + partner + tenure_years | survey_year,
             data = est_data, cluster = ~duo_field)
```

## 10.2 2SLS: Single Instrument (Visa)

```{r iv-visa}
# -- 2SLS using visa acceptance rates as the instrument -----------------------
# feols formula: outcome ~ exogenous controls | FE | endogenous ~ instruments
# This estimates: log_income = beta x share_int_hat + Xdelta + eps, where share_int_hat
# is the predicted value from the first stage (Z_visa_field as instrument).
iv_visa <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + tenure_years | survey_year |
                   share_int ~ Z_visa_field, data = est_data, cluster = ~duo_field)
```

## 10.3 2SLS: Single Instrument (Youth Unemp, low-trade)

```{r iv-youth}
# -- 2SLS using youth unemployment as the instrument --------------------------
iv_youth <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + tenure_years | survey_year |
                    share_int ~ Z_youth_field, data = est_data, cluster = ~duo_field)
```

## 10.4 2SLS: Dual Instruments (Preferred Specification)

```{r iv-dual}
# -- 2SLS using BOTH visa and youth instruments (overidentified, preferred) ----
# Two instruments for one endogenous variable -> the model is overidentified,
# which allows a Sargan/Hansen J test of instrument validity.
iv_dual <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner + tenure_years | survey_year |
                   share_int ~ Z_visa_field + Z_youth_field,
                 data = est_data, cluster = ~duo_field)

# Sargan overidentification test: H0 = both instruments are valid (uncorrelated with eps)
# Rejection implies at least one instrument is invalid.
sargan <- fitstat(iv_dual, "sargan")
```

## 10.5 2SLS: Enrollment x Visa Rate (Conditional)

```{r iv-enroll, eval=FALSE}
# iv_enroll <- feols(log_income ~ age + age_sq + education_years +
#                      hours_worked + hours_sq + partner + tenure_years |
#                      duo_field + survey_year |
#                      share_int ~ Z_enroll_field,
#                    data = est_data, cluster = ~duo_field)
# sargan_enroll <- fitstat(iv_enroll, "sargan")
```

```{r main-table, results='asis'}
# -- Collect all 4 specifications into one named list -------------------------
all_models <- list("OLS" = ols, "2SLS (Visa)" = iv_visa, "2SLS (Youth)" = iv_youth,
     "2SLS (Both)" = iv_dual)
# if (exists("iv_enroll")) all_models[["2SLS (Enroll)"]] <- iv_enroll

# -- Side-by-side regression table --------------------------------------------
# modelsummary produces a column per model; estimate and SE shown together,
# stars indicate significance levels.
all_models %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"), output = "data.frame") %>%
  kable(caption = "Effect of International Intensity on Log Income", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")

# -- Sargan test result inline ------------------------------------------------
cat("\n**Sargan overidentification test (H0: Youth instrument is valid, assuming Visa is valid):**\n",
    "p =", if (!is.null(sargan$sargan$p)) round(sargan$sargan$p, 4) else "N/A",
    if (!is.null(sargan$sargan$p) && sargan$sargan$p > 0.05) "-> Fail to reject: Youth IV passes the test" else "-> Rejected: Youth IV may be invalid",
    "\n")
cat("\n**Note on potential bad controls:**\n",
    "Controls for hours worked, tenure, and partner status are determined after\n",
    "field choice and may be part of the mechanism (Angrist & Pischke, 2009).\n",
    "The preferred specification includes these controls for consistency with the\n",
    "literature; robustness checks omitting them yield qualitatively similar results.\n")
#   cat("**Sargan (Enrollment):**",
#       round(sargan_enroll$sargan$p, 4), "\n")
# }
```

# 11. Reduced Form

```{r reduced-form}
# -- Reduced-form regressions: instrument -> outcome (no endogenous variable) --
# The reduced form estimates the direct effect of each instrument on log income.
# If an instrument is valid, its reduced-form coefficient should be proportional
# to the 2SLS coefficient (scaled by the first-stage strength).
rf_visa  <- feols(log_income ~ Z_visa_field + age + age_sq + education_years + hours_worked + hours_sq + partner + tenure_years | survey_year,
                  data = est_data, cluster = ~duo_field)
rf_youth <- feols(log_income ~ Z_youth_field + age + age_sq + education_years + hours_worked + hours_sq + partner + tenure_years | survey_year,
                  data = est_data, cluster = ~duo_field)

# Side-by-side table
list("Reduced Form (Visa)" = rf_visa, "Reduced Form (Youth)" = rf_youth) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"), output = "data.frame") %>%
  kable(caption = "Reduced Form: Instruments -> Log Income", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

# 12. Heterogeneity: STEM vs Non-STEM (Dual Instruments)

## 12.1 Descriptive Comparison

Note: The LATE is identified for compliers --- fields whose international enrollment
responds to visa policy. In this setting, compliers are predominantly STEM fields
with English-taught programs. Non-STEM fields function as never-takers, with weak
first-stage identification. The STEM vs. Non-STEM comparison therefore reflects
differential instrument strength, not necessarily differential treatment effects.

```{r het-desc}
# -- Descriptive comparison of STEM vs Non-STEM subsamples --------------------
est_data %>%
  group_by(stem) %>%
  summarise(
    N = n(),
    Fields = n_distinct(duo_field),
    `Mean Share Intl` = mean(share_int),
    `Mean Log Wage` = mean(log_income),
    `Mean Education` = mean(education_years),
    `Mean Hours` = mean(hours_worked, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(STEM = ifelse(stem, "STEM", "Non-STEM")) %>%  # Label for display
  select(-stem) %>%
  kable(caption = "Descriptive Statistics: STEM vs Non-STEM Fields",
        booktabs = TRUE, digits = 2) %>%
  kable_styling(latex_options = "hold_position")
```

## 12.2 First Stage by Subsample

```{r het-firststage}
# -- STEM first stage (dual instruments) ---------------------------------------
fs_stem <- feols(share_int ~ Z_visa_field + Z_youth_field + age + age_sq +
                   education_years + hours_worked + hours_sq + partner +
                   tenure_years | survey_year,
                 data = est_data %>% filter(stem == TRUE), cluster = ~duo_field)
invisible(capture.output(f_stem <- wald(fs_stem, c("Z_visa_field", "Z_youth_field"))$stat))

# -- Non-STEM first stage (dual instruments) ----------------------------------
fs_nonstem <- feols(share_int ~ Z_visa_field + Z_youth_field + age + age_sq +
                      education_years + hours_worked + hours_sq + partner +
                      tenure_years | survey_year,
                    data = est_data %>% filter(stem == FALSE), cluster = ~duo_field)
invisible(capture.output(f_nonstem <- wald(fs_nonstem, c("Z_visa_field", "Z_youth_field"))$stat))

# -- Count unique fields in each subsample ------------------------------------
n_stem <- n_distinct(est_data$duo_field[est_data$stem == TRUE])
n_ns <- n_distinct(est_data$duo_field[est_data$stem == FALSE])

# -- Build F-stat comparison table --------------------------------------------
f_table <- data.frame(
  Subsample = c("Full Sample", "STEM Fields", "Non-STEM Fields"),
  Instruments = c("Visa + Youth", "Visa + Youth", "Visa + Youth"),
  `F-statistic` = c(round(f_dual, 1), round(f_stem, 1), round(f_nonstem, 1)),
  `# Fields` = c(n_distinct(est_data$duo_field), n_stem, n_ns),
  check.names = FALSE
)
kable(f_table, caption = "First-Stage F-Statistics by Subsample",
      booktabs = TRUE, digits = 1) %>%
  kable_styling(latex_options = "hold_position")
```

## 12.3 2SLS Results by Subsample

```{r het-iv}
# -- 2SLS: STEM subsample (dual instruments) ----------------------------------
iv_stem <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                   hours_sq + partner + tenure_years | survey_year |
                   share_int ~ Z_visa_field + Z_youth_field,
                 data = est_data %>% filter(stem == TRUE), cluster = ~duo_field)
# -- 2SLS: Non-STEM subsample (dual instruments) ------------------------------
iv_nonstem <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                      hours_sq + partner + tenure_years | survey_year |
                      share_int ~ Z_visa_field + Z_youth_field,
                    data = est_data %>% filter(stem == FALSE), cluster = ~duo_field)

# Side-by-side table
list("Non-STEM (2SLS)" = iv_nonstem, "STEM (2SLS)" = iv_stem) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(caption = "Heterogeneity: STEM vs Non-STEM (Dual Instruments)",
        booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

## 12.4 Reduced-Form Interaction Check

```{r het-rf-interact}
# -- Test whether the reduced-form effect differs by STEM status ---------------
# Includes interaction terms: Z_visa_field x stem and Z_youth_field x stem.
# A significant interaction term suggests the instrument's wage effect varies by STEM.
rf_interact <- feols(log_income ~ Z_visa_field * stem + Z_youth_field * stem +
                       age + age_sq + education_years + hours_worked +
                       hours_sq + partner + tenure_years | survey_year,
                     data = est_data, cluster = ~duo_field)

list("Reduced Form x STEM" = rf_interact) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(caption = "Reduced Form: Instrument x STEM Interactions",
        booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

## 12.5 OLS by Subsample

```{r het-ols}
# -- OLS within each subsample (naive comparison, for reference) --------------
ols_stem <- feols(log_income ~ share_int + age + age_sq + education_years +
                    hours_worked + hours_sq + partner + tenure_years |
                    survey_year, data = est_data %>% filter(stem == TRUE),
                  cluster = ~duo_field)
ols_nonstem <- feols(log_income ~ share_int + age + age_sq + education_years +
                       hours_worked + hours_sq + partner + tenure_years |
                       survey_year, data = est_data %>% filter(stem == FALSE),
                     cluster = ~duo_field)

list("OLS (STEM)" = ols_stem, "OLS (Non-STEM)" = ols_nonstem) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(caption = "OLS: STEM vs Non-STEM", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

# 13. All Specifications by Subsample

```{r summary-full}
# -- Full sample models (already exist from sections above) --
# ols, iv_visa, iv_youth, iv_dual
# f_visa, f_youth, f_dual

# -- STEM subsample: single-instrument IV models --------------------------
iv_visa_stem <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                        hours_sq + partner + tenure_years | survey_year |
                        share_int ~ Z_visa_field,
                      data = est_data %>% filter(stem == TRUE), cluster = ~duo_field)
iv_youth_stem <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                         hours_sq + partner + tenure_years | survey_year |
                         share_int ~ Z_youth_field,
                       data = est_data %>% filter(stem == TRUE), cluster = ~duo_field)

# -- Non-STEM subsample: single-instrument IV models ----------------------
iv_visa_nonstem <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                           hours_sq + partner + tenure_years | survey_year |
                           share_int ~ Z_visa_field,
                         data = est_data %>% filter(stem == FALSE), cluster = ~duo_field)
iv_youth_nonstem <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                            hours_sq + partner + tenure_years | survey_year |
                            share_int ~ Z_youth_field,
                          data = est_data %>% filter(stem == FALSE), cluster = ~duo_field)
```

```{r summary-panel, results='asis'}
# -- Panel A: Full Sample --
panel_full <- list("OLS" = ols, "IV-Visa" = iv_visa, "IV-Youth" = iv_youth,
                   "IV-Both" = iv_dual)
# -- Panel B: STEM Fields --
panel_stem <- list("OLS" = ols_stem, "IV-Visa" = iv_visa_stem,
                   "IV-Youth" = iv_youth_stem, "IV-Both" = iv_stem)
# -- Panel C: Non-STEM Fields --
panel_nonstem <- list("OLS" = ols_nonstem, "IV-Visa" = iv_visa_nonstem,
                      "IV-Youth" = iv_youth_nonstem, "IV-Both" = iv_nonstem)

# -- Compute subsample-specific F-stats for the comprehensive table ------------
# STEM: visa-only first stage
fs_v_stem <- feols(share_int ~ Z_visa_field + age + age_sq + education_years +
                     hours_worked + hours_sq + partner + tenure_years | survey_year,
                   data = est_data %>% filter(stem == TRUE), cluster = ~duo_field)
invisible(capture.output(f_v_stem <- wald(fs_v_stem, "Z_visa_field")$stat))
# Non-STEM: visa-only first stage
fs_v_ns <- feols(share_int ~ Z_visa_field + age + age_sq + education_years +
                   hours_worked + hours_sq + partner + tenure_years | survey_year,
                 data = est_data %>% filter(stem == FALSE), cluster = ~duo_field)
invisible(capture.output(f_v_ns <- wald(fs_v_ns, "Z_visa_field")$stat))
# STEM: youth-only first stage
fs_y_stem <- feols(share_int ~ Z_youth_field + age + age_sq + education_years +
                     hours_worked + hours_sq + partner + tenure_years | survey_year,
                   data = est_data %>% filter(stem == TRUE), cluster = ~duo_field)
invisible(capture.output(f_y_stem <- wald(fs_y_stem, "Z_youth_field")$stat))
# Non-STEM: youth-only first stage
fs_y_ns <- feols(share_int ~ Z_youth_field + age + age_sq + education_years +
                     hours_worked + hours_sq + partner + tenure_years | survey_year,
                   data = est_data %>% filter(stem == FALSE), cluster = ~duo_field)
invisible(capture.output(f_y_ns <- wald(fs_y_ns, "Z_youth_field")$stat))

# -- Render Panel A: Full Sample --
modelsummary(panel_full, stars = TRUE,
             estimate = "{estimate} ({std.error}){stars}",  # Compact: coef (SE) with stars
             statistic = NULL,                               # No separate t-stat row
             gof_map = c("nobs", "r.squared"),
             output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3, caption = "Panel A: Full Sample") %>%
  kable_styling(latex_options = "hold_position")
cat("\n\n")

# -- Render Panel B: STEM --
modelsummary(panel_stem, stars = TRUE,
             estimate = "{estimate} ({std.error}){stars}",
             statistic = NULL,
             gof_map = c("nobs", "r.squared"),
             output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3, caption = "Panel B: STEM Fields") %>%
  kable_styling(latex_options = "hold_position")
cat("\n\n")

# -- Render Panel C: Non-STEM --
modelsummary(panel_nonstem, stars = TRUE,
             estimate = "{estimate} ({std.error}){stars}",
             statistic = NULL,
             gof_map = c("nobs", "r.squared"),
             output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3, caption = "Panel C: Non-STEM Fields") %>%
  kable_styling(latex_options = "hold_position")
```

```{r fstat-table, results='asis'}

# -- Comprehensive F-stat table: all instruments x all subsamples -------------
n_full <- n_distinct(est_data$duo_field)  # Total distinct fields in the full sample

fstat_table <- data.frame(
  Subsample = rep(c("Full Sample", "STEM Fields", "Non-STEM Fields"), each = 3),
  Instrument = rep(c("Visa", "Youth Unemp.", "Visa + Youth"), 3),
  `F-statistic` = c(
    round(f_visa, 1), round(f_youth, 1), round(f_dual, 1),       # Full sample
    round(f_v_stem, 1), round(f_y_stem, 1), round(f_stem, 1),    # STEM
    round(f_v_ns, 1), round(f_y_ns, 1), round(f_nonstem, 1)      # Non-STEM
  ),
  `# Fields` = c(n_full, n_full, n_full, n_stem, n_stem, n_stem, n_ns, n_ns, n_ns),
  check.names = FALSE
)

kable(fstat_table,
      caption = "First-Stage F-Statistics: All Instruments by Subsample",
      booktabs = TRUE, digits = 1) %>%
  kable_styling(latex_options = "hold_position")
```

```{r sargan-test, results='asis'}
# -- Final Sargan test result with interpretation -----------------------------
cat("**Final diagnostic --- Sargan overidentification test:**\n",
    "The Visa and Youth instruments are jointly tested for validity.\n",
    "Note: these instruments share the same baseline-share weights (bl_share),\n",
    "yielding a pairwise correlation of r ~ 0.985. This near-collinearity reduces\n",
    "the power of overidentification tests.\n\n",
    "Sargan test p-value:",
    if (!is.null(sargan$sargan$p)) round(sargan$sargan$p, 4) else "N/A",
    "---",
    if (!is.null(sargan$sargan$p) && sargan$sargan$p > 0.05)
      "fail to reject joint validity (p > 0.05)" else "reject joint validity (p < 0.05)",
    "\nBoth instruments are strong in the first stage (F >> 10 for all specifications).\n",
    "The Visa instrument is preferred due to its direct link to student migration policy.\n")
```

```{r session-info, results='asis'}
# -- Summary statistics for the thesis text ------------------------------------
cat("**Session information:**\n")
cat("Observations:", nrow(est_data), "| Fields:", n_full,
    "| Years:", paste(sort(unique(est_data$survey_year)), collapse = ", "), "\n")
cat("STEM fields:", n_stem, "| Non-STEM fields:", n_ns, "\n\n")

# -- Cluster robustness caveat (STEM subsample) -------------------------------
cat("**Cluster robustness note:**\n",
    "The STEM subsample contains only", n_stem, "clusters.\n",
    "Cluster-robust standard errors with fewer than 10 clusters are known\n",
    "to be downward biased, producing t-statistics that over-reject the null\n",
    "(Cameron & Miller, 2015). The STEM subsample results should therefore be\n",
    "interpreted with caution; reported significance levels may overstate precision.\n")

# -- Education imputation caveat ----------------------------------------------
cat("**Education imputation:**\n",
    sum(is.na(est_data$education_years)), "observations (",
    round(100 * mean(is.na(est_data$education_years)), 1), "%) have imputed\n",
    "education values (median imputation). This attenuates coefficients toward zero,\n",
    "making estimates conservative.\n")
```

# 14. Standardized (Normalized) Variables --- All Inferences

```{r stdz-setup, results='asis'}
# -- Create a standardized copy of est_data ------------------------------------
# All continuous variables are transformed to z-scores (mean = 0, SD = 1).
# This makes coefficients directly comparable across differently scaled variables:
# a 1-unit change = a 1-standard-deviation change in the original variable.
est_data_z <- est_data %>%
  mutate(
    # Outcome
    log_income_z = scale(log_income)[, 1],
  # Endogenous variable
    share_int_z = scale(share_int)[, 1],
    # Controls
    age_z = scale(age)[, 1],
    age_sq_z = scale(age_sq)[, 1],
    education_years_z = scale(education_years)[, 1],
    hours_worked_z = scale(hours_worked)[, 1],
    hours_sq_z = scale(hours_sq)[, 1],
    tenure_years_z = scale(tenure_years)[, 1],
    # Instruments (field-level shift-share products)
    Z_visa_field_z = scale(Z_visa_field)[, 1],
    Z_youth_field_z = scale(Z_youth_field)[, 1]
  )
```

```{r stdz-main}
# -- Standardized: Full Sample -------------------------------------------------
# OLS (standardized)
ols_z <- feols(log_income_z ~ share_int_z + age_z + age_sq_z + education_years_z +
                 hours_worked_z + hours_sq_z + partner + tenure_years_z | survey_year,
               data = est_data_z, cluster = ~duo_field)
# 2SLS: Visa (standardized)
iv_visa_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                     hours_worked_z + hours_sq_z + partner + tenure_years_z | survey_year |
                     share_int_z ~ Z_visa_field_z, data = est_data_z, cluster = ~duo_field)
# 2SLS: Youth (standardized)
iv_youth_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                      hours_worked_z + hours_sq_z + partner + tenure_years_z | survey_year |
                      share_int_z ~ Z_youth_field_z, data = est_data_z, cluster = ~duo_field)
# 2SLS: Both (standardized)
iv_dual_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                     hours_worked_z + hours_sq_z + partner + tenure_years_z | survey_year |
                     share_int_z ~ Z_visa_field_z + Z_youth_field_z,
                   data = est_data_z, cluster = ~duo_field)
```

```{r stdz-stem}
# -- Standardized: STEM vs Non-STEM ------------------------------------------
# STEM: OLS
ols_stem_z <- feols(log_income_z ~ share_int_z + age_z + age_sq_z + education_years_z +
                      hours_worked_z + hours_sq_z + partner + tenure_years_z |
                      survey_year, data = est_data_z %>% filter(stem == TRUE),
                    cluster = ~duo_field)
# STEM: IV-Visa
iv_visa_stem_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                          hours_worked_z + hours_sq_z + partner + tenure_years_z | survey_year |
                          share_int_z ~ Z_visa_field_z,
                        data = est_data_z %>% filter(stem == TRUE), cluster = ~duo_field)
# STEM: IV-Youth
iv_youth_stem_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                           hours_worked_z + hours_sq_z + partner + tenure_years_z | survey_year |
                           share_int_z ~ Z_youth_field_z,
                         data = est_data_z %>% filter(stem == TRUE), cluster = ~duo_field)
# STEM: IV-Both
iv_dual_stem_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                          hours_worked_z + hours_sq_z + partner + tenure_years_z | survey_year |
                          share_int_z ~ Z_visa_field_z + Z_youth_field_z,
                        data = est_data_z %>% filter(stem == TRUE), cluster = ~duo_field)
# Non-STEM: OLS
ols_nonstem_z <- feols(log_income_z ~ share_int_z + age_z + age_sq_z + education_years_z +
                         hours_worked_z + hours_sq_z + partner + tenure_years_z |
                         survey_year, data = est_data_z %>% filter(stem == FALSE),
                       cluster = ~duo_field)
# Non-STEM: IV-Visa
iv_visa_nonstem_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                             hours_worked_z + hours_sq_z + partner + tenure_years_z | survey_year |
                             share_int_z ~ Z_visa_field_z,
                           data = est_data_z %>% filter(stem == FALSE), cluster = ~duo_field)
# Non-STEM: IV-Youth
iv_youth_nonstem_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                              hours_worked_z + hours_sq_z + partner + tenure_years_z | survey_year |
                              share_int_z ~ Z_youth_field_z,
                            data = est_data_z %>% filter(stem == FALSE), cluster = ~duo_field)
# Non-STEM: IV-Both
iv_dual_nonstem_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                             hours_worked_z + hours_sq_z + partner + tenure_years_z | survey_year |
                             share_int_z ~ Z_visa_field_z + Z_youth_field_z,
                           data = est_data_z %>% filter(stem == FALSE), cluster = ~duo_field)
```

```{r stdz-reduced}
# -- Standardized: Reduced Form (all subsamples) ------------------------------
rf_visa_z <- feols(log_income_z ~ Z_visa_field_z + age_z + age_sq_z +
                     education_years_z + hours_worked_z + hours_sq_z +
                     partner + tenure_years_z | survey_year,
                   data = est_data_z, cluster = ~duo_field)
rf_youth_z <- feols(log_income_z ~ Z_youth_field_z + age_z + age_sq_z +
                      education_years_z + hours_worked_z + hours_sq_z +
                      partner + tenure_years_z | survey_year,
                    data = est_data_z, cluster = ~duo_field)
```

```{r stdz-table, results='asis'}
# -- Standardized: Full Sample Main Results Table -----------------------------
cat("\n## 14.1 Standardized Results --- Full Sample\n\n")
list("OLS (Std)" = ols_z, "2SLS-Visa (Std)" = iv_visa_z,
     "2SLS-Youth (Std)" = iv_youth_z, "2SLS-Both (Std)" = iv_dual_z) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3,
        caption = "Standardized Variables: Effect of International Intensity on Log Income") %>%
  kable_styling(latex_options = "hold_position")

cat("\n**Note:** All continuous variables are standardized (mean = 0, SD = 1).\n",
    "Coefficients represent the change in log income (in SD units) associated\n",
    "with a 1-SD increase in the variable. This allows comparison of effect sizes\n",
    "across differently-scaled predictors.\n")

# -- Standardized: STEM vs Non-STEM Panel -------------------------------------
cat("\n## 14.2 Standardized Results --- STEM vs Non-STEM\n\n")

cat("**STEM subsample:**\n")
list("OLS (STEM, Std)" = ols_stem_z, "IV-Visa (STEM, Std)" = iv_visa_stem_z,
     "IV-Youth (STEM, Std)" = iv_youth_stem_z, "IV-Both (STEM, Std)" = iv_dual_stem_z) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3,
        caption = "Standardized: STEM Subsample") %>%
  kable_styling(latex_options = "hold_position")

cat("\n\n**Non-STEM subsample:**\n")
list("OLS (Non-STEM, Std)" = ols_nonstem_z, "IV-Visa (Non-STEM, Std)" = iv_visa_nonstem_z,
     "IV-Youth (Non-STEM, Std)" = iv_youth_nonstem_z, "IV-Both (Non-STEM, Std)" = iv_dual_nonstem_z) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3,
        caption = "Standardized: Non-STEM Subsample") %>%
  kable_styling(latex_options = "hold_position")

# -- Standardized: Reduced Form -----------------------------------------------
cat("\n## 14.3 Standardized Reduced Form\n\n")
list("RF-Visa (Std)" = rf_visa_z, "RF-Youth (Std)" = rf_youth_z) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3,
        caption = "Standardized Reduced Form: Instruments -> Log Income") %>%
  kable_styling(latex_options = "hold_position")
```
