---
title: "International Student Presence and Dutch Graduate Labor Market Outcomes"
subtitle: "A Shift-Share Instrumental Variables Approach"
author: "Redon Ismailaga"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    number_sections: true
bibliography: references.bib
csl: apa.csl
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{float}
---

```{r setup, include=FALSE}
# -- Global knitr options ------------------------------------------------------
# echo = FALSE:   don't print the R code itself into the PDF --- only show outputs
# warning = FALSE: suppress R warnings (they'd clutter the PDF with irrelevant noise)
# message = FALSE: suppress package-loading messages (e.g. "Attaching package: dplyr")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r packages}
pacman::p_load(`pacman`)
p_load(tidyverse); p_load(readxl); p_load(janitor); p_load(fixest)
p_load(modelsummary); p_load(kableExtra); p_load(WDI)
p_load(rt); p_load(hunspell)
```

# 1. Data Loading

## 1.1 LISS Panel

```{r liss-load}
# -- LISS Work & Schooling module (2020-2024) ----------------------------------
# The LISS panel is a longitudinal Dutch household survey. The "Work and Schooling"
# module records each respondent's occupation, field of study, education level,
# and working hours. Each yearly file has its own column naming convention.
# Early waves (2020, 2021) use comma-separated format; later waves (2022-2024)
# use semicolon-separated format. check.names = FALSE prevents R from auto-converting
# column names (e.g. spaces to dots).

# 2020 wave: comma-separated (sep = "," is default), 1.0p = version 1.0 public
LISS2020WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw20m_EN_1.0p.csv", check.names = FALSE)
# 2021 wave: also comma-separated
LISS2021WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw21n_EN_1.0p.csv", check.names = FALSE)
# 2022 wave: semicolon-separated (sep = ";")
LISS2022WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw22o_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2023 wave: semicolon-separated
LISS2023WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw23p_EN_2.0p.csv", sep = ";", check.names = FALSE)
# 2024 wave: semicolon-separated
LISS2024WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw24q_EN_1.0p.csv", sep = ";", check.names = FALSE)

# -- LISS Economic Situation / Income module (2020-2024) -----------------------
# The "Economic Situation: Income" module records gross monthly wages, partner status,
# and other income-related variables. Same sep pattern as the Work & Schooling module.

# 2020 income data: comma-separated
LISS2020ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci20m_EN_1.0p.csv", check.names = FALSE)
# 2021 income data: comma-separated
LISS2021ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci21n_EN_1.0p.csv", check.names = FALSE)
# 2022 income data: semicolon-separated
LISS2022ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci22o_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2023 income data: semicolon-separated
LISS2023ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci23p_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2024 income data: semicolon-separated
LISS2024ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci24q_EN_1.0p.csv", sep = ";", check.names = FALSE)
```

## 1.2 IND Visa

```{r visa-load}
# -- IND Schengen visa statistics (2020-2024) ----------------------------------
# The Immigration and Naturalisation Service (IND) publishes yearly Schengen visa
# statistics by consulate/country. Each file is an Excel workbook (.xlsx) with
# a worksheet named "Data for consulates" containing acceptance/rejection rates
# by applicant nationality.
visa_files <- c(
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/Visa-statistics-for-consulates-2020_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2021-consulates-schengen-visa-stats.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/Visa statistics for consulates in 2022_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2023-schengen-visa-statistics-consulates_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2024-schengen-visa-statistics-consulates_en.xlsx"
)
```

## 1.4 CBS Nationality Data

```{r cbs-load}
# -- CBS temporary residence permits by nationality (2010) ---------------------
# CBS (Centraal Bureau voor de Statistiek) publishes counts of temporary residence
# permits (verblijfsvergunningen voor bepaalde tijd) by nationality and year.
# This data is used to construct the baseline nationality shares for the shift-share
# instrument. The file name includes a timestamp from when it was downloaded.
EthnicalMakeup <- read.csv("../../../Data/CBS Incoming Student Ethnical Makeup/Verblijfsvergunningen_voor_bepaalde_tijd_19052026_175309.csv", sep = ";")
```

# 2. Process IND Visa

```{r process-visa}
# -- IND Schengen visa acceptance rates (2020-2024) ---------------------------
# Load each year's Excel file, keep only Netherlands consulates, and extract
# acceptance rates by applicant nationality. Column names are selected after
# inspecting the raw datasets.

years <- 2020:2024
visa_list <- list()

for (i in seq_along(visa_files)) {
  if (!file.exists(visa_files[i])) next

  df <- read_excel(visa_files[i], sheet = "Data for consulates") %>%
    clean_names() %>%
    mutate(year = years[i])

  # Filter to Netherlands consulates and select key columns
  # (adjust the exact column names below after viewing the data in R)
  nl <- df %>%
    filter(schengen_state == "Netherlands") %>%
    select(
      country = country_where_consulate_is_located,
      not_issued_rate = not_issued_rate_for_uniform_visas,
      year
    ) %>%
    mutate(
      acceptance_rate = 1 - not_issued_rate,
      country = str_to_upper(str_trim(country)),
      # Turkey changed its English name to "Türkiye" in 2022.
      # The IND files use "TURKEY" (2020-2021) and "TÜRKIYE" (2022-2024).
      # Normalize to "TURKEY" so the inner_join with CBS country_std works.
      country = str_replace(country, "TÜRKIYE", "TURKEY")
    )

  if (nrow(nl) > 0) visa_list[[i]] <- nl
}

Consulates <- bind_rows(visa_list) %>%
  # Average across consulates within each country (e.g. US has DC + NY + Chicago)
  group_by(country, year) %>%
  summarise(acceptance_rate = mean(acceptance_rate, na.rm = TRUE), .groups = "drop")
```

# 3. Field-Level International Share (CBS)
```{r process-cbs-share}
# -- CBS international-background senior share (parents born outside NL) -----
# Uses the subgroups file with granular 3-digit CBS sub-field codes, senior year.
# All CBS sub-fields map to one of the 11 field keys (education through services).
# Field unknown set to NA (drops from merge — no bias from misassignment).

base <- "../../../Data/CBS Incoming Student Ethnical Makeup"

cbs_raw <- read_delim(file.path(base, "Born in or out NL, FoS, gender, subgroups, Senior year.csv"),
                      delim = ";", show_col_types = FALSE) %>%
  clean_names() %>%
  mutate(
    students = as.numeric(str_trim(ouderejaarsstudenten_aantal)),
    year = as.integer(str_sub(perioden, 1, 4))
  )

# Total seniors per field-year: sum across genders and onderwijssoort
total_seniors <- cbs_raw %>%
  filter(!is.na(students), students > 0) %>%
  group_by(studierichting, onderwijssoort, year) %>%
  summarise(total_seniors = sum(students, na.rm = TRUE), .groups = "drop")

# International-background seniors (parents born outside NL)
# Note: this includes Nederlandse Cariben students; they are proportionally
# removed below using the Internationals file's herkomstland breakdown.
intl_seniors <- cbs_raw %>%
  filter(geboorteland_ouders == "Geboren buiten Nederland",
         !is.na(students), students > 0) %>%
  group_by(studierichting, onderwijssoort, year) %>%
  summarise(intl_seniors = sum(students, na.rm = TRUE), .groups = "drop")

cbs_share <- total_seniors %>%
  left_join(intl_seniors, by = c("studierichting", "onderwijssoort", "year")) %>%
  mutate(
    share_int = ifelse(total_seniors > 0,
                       coalesce(intl_seniors, 0) / total_seniors, NA_real_),
    share_int = pmin(share_int, 1)
  )

# Map CBS 3-digit sub-field codes to CBS field groups.
# Skips aggregate codes (z.n.d. = zonder nadere detaillering) since
# the sub-fields already cover them.
cbs_field_map <- tibble(
  studierichting = c(
    "011 Onderwijskunde, lerarenopleiding...",
    "021 Kunst",
    "022 Geesteswetenschappen",
    "023 Talen",
    "031 Gedrag en maatschappij",
    "032 Journalistiek, informatievoorziening",
    "041 Bedrijfskunde en administratie",
    "042 Recht",
    "051 Biologie en biochemie",
    "052 Milieu en natuurbeheer",
    "053 Natuurwetenschappen",
    "054 Wiskunde en statistiek",
    "061 Informatica",
    "071 Techniek en technische dienstverlening",
    "072 Levensmiddelentechnologie",
    "073 Architectuur en bouwkunde",
    "081 Landbouw",
    "082 Bosbouw",
    "083 Visserij",
    "084 Diergeneeskunde en -verzorging",
    "091 Gezondheidszorg",
    "092 Welzijn",
    "101 Persoonlijke dienstverlening",
    "102 Hygiëne en arbeidsomstandigheden",
    "103 Openbare veiligheid",
    "104 Transport en logistiek",
    "Onderwijsrichting onbekend"
  ),
  field = c(
    "education",
    "arts_languages",
    "arts_languages",
    "arts_languages",
    "behavior_society",
    "behavior_society",
    "business_admin",
    "law",
    "natural_sciences",
    "natural_sciences",
    "natural_sciences",
    "natural_sciences",
    "engineering",
    "engineering",
    "engineering",
    "engineering",
    "agriculture",
    "agriculture",
    "agriculture",
    "agriculture",
    "healthcare",
    "healthcare",
    "services",
    "services",
    "services",
    "services",
    NA_character_
  )
)

# Map aggregate CBS fields to CBS field groups (used by the Internationals file)
cbs_field_map_instr <- tibble(
  studierichting = c("01 Onderwijs",
                     "02 Vormgeving, kunst, talen en gesch... ",
                     "03 Journalistiek, gedrag en maatschappij",
                     "041 Bedrijfskunde en administratie",
                     "042 Recht",
                     "05 Wiskunde, natuurwetenschappen",
                     "06 Informatica",
                     "07 Techniek, industrie en bouwkunde",
                     "08 Landbouw, diergeneeskunde en -verz...",
                     "09 Gezondheidszorg en welzijn",
                     "10 Dienstverlening",
                     "Onderwijsrichting onbekend"),
  field = c("education", "arts_languages", "behavior_society",
                "business_admin", "law",
                "natural_sciences", "engineering", "engineering",
                "agriculture", "healthcare",
                "services", NA_character_)
)

# Map CBS shares to CBS field groups (aggregates sub-fields to CBS field level)
field_shares <- cbs_share %>%
  inner_join(cbs_field_map, by = "studierichting") %>%
  group_by(field, onderwijssoort, year) %>%
  summarise(share_int = weighted.mean(share_int, total_seniors, na.rm = TRUE),
            .groups = "drop")

# Aggregate share_int for the joint (pooled) specification
field_shares_agg <- cbs_share %>%
  inner_join(cbs_field_map, by = "studierichting") %>%
  group_by(field, year) %>%
  summarise(share_int = weighted.mean(share_int, total_seniors, na.rm = TRUE),
            .groups = "drop")

# -- Proportional Cariben correction -------------------------------------------
# Caribbean students are Kingdom citizens who don't need Schengen visas.
# The instrument cannot shift their enrollment, so remove their proportional
# contribution from share_int using the Internationals file's breakdown.
# This preserves the subgroups file's foreign counts (which give the correct
# F-stat) while proportionally removing only the Cariben portion.

# Cariben foreign students per aggregate field from the Internationals file
cariben_raw <- read_delim(file.path(base, "Students by homeland, gender, FoS, Senior year, Subgroupped.csv"),
                          delim = ";", show_col_types = FALSE) %>%
  clean_names() %>%
  filter(herkomstland == "Nederlandse Cariben",
         geboorteland_ouders == "Geboren buiten Nederland") %>%
  mutate(
    count = as.numeric(str_trim(ouderejaarsstudenten_aantal)),
    year = as.integer(str_sub(perioden, 1, 4))
  ) %>%
  filter(!is.na(count), count > 0) %>%
  group_by(studierichting, onderwijssoort, year) %>%
  summarise(cariben = sum(count, na.rm = TRUE), .groups = "drop")

# Total foreign students per aggregate field from the Internationals file
intl_total_raw <- read_delim(file.path(base, "Students by homeland, gender, FoS, Senior year, Subgroupped.csv"),
                             delim = ";", show_col_types = FALSE) %>%
  clean_names() %>%
  filter(geboorteland_ouders == "Geboren buiten Nederland") %>%
  mutate(
    count = as.numeric(str_trim(ouderejaarsstudenten_aantal)),
    year = as.integer(str_sub(perioden, 1, 4))
  ) %>%
  filter(!is.na(count), count > 0) %>%
  group_by(studierichting, onderwijssoort, year) %>%
  summarise(intl_total = sum(count, na.rm = TRUE), .groups = "drop")

# Map both to CBS fields
cariben_field <- cariben_raw %>%
  inner_join(cbs_field_map_instr, by = "studierichting") %>%
  filter(!is.na(field)) %>%
  group_by(field, onderwijssoort, year) %>%
  summarise(cariben = sum(cariben, na.rm = TRUE), .groups = "drop")

intl_total_field <- intl_total_raw %>%
  inner_join(cbs_field_map_instr, by = "studierichting") %>%
  filter(!is.na(field)) %>%
  group_by(field, onderwijssoort, year) %>%
  summarise(intl_total = sum(intl_total, na.rm = TRUE), .groups = "drop")

# Cariben share of foreign students per CBS field
cariben_share <- cariben_field %>%
  left_join(intl_total_field, by = c("field", "onderwijssoort", "year")) %>%
  mutate(cariben_share = ifelse(intl_total > 0, cariben / intl_total, 0)) %>%
  select(field, onderwijssoort, year, cariben_share)

cariben_share_agg <- cariben_field %>%
  left_join(intl_total_field, by = c("field", "onderwijssoort", "year")) %>%
  group_by(field, year) %>%
  summarise(
    cariben = sum(cariben, na.rm = TRUE),
    intl_total = sum(intl_total, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(cariben_share = ifelse(intl_total > 0, cariben / intl_total, 0)) %>%
  select(field, year, cariben_share)

# Apply correction: share_int = share_int * (1 - cariben_share)
field_shares <- field_shares %>%
  left_join(cariben_share, by = c("field", "onderwijssoort", "year")) %>%
  mutate(
    cariben_share = coalesce(cariben_share, 0),
    share_int = share_int * (1 - cariben_share)
  ) %>%
  select(field, onderwijssoort, year, share_int)

field_shares_agg <- field_shares_agg %>%
  left_join(cariben_share_agg, by = c("field", "year")) %>%
  mutate(
    cariben_share = coalesce(cariben_share, 0),
    share_int = share_int * (1 - cariben_share)
  ) %>%
  select(field, year, share_int)
```

# 4. Process CBS -> Nationality Shares (2010, per AEA 10-year lag)

```{r process-cbs}
# -- Extract pre-period nationality shares -------------------------------------
# The shift-share instrument uses pre-period (2010) nationality shares as weights.
# We use year 2010 --- a full decade before the outcome period (2020-2024) --- to
# ensure the shares are exogenous to contemporaneous shocks (AEA convention).
# -- Clean and parse directly from EthnicalMakeup -----------------------------
cbs_shares <- EthnicalMakeup %>%
  mutate(
    nationality = str_to_title(str_trim(Nationaliteit)),
    permits = as.numeric(str_trim(Verblijfsvergunning.regulier.naar.motief.Studie..aantal.)),
    year = as.integer(str_extract(as.character(Perioden), "^[0-9]{4}"))
  ) %>%
  filter(!is.na(permits), permits > 0, !is.na(year))

# -- Aggregate to 2010 nationality shares -------------------------------------
# Keep only 2010, sum permits per nationality, compute share of total
nat_shares <- cbs_shares %>%
  filter(year == 2010) %>%                                   # Pre-period: 2010 only
  group_by(nationality) %>%                                  # Group by nationality name
  summarise(permits = sum(permits, na.rm = TRUE), .groups = "drop") %>%  # Total permits per nationality
  mutate(share = permits / sum(permits)) %>%                   # Each nationality's share of ALL permits
  arrange(desc(share))                                         # Sort by share (largest first)

# -- Map Dutch nationality adjectives to English country names -----------------
# CBS uses Dutch nationality adjectives (e.g. "Chinees", "Indiaas"), while IND
# visa data uses English country names. Complete named-vector lookup covering
# all 80 nationalities in the CBS permit data.

# Manual overrides for Dutch adjectives that CLDR regex can't resolve cleanly
nl_adj_to_en <- c(
  "Chinees"        = "CHINA",
  "Indiaas"        = "INDIA",
  "Turks"          = "TURKEY",
  "Turkse"         = "TURKEY",
  "Amerikaans"     = "UNITED STATES",
  "Brits"          = "UNITED KINGDOM",
  "Duits"          = "GERMANY",
  "Frans"          = "FRANCE",
  "Italiaans"      = "ITALY",
  "Spaans"         = "SPAIN",
  "Pools"          = "POLAND",
  "Roemeens"       = "ROMANIA",
  "Bulgaars"       = "BULGARIA",
  "Grieks"         = "GREECE",
  "Portugees"      = "PORTUGAL",
  "Hongaars"       = "HUNGARY",
  "Litouws"        = "LITHUANIA",
  "Russisch"       = "RUSSIA",
  "Braziliaans"    = "BRAZIL",
  "Japans"         = "JAPAN",
  "Mexicaans"      = "MEXICO",
  "Koreaans"       = "SOUTH KOREA",
  "Zuid-Koreaans"  = "SOUTH KOREA",
  "Indonesisch"    = "INDONESIA",
  "Iraans"         = "IRAN",
  "Taiwanees"      = "TAIWAN",
  "Maledivisch"    = "MALDIVES",
  "Surinaams"      = "SURINAME",
  "Oekraïens"      = "UKRAINE",
  "Bengalees"      = "BANGLADESH",
  "Filippijns"     = "PHILIPPINES",
  "Burger Servië"  = "SERBIA",
  "Ugandees"       = "UGANDA",
  "Argentijns"     = "ARGENTINA",
  "Armeens"        = "ARMENIA",
  "Mongools"       = "MONGOLIA",
  "Venezolaans"    = "VENEZUELA",
  "Wit-Russisch"   = "BELARUS",
  "Kosovaars"      = "KOSOVO",
  "Oezbeeks"       = "UZBEKISTAN",
  "Algerijns"      = "ALGERIA",
  "Jemenitisch"    = "YEMEN",
  "Canadees"       = "CANADA",
  "Australisch"    = "AUSTRALIA",
  "Singaporees"    = "SINGAPORE",
  "Nepalees"       = "NEPAL",
  "Marokkaans"     = "MOROCCO",
  "Nieuw-Zeelands" = "NEW ZEALAND",
  "Zuid-Afrikaans" = "SOUTH AFRICA",
  "Sierra Leoons"  = "SIERRA LEONE",
  "Salvadoraans"   = "EL SALVADOR",
  "Omanitisch"     = "OMAN",
  "Myanmarees"     = "MYANMAR",
  "Chileens"       = "CHILE",
  "Bhutaans"       = "BHUTAN",
  "Kazaks"         = "KAZAKHSTAN",
  "Sri Lankaans"   = "SRI LANKA",
  "Ecuadoraans"    = "ECUADOR",
  "Costa Ricaans"  = "COSTA RICA",
  "Macedonisch"    = "NORTH MACEDONIA",
  "Boliviaans"     = "BOLIVIA",
  "Moldavisch"     = "MOLDOVA",
  "Eritrees"       = "ERITREA",
  "Cubaans"        = "CUBA",
  "Jordaans"       = "JORDAN",
  "Liberiaans"     = "LIBERIA",
  "Syrisch"        = "SYRIA",
  "Tunesisch"      = "TUNISIA",
  "Albanees"       = "ALBANIA",
  "Kroatisch"      = "CROATIA",
  "Maleisisch"     = "MALAYSIA",
  "Bosnisch"       = "BOSNIA AND HERZEGOVINA",
  "Libanees"       = "LEBANON",
  "Namibisch"      = "NAMIBIA",
  "Soedanees"      = "SUDAN",
  "Azerbeidzjaans" = "AZERBAIJAN",
  "Georgisch"      = "GEORGIA",
  "Colombiaans"    = "COLOMBIA",
  "Ethiopisch"     = "ETHIOPIA",
  "Vietnamees"     = "VIETNAM",
  "Nigeriaans"     = "NIGERIA",
  "Ghanees"        = "GHANA",
  "Keniaans"       = "KENYA",
  "Peruaans"       = "PERU",
  "Tanzaniaans"    = "TANZANIA",
  "Rwandees"       = "RWANDA",
  "Zimbabwaans"    = "ZIMBABWE",
  "Thais"          = "THAILAND",
  "Egyptisch"      = "EGYPT",
  "Israëlisch"     = "ISRAEL",
  "Kameroens"      = "CAMEROON",
  "Pakistaans"     = "PAKISTAN",
  "Zambiaans"      = "ZAMBIA",
  "Afghaans"       = "AFGHANISTAN"
)

nat_shares <- nat_shares %>%
  mutate(country_std = coalesce(nl_adj_to_en[nationality],
                                str_to_upper(nationality)))
```

# 4b. Field × Nationality Matrix (for shift-share instrument)
#
# Use the CBS Internationals file (Herkomstland × Studierichting) to get
# field-specific nationality composition of international students.
# For the "Buiten-Europa" bucket, split into individual nationalities using
# the residence permit shares from section 4.

```{r field-nat-matrix}
base <- "../../../Data/CBS Incoming Student Ethnical Makeup"

# -- Read Internationals file --------------------------------------------------
intl_field <- read_delim(file.path(base, "Students by homeland, gender, FoS, Senior year, Subgroupped.csv"),
                         delim = ";", show_col_types = FALSE) %>%
  clean_names() %>%
  mutate(
    count = as.numeric(str_trim(ouderejaarsstudenten_aantal)),
    year = as.integer(str_sub(perioden, 1, 4))
  ) %>%
  # Keep onderwijssoort (HBO/WO) separate for level-specific instruments
  filter(!is.na(count), count > 0) %>%
  group_by(studierichting, herkomstland, onderwijssoort, year) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop")

# -- Aggregate across onderwijssoort for the joint instrument -----------------
intl_field_agg <- intl_field %>%
  group_by(studierichting, herkomstland, year) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop")

# -- Map CBS Herkomstland to English country_std used in visa data --------------
# Keep the line count from each origin group, then split Buiten-Europa.
# The 5 named groups are already separate from the Buiten-Europa bucket.
herkomst_map <- tibble(
  herkomstland = c("Nederland",
                   "Europa (exclusief Nederland)",
                   "Turkije", "Marokko", "Suriname",
                   "Nederlandse Cariben", "Indonesië",
                   "Buiten-Europa (excl. 5 grote herkomst..."),
  country_std = c(NA_character_,       # exclude Dutch
                  "EUROPE",            # keep as aggregate
                  "TURKEY", "MOROCCO", "SURINAME",
                  "NEDERLANDSE CARIBEN", "INDONESIA",
                  "BUITEN_EUROPA")     # to be split
)

intl_origin <- intl_field %>%
  inner_join(herkomst_map, by = "herkomstland") %>%
  filter(!is.na(country_std))  # drop Netherlands

intl_origin_agg <- intl_field_agg %>%
  inner_join(herkomst_map, by = "herkomstland") %>%
  filter(!is.na(country_std))

# Non-EU/EEA nationalities from the permit data that are NOT among the 5
# named groups (Turkey, Morocco, Suriname, Dutch Caribbean, Indonesia).
# CBS codes "Europa (exclusief Nederland)" as EU/EEA aggregate, so we
# need to exclude EU + EEA member states from the Buiten-Europa split.
# UK included too — was in the EU at baseline (2011) when shift-share
# weights are computed. UK Schengen visa rates also irrelevant (visa-free).
eu_eea_countries <- c("EUROPE",
  "AUSTRIA", "BELGIUM", "BULGARIA", "CROATIA", "CYPRUS",
  "CZECH REPUBLIC", "DENMARK", "ESTONIA", "FINLAND", "FRANCE",
  "GERMANY", "GREECE", "HUNGARY", "ICELAND", "IRELAND", "ITALY",
  "LATVIA", "LIECHTENSTEIN", "LITHUANIA", "LUXEMBOURG", "MALTA",
  "NETHERLANDS", "NORWAY", "POLAND", "PORTUGAL", "ROMANIA",
  "SLOVAKIA", "SLOVENIA", "SPAIN", "SWEDEN", "SWITZERLAND",
  "UNITED KINGDOM")

five_named_groups <- c("TURKEY", "MOROCCO", "SURINAME",
                       "NEDERLANDSE CARIBEN", "INDONESIA")

non_eu_countries <- nat_shares %>%
  filter(!is.na(country_std)) %>%
  filter(!country_std %in% c(eu_eea_countries, five_named_groups, "NETHERLANDS", "NEDERLAND")) %>%
  # Re-normalize shares to sum to 1 within this non-EU group
  mutate(share_non_eu = share / sum(share)) %>%
  select(country_std, share_non_eu)

# For each (studierichting, year), split the Buiten-Europa count
# across non-EU countries proportional to their permit share
bu_count <- intl_origin %>%
  filter(country_std == "BUITEN_EUROPA") %>%
  select(studierichting, onderwijssoort, year, bu_count = count)

bu_count_agg <- intl_origin_agg %>%
  filter(country_std == "BUITEN_EUROPA") %>%
  select(studierichting, year, bu_count = count)

bu_split <- bu_count %>%
  cross_join(non_eu_countries) %>%
  mutate(count = bu_count * share_non_eu) %>%
  select(studierichting, onderwijssoort, year, country_std, count)

bu_split_agg <- bu_count_agg %>%
  cross_join(non_eu_countries) %>%
  mutate(count = bu_count * share_non_eu) %>%
  select(studierichting, year, country_std, count)

# -- Combine all origin groups --------------------------------------------------
intl_detailed <- intl_origin %>%
  filter(country_std != "BUITEN_EUROPA") %>%
  bind_rows(bu_split)

intl_detailed_agg <- intl_origin_agg %>%
  filter(country_std != "BUITEN_EUROPA") %>%
  bind_rows(bu_split_agg)

# -- Map CBS Studierichting to CBS field groups ----------------------------------
cbs_field_map_instr <- tibble(
  studierichting = c("01 Onderwijs",
                     "02 Vormgeving, kunst, talen en gesch... ",
                     "03 Journalistiek, gedrag en maatschappij",
                     "041 Bedrijfskunde en administratie",
                     "042 Recht",
                     "05 Wiskunde, natuurwetenschappen",
                     "06 Informatica",
                     "07 Techniek, industrie en bouwkunde",
                     "08 Landbouw, diergeneeskunde en -verz...",
                     "09 Gezondheidszorg en welzijn",
                     "10 Dienstverlening",
                     "Onderwijsrichting onbekend"),
  field = c("education", "arts_languages", "behavior_society",
                "business_admin", "law",
                "natural_sciences", "engineering", "engineering",
                "agriculture", "healthcare",
                "services", NA_character_)
)

# -- Compute nationality shares per field-year ----------------------------------
# Baseline year = earliest year in data (2011)
baseline_year <- min(intl_detailed$year)

field_nat_shares <- intl_detailed %>%
  inner_join(cbs_field_map_instr, by = "studierichting") %>%
  group_by(field, onderwijssoort, year, country_std) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop") %>%
  group_by(field, onderwijssoort, year) %>%
  mutate(share_in_field = count / sum(count)) %>%
  ungroup()

field_nat_shares_agg <- intl_detailed_agg %>%
  inner_join(cbs_field_map_instr, by = "studierichting") %>%
  group_by(field, year, country_std) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop") %>%
  group_by(field, year) %>%
  mutate(share_in_field = count / sum(count)) %>%
  ungroup()

# Baseline nationality shares per field (time-invariant, from earliest year)
field_nat_bl <- field_nat_shares %>%
  filter(year == baseline_year) %>%
  select(field, onderwijssoort, country_std, share_jf = share_in_field)

field_nat_bl_agg <- field_nat_shares_agg %>%
  filter(year == baseline_year) %>%
  select(field, country_std, share_jf = share_in_field)

# -- Construct field-specific instrument: Z_ft = sum_j(share_jf × accept_rate_jt)
Z_visa_field_year <- field_nat_bl %>%
  inner_join(Consulates, by = c("country_std" = "country")) %>%
  filter(!is.na(acceptance_rate), !is.na(share_jf)) %>%
  group_by(field, onderwijssoort, year) %>%
  summarise(Z_visa_field = sum(share_jf * acceptance_rate, na.rm = TRUE),
            .groups = "drop")

# Joint instrument: aggregate across onderwijssoort for the pooled specification
Z_visa_field_agg <- field_nat_bl_agg %>%
  inner_join(Consulates, by = c("country_std" = "country")) %>%
  filter(!is.na(acceptance_rate), !is.na(share_jf)) %>%
  group_by(field, year) %>%
  summarise(Z_visa_field = sum(share_jf * acceptance_rate, na.rm = TRUE),
            .groups = "drop")
```

# 5. Construct Instruments

## 5.1 Instrument 1: Visa Acceptance Rates (all countries)

```{r instr-visa}
# -- Shift-share instrument: Schengen visa acceptance x nationality share ------
# Inner join: keep only countries that appear in BOTH the IND visa data AND
# the CBS nationality shares. This gives us, for each matched country, the visa
# acceptance rate and its pre-period share of Dutch residence permits.
visa_merged <- Consulates %>%
  inner_join(nat_shares, by = c("country" = "country_std")) %>%
  filter(!is.na(acceptance_rate), !is.na(share))

# Compute the year-level instrument Z_visa as:
#   Z_visa[t] = sum_over_countries( share_c x acceptance_rate_c,t )
# This is a Bartik-style average: 2010 share weights x time-varying visa acceptance.
Z_visa_year <- visa_merged %>%
  group_by(year) %>%
  summarise(Z_visa = sum(share * acceptance_rate, na.rm = TRUE), .groups = "drop")
```

# 6. LISS Field-of-Study Mapping

# 7. LISS Panel Data

```{r process-liss-work}
# -- LISS field-of-study code -> CBS field group map -----------------------------
# LISS uses numeric codes (011-027) for fields of study. We map each code to the
# corresponding CBS field group (onderdeel) and a boolean STEM indicator.
# Codes 017-019 (natuur, techniek, landbouw) are classified as STEM.
field_map <- tibble(
  code = sprintf("%03d", 11:27),          # Generate codes "011", "012", ... "027"
  field = c(NA, "education", "arts_languages", "arts_languages",
                "behavior_society", "business_admin", "law",
                "natural_sciences", "engineering", "agriculture",
                "healthcare", "behavior_society",
                "behavior_society", "engineering", "engineering",
                "law", "services"),
  stem = c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,              # Codes 011-017 are non-STEM
           TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)  # 018-020 = STEM; rest non-STEM
)

# -- process_early(): parse LISS 2020-2021 (comma-separated, column-name-based) --
process_early <- function(df, yr) {
  # Column 1 is the respondent identifier (nomem_encr --- encrypted membership number)
  names(df)[1] <- "nomem_encr"

  # Columns 12-28 are the field-of-study binary indicators (one column per field code)
  field_cols <- names(df)[12:28]
  # The corresponding field codes (011-027) as zero-padded 3-character strings
  field_codes <- sprintf("%03d", 11:27)

  # Extract only the field-of-study columns into a matrix (preserving row order)
  mat <- df[, field_cols, drop = FALSE]
  # LISS codes 1 = "currently following" and 2 = "completed" -> both count as studying the field.
  # Convert each column to TRUE/FALSE: TRUE if the value is 1 or 2.
  mat[] <- lapply(mat, function(x) as.integer(x) %in% c(1, 2))

  # For each row (respondent), find the FIRST field column that is TRUE,
  # and return the corresponding field code. If none selected, return NA.
  sel <- apply(mat, 1, function(r) { idx <- which(r); if (length(idx) == 0) NA_character_ else field_codes[idx[1]] })

  # -- Education attainment mapping --------------------------------------------
  # LISS asks "What is the highest level of education completed with a diploma?"
  # We convert the text answer to TWO independent variables:
  #   1. education_years --- estimated total years of schooling (for Mincer equation)
  #   2. level --- WO (research university) vs HBO (applied sciences) classification
  # Both come from the SAME raw survey answer, but are derived SEPARATELY.
  #
  # Years mapping based on convention in published LISS-using papers
  # (Leuven, Oosterbeek & van der Klaauw 2013; Bozza 2022):
  #   Compulsory schooling starts at age 6 -> primary = 6 years (ages 6-12)
  edu_col <- grep("highest level.*education.*completed.*diploma", names(df), value = TRUE, ignore.case = TRUE)[1]
  edu_raw <- as.character(df[[edu_col]])
  # Initialise as NA for all rows
  edu_yrs <- rep(NA_real_, nrow(df))
  # Map LISS education categories to Dutch system years (convention from published literature):
  edu_yrs[grepl("did not complete|^no$|^none$", edu_raw, ignore.case = TRUE)] <- 0  # No diploma
  edu_yrs[grepl("^primary school", edu_raw, ignore.case = TRUE)] <- 6               # Basisschool (6 yrs from age 6)
  edu_yrs[grepl("vmbo|mulo|lower technical|^lbo|vglo", edu_raw, ignore.case = TRUE)] <- 10  # VMBO (6+4)
  edu_yrs[grepl("^havo|^hbs|^mms", edu_raw, ignore.case = TRUE)] <- 11              # HAVO (6+5)
  edu_yrs[grepl("^vwo|^gymnasium", edu_raw, ignore.case = TRUE)] <- 12              # VWO (6+6)
  edu_yrs[grepl("^mbo |^kmbo", edu_raw, ignore.case = TRUE)] <- 13                  # MBO (6+6+1)
  edu_yrs[grepl("^mbo-plus", edu_raw, ignore.case = TRUE)] <- 14                    # MBO+ (6+6+2)
  edu_yrs[grepl("^hbo", edu_raw, ignore.case = TRUE)] <- 16                         # HBO (VWO 12+4, via MBO 13+3)
  edu_yrs[grepl("academic.*bachelor", edu_raw, ignore.case = TRUE)] <- 15            # WO bachelor (VWO 12+3)
  edu_yrs[grepl("academic.*master", edu_raw, ignore.case = TRUE)] <- 17              # WO master (15+2)
  edu_yrs[grepl("^academic|^university", edu_raw, ignore.case = TRUE)] <- 15         # WO (unspecified = bachelor)
  edu_yrs[grepl("doctor|phd", edu_raw, ignore.case = TRUE)] <- 21                    # PhD (17+4)
  # Additional mappings for values present in the LISS early-wave data:
  edu_yrs[grepl("teacher training", edu_raw, ignore.case = TRUE)] <- 16              # Pabo / teacher training = HBO (16 yrs)
  edu_yrs[grepl("conservatory|art academy", edu_raw, ignore.case = TRUE)] <- 15      # Conservatorium = WO bachelor (15 yrs)
  edu_yrs[grepl("special education", edu_raw, ignore.case = TRUE)] <- 10            # Special education = VMBO level (10 yrs)

  # -- Education level (HBO vs WO) from raw education answer --------------------
  onderwijssoort <- rep(NA_character_, nrow(df))
  onderwijssoort[grepl("hbo", edu_raw, ignore.case = TRUE)] <- "Hoger beroepsonderwijs"
  onderwijssoort[grepl("academic.*bachelor|academic.*master|^academic|^university|doctor|phd",
                       edu_raw, ignore.case = TRUE)] <- "Wetenschappelijk onderwijs"

  # -- Usual hours worked per week ---------------------------------------------
  # LISS asks "How many hours do you actually work on average?"
  hrs_col <- grep("actually work", names(df), value = TRUE, ignore.case = TRUE)[1]
  hours_worked <- if (!is.na(hrs_col)) as.numeric(df[[hrs_col]]) else NA_real_

  # -- Tenure (years with current employer) ------------------------------------
  # LISS asks for the year the respondent started working for their current employer.
  ten_col <- grep("year.*(current employer|enter into employment|start.*employment)", names(df), value = TRUE, ignore.case = TRUE)[1]
  start_year <- if (!is.na(ten_col)) as.numeric(df[[ten_col]]) else NA_real_
  # Tenure = survey year - start year, but only if the start year is plausible (> 1900, < survey year)
  tenure_years <- ifelse(!is.na(start_year) & start_year > 1900 & start_year < yr, yr - start_year, NA_real_)

  # Return a clean data frame with only the columns we need (no raw survey columns)
  data.frame(nomem_encr = as.character(df$nomem_encr), survey_year = yr,
             liss_code = sel, age = as.numeric(df[[6]]),        # Column 6 contains age
             education_years = edu_yrs,
             onderwijssoort = onderwijssoort,
             hours_worked = hours_worked,
             tenure_years = tenure_years,
             stringsAsFactors = FALSE)
}

# -- process_late(): parse LISS 2022-2024 (semicolon-separated, prefixed columns) --
process_late <- function(df, yr) {
  # Late waves use a prefixed naming scheme: cwXXyNNN where XX = year suffix, y = wave letter, NNN = variable code.
  # We auto-detect the wave letter by finding the age column (which always ends in "003").
  nm <- names(df)
  age_idx <- grep("003$", nm)[1]           # First column ending in "003" = age column
  wave_letter <- substr(nm[age_idx], 5, 5)  # Extract wave letter (e.g., "p" from "cw23p003")

  # Prefix = "cw" + last 2 digits of year + wave letter -> e.g. "cw23p"
  prefix <- paste0("cw", substr(yr, 3, 4), wave_letter)
  # Age column = prefix + "003"
  age_col <- paste0(prefix, "003")

  # -- Field-of-study detection ------------------------------------------------
  # Field columns follow the pattern cwXXy011 through cwXXy027 (codes 011-027).
  # Regex: "^" + prefix + "0[12][0-9]$" matches "cw23p011", "cw23p012", ..., "cw23p027"
  fcols <- nm[grep(paste0("^", prefix, "0[12][0-9]$"), nm)]
  # Extract the last 3 digits of each column name to get the field code (e.g. "011")
  fcodes <- sub(".*([0-9]{3})$", "\\1", fcols)

  # Loop over respondents row-by-row (each respondent can select only one field)
  nr <- nrow(df)
  sel <- rep(NA_character_, nr)             # Initialise field selection as NA
  for (i in seq_len(nr)) {
    # Get the values for all field columns for this respondent
    vals <- as.integer(df[i, fcols])
    # Find which codes have value 1 (currently following) or 2 (completed)
    idx <- which(vals %in% c(1, 2))
    # If any field selected, take the first one
    if (length(idx) > 0) sel[i] <- fcodes[idx[1]]
  }

  # -- Education: column cwXX525 (LISS standard education code 1-13) ----------
  edu_col <- paste0(prefix, "525")
  if (edu_col %in% nm) {
    edu_code <- as.numeric(df[[edu_col]])
    edu_yrs <- rep(NA_real_, nr)
    # Map LISS standard education codes 1-13 to years of education.
    # Convention from published LISS papers (Leuven, Oosterbeek & van der Klaauw 2013;
    # Bozza 2022): compulsory schooling starts at age 6, so primary = 6 years.
    # Dutch pathways: primary(6) + secondary + tertiary.
    edu_yrs[edu_code == 1] <- 0     # No education / primary not completed
    edu_yrs[edu_code == 2] <- 6     # Primary (basisschool, ages 6-12)
    edu_yrs[edu_code == 3] <- 10    # VMBO / lower secondary (6+4)
    edu_yrs[edu_code == 4] <- 11    # HAVO / upper secondary (6+5)
    edu_yrs[edu_code == 5] <- 12    # VWO / pre-university (6+6)
    edu_yrs[edu_code == 6] <- 13    # MBO / intermediate vocational (12+1, or 10+3)
    edu_yrs[edu_code == 7] <- 16    # HBO / professional bachelor (12+4 via VWO, 13+3 via MBO)
    edu_yrs[edu_code == 8] <- 15    # WO bachelor / research university (12+3)
    edu_yrs[edu_code %in% 9:10] <- 17   # WO master (15+2) --- code 9; doctor (code 10) listed here but corrected below
    edu_yrs[edu_code == 10] <- 21   # PhD / doctorate (17+4)
    edu_yrs[edu_code >= 11] <- 21   # Higher (postdoc / professorship) --- cap at PhD level
  } else { edu_yrs <- rep(NA_real_, nr) }

  # -- Education level (HBO vs WO) from standardized education code ------------
  if (exists("edu_code") && length(edu_code) > 0) {
    onderwijssoort <- rep(NA_character_, nr)
    onderwijssoort[edu_code == 7] <- "Hoger beroepsonderwijs"
    onderwijssoort[edu_code %in% 8:10] <- "Wetenschappelijk onderwijs"
  } else {
    onderwijssoort <- rep(NA_character_, nr)
  }

  # -- Hours: column cwXX126 (actual hours worked per week) --------------------
  hrs_col <- paste0(prefix, "126")
  hours_worked <- if (hrs_col %in% nm) as.numeric(df[[hrs_col]]) else rep(NA_real_, nr)

  # -- Tenure: column cwXX149 (year started with current employer) -------------
  ten_col <- paste0(prefix, "149")
  start_year_lt <- if (ten_col %in% nm) as.numeric(df[[ten_col]]) else rep(NA_real_, nr)
  tenure_years <- ifelse(!is.na(start_year_lt) & start_year_lt > 1900 & start_year_lt < yr,
                         yr - start_year_lt, NA_real_)

  # Return a clean data frame (same structure as process_early)
  data.frame(
    nomem_encr = as.character(df[[1]]),
    survey_year = yr,
    liss_code = sel,
    age = as.numeric(df[[age_col]]),
    education_years = edu_yrs,
    onderwijssoort = onderwijssoort,
    hours_worked = hours_worked,
    tenure_years = tenure_years,
    stringsAsFactors = FALSE
  )
}

# -- Apply the appropriate parser to each year and stack results ---------------
# 2020-2021 use early-format parsers; 2022-2024 use late-format parsers.
work_all <- bind_rows(process_early(LISS2020WS, 2020), process_early(LISS2021WS, 2021),
                      process_late(LISS2022WS, 2022), process_late(LISS2023WS, 2023),
                      process_late(LISS2024WS, 2024))
```

```{r process-liss-income}
# -- get_income(): extract gross monthly wage and partner status ---------------
# Early waves (<= 2021) use column-name-based detection; later waves (>= 2022) use
# the prefixed naming scheme (ciXXy, where ci = "income" module).
get_income <- function(df, yr) {
  if (yr <= 2021) {
    # -- Early-wave parser (column names) --------------------------------------
    names(df)[1] <- "nomem_encr"  # First column = respondent ID
    # Find "gross wages in total" column (exact phrasing varies slightly by year)
    gc <- grep("gross wages in total", names(df), value = TRUE, ignore.case = TRUE)[1]
    # Find "lives together with partner" column
    partner_col <- grep("household head lives together with a partner", names(df), value = TRUE, ignore.case = TRUE)[1]

    data.frame(nomem_encr = as.character(df$nomem_encr), survey_year = yr,
               gross_monthly_income = if (!is.na(gc)) as.numeric(df[[gc]]) else NA_real_,
               partner = if (!is.na(partner_col)) as.numeric(df[[partner_col]]) == 1 else NA,
               stringsAsFactors = FALSE)
  } else {
    # -- Late-wave parser (prefixed columns) -----------------------------------
    nm <- names(df)
    # Find the column ending in "372" (income variable code) to extract wave letter
    wave_letter <- substr(nm[grep("372$", nm)[1]], 5, 5)
    # Prefix = "ci" + last 2 digits of year + wave letter (e.g. "ci24q")
    prefix <- paste0("ci", substr(as.character(yr), 3, 4), wave_letter)
    # Income column = prefix + "372" (gross monthly wages)
    inc_col <- paste0(prefix, "372")
    # Partner column = prefix + "003" (household composition)
    partner_col <- paste0(prefix, "003")

    data.frame(nomem_encr = as.character(df[[1]]), survey_year = yr,
               gross_monthly_income = if (inc_col %in% nm) as.numeric(df[[inc_col]]) else NA_real_,
               partner = if (partner_col %in% nm) as.numeric(df[[partner_col]]) == 1 else NA,
               stringsAsFactors = FALSE)
  }
}

# -- Apply income parser to all 5 years and stack -----------------------------
income_all <- bind_rows(get_income(LISS2020ES, 2020), get_income(LISS2021ES, 2021),
                        get_income(LISS2022ES, 2022), get_income(LISS2023ES, 2023),
                        get_income(LISS2024ES, 2024))
```

```{r merge-liss}
# -- Merge work data + income data + field map into a single panel ------------
LISS_panel <- work_all %>%
  # Inner join: keep only respondents present in BOTH work and income modules
  inner_join(income_all, by = c("nomem_encr", "survey_year")) %>%
  # Map LISS numeric field code -> CBS field group + STEM indicator
  left_join(field_map, by = c("liss_code" = "code")) %>%
  # Drop rows missing essential variables (can't estimate without these)
  filter(!is.na(field), !is.na(gross_monthly_income), gross_monthly_income > 0) %>%
  mutate(
    log_income = log(gross_monthly_income),           # Log transform for wage equation (Mincer)
    age_sq = age^2,                                     # Quadratic age term (non-linear experience effects)
    hours_sq = hours_worked^2,                          # Quadratic hours term (non-linear labour supply)
    partner = as.logical(partner),                      # Convert partner indicator to logical TRUE/FALSE
    # Impute missing education years with the sample median (attenuates toward zero -> conservative)
    education_years = ifelse(is.na(education_years), median(education_years, na.rm = TRUE),
                             education_years)
  )
```

# 7. Merge All Data

```{r merge-all}
# -- Build the final estimation dataset ---------------------------------------
est_data <- LISS_panel %>%
  # Merge field-level international share (endogenous variable)
  # Merge level-specific share_int (matches by field x onderwijssoort x year)
  # Merge level-specific share_int (matches by field x onderwijssoort x year)
  left_join(field_shares, by = c("field", "onderwijssoort", "survey_year" = "year")) %>%
  # Merge aggregate share_int for the joint specification
  left_join(field_shares_agg %>% rename(share_int_agg = share_int),
            by = c("field", "survey_year" = "year")) %>%
  # Merge level-specific instrument
  left_join(Z_visa_field_year, by = c("field", "onderwijssoort", "survey_year" = "year")) %>%
  # Merge aggregate instrument for the joint specification
  left_join(Z_visa_field_agg %>% rename(Z_visa_field_agg = Z_visa_field),
            by = c("field", "survey_year" = "year"))

est_data <- est_data %>%
  # Drop any row missing a key variable (complete-case analysis)
  # Filter on aggregate vars for the main joint specification;
  # level-specific (share_int, Z_visa_field) may have NAs for non-matched onderwijssoort
  filter(!is.na(share_int_agg), !is.na(Z_visa_field_agg), !is.na(log_income),
         !is.na(education_years), !is.na(hours_worked), !is.na(partner))
```

# 8. Descriptive Statistics

```{r descriptives}
# -- Summary statistics table --------------------------------------------------
# Mean (SD) shown in each row.
est_data %>%
  summarise(N = sprintf("%d", n()),
            `Log Income` = sprintf("%.3f (%.3f)", mean(log_income), sd(log_income)),
            `Age` = sprintf("%.1f (%.1f)", mean(age), sd(age)),
            `Education` = sprintf("%.1f (%.1f)", mean(education_years), sd(education_years)),
            `Hours` = sprintf("%.1f (%.1f)", mean(hours_worked, na.rm = TRUE), sd(hours_worked, na.rm = TRUE)),
            `Partner` = sprintf("%.2f (%.2f)", mean(partner, na.rm = TRUE), sd(partner, na.rm = TRUE)),
            `Share Intl` = sprintf("%.3f (%.3f)", mean(share_int_agg), sd(share_int_agg)),
            `Z_visa_field` = sprintf("%.4f (%.4f)", mean(Z_visa_field_agg, na.rm = TRUE), sd(Z_visa_field_agg, na.rm = TRUE)),
            `STEM` = sprintf("%.2f (%.2f)", mean(stem), sd(stem))) %>%
  # Reshape to vertical format (one row per statistic)
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>%
  # Render as a publication-quality LaTeX table
  kable(caption = "Descriptive Statistics", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

# Remaining sections (First Stage, Main Results, Heterogeneity,
# All Specifications, Standardized, Lagged Instruments) removed
# for clean rebuild. Add new regression sections below.
