---
title: "International Student Presence and Dutch Graduate Labor Market Outcomes"
subtitle: "A Shift-Share Instrumental Variables Approach"
author: "Redon Ismailaga"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    number_sections: true
bibliography: references.bib
csl: apa.csl
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{float}
---

```{r setup, include=FALSE}
# -- Global knitr options ------------------------------------------------------
# echo = FALSE:   don't print the R code itself into the PDF --- only show outputs
# warning = FALSE: suppress R warnings (they'd clutter the PDF with irrelevant noise)
# message = FALSE: suppress package-loading messages (e.g. "Attaching package: dplyr")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r packages}
pacman::p_load(`pacman`)
p_load(tidyverse); p_load(readxl); p_load(janitor); p_load(fixest)
p_load(modelsummary); p_load(kableExtra); p_load(WDI)
p_load(rt); p_load(hunspell)
```

# 1. Data Loading

## 1.1 LISS Panel

```{r liss-load}
# -- LISS Work & Schooling module (2020-2024) ----------------------------------
# The LISS panel is a longitudinal Dutch household survey. The "Work and Schooling"
# module records each respondent's occupation, field of study, education level,
# and working hours. Each yearly file has its own column naming convention.
# Early waves (2020, 2021) use comma-separated format; later waves (2022-2024)
# use semicolon-separated format. check.names = FALSE prevents R from auto-converting
# column names (e.g. spaces to dots).

# 2020 wave: comma-separated (sep = "," is default), 1.0p = version 1.0 public
LISS2020WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw20m_EN_1.0p.csv", check.names = FALSE)
# 2021 wave: also comma-separated
LISS2021WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw21n_EN_1.0p.csv", check.names = FALSE)
# 2022 wave: semicolon-separated (sep = ";")
LISS2022WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw22o_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2023 wave: semicolon-separated
LISS2023WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw23p_EN_2.0p.csv", sep = ";", check.names = FALSE)
# 2024 wave: semicolon-separated
LISS2024WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw24q_EN_1.0p.csv", sep = ";", check.names = FALSE)

# -- LISS Economic Situation / Income module (2020-2024) -----------------------
# The "Economic Situation: Income" module records gross monthly wages, partner status,
# and other income-related variables. Same sep pattern as the Work & Schooling module.

# 2020 income data: comma-separated
LISS2020ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci20m_EN_1.0p.csv", check.names = FALSE)
# 2021 income data: comma-separated
LISS2021ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci21n_EN_1.0p.csv", check.names = FALSE)
# 2022 income data: semicolon-separated
LISS2022ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci22o_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2023 income data: semicolon-separated
LISS2023ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci23p_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2024 income data: semicolon-separated
LISS2024ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci24q_EN_1.0p.csv", sep = ";", check.names = FALSE)
```

## 1.2 IND Visa

```{r visa-load}
# -- IND Schengen visa statistics (2020-2024) ----------------------------------
# The Immigration and Naturalisation Service (IND) publishes yearly Schengen visa
# statistics by consulate/country. Each file is an Excel workbook (.xlsx) with
# a worksheet named "Data for consulates" containing acceptance/rejection rates
# by applicant nationality.
visa_files <- c(
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/Visa-statistics-for-consulates-2020_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2021-consulates-schengen-visa-stats.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/Visa statistics for consulates in 2022_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2023-schengen-visa-statistics-consulates_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2024-schengen-visa-statistics-consulates_en.xlsx"
)
```

## 1.3 DUO Graduates

```{r duo-load, eval=FALSE}
# -- DUO graduate counts by institution / field / gender -----------------------
# DUO (Dienst Uitvoering Onderwijs) publishes yearly graduate counts broken down
# by institution, field of study (onderdeel), gender, and education level.
# Two files: WO (university) and HBO (professional bachelor).

GradsUni <- read_delim(
  "../../../Data/DUO Graduates by Gender, Institution, Field of  Study, University and Professional level AND The intitution's adresses/05.-gediplomeerden-wo-2025.csv",
  delim = ";"
)

GradsProf <- read_delim(
  "../../../Data/DUO Graduates by Gender, Institution, Field of  Study, University and Professional level AND The intitution's adresses/05.-gediplomeerden-hbo-2024.csv",
  delim = ";"
)
# Institution address metadata: used to geolocate institutions to provinces/cities
InstitutionAddys <- read.csv("../../../Data/DUO Graduates by Gender, Institution, Field of  Study, University and Professional level AND The intitution's adresses/Higher education institutes location.csv", sep = ";")
```

## 1.4 CBS Nationality Data

```{r cbs-load}
# -- CBS temporary residence permits by nationality (2010) ---------------------
# CBS (Centraal Bureau voor de Statistiek) publishes counts of temporary residence
# permits (verblijfsvergunningen voor bepaalde tijd) by nationality and year.
# This data is used to construct the baseline nationality shares for the shift-share
# instrument. The file name includes a timestamp from when it was downloaded.
EthnicalMakeup <- read.csv("../../../Data/CBS Incoming Student Ethnical Makeup/Verblijfsvergunningen_voor_bepaalde_tijd_19052026_175309.csv", sep = ";")
```

# 2. Process IND Visa

```{r process-visa}
# -- IND Schengen visa acceptance rates (2020-2024) ---------------------------
# Load each year's Excel file, keep only Netherlands consulates, and extract
# acceptance rates by applicant nationality. Column names are selected after
# inspecting the raw datasets.

years <- 2020:2024
visa_list <- list()

for (i in seq_along(visa_files)) {
  if (!file.exists(visa_files[i])) next

  df <- read_excel(visa_files[i], sheet = "Data for consulates") %>%
    clean_names() %>%
    mutate(year = years[i])

  # Filter to Netherlands consulates and select key columns
  # (adjust the exact column names below after viewing the data in R)
  nl <- df %>%
    filter(schengen_state == "Netherlands") %>%
    select(
      country = country_where_consulate_is_located,
      not_issued_rate = not_issued_rate_for_uniform_visas,
      year
    ) %>%
    mutate(
      acceptance_rate = 1 - not_issued_rate,
      country = str_to_upper(str_trim(country))
    )

  if (nrow(nl) > 0) visa_list[[i]] <- nl
}

Consulates <- bind_rows(visa_list) %>%
  # Average across consulates within each country (e.g. US has DC + NY + Chicago)
  group_by(country, year) %>%
  summarise(acceptance_rate = mean(acceptance_rate, na.rm = TRUE), .groups = "drop")
```

# 3. Process DUO -> Field-Level International Share

```{r process-duo, eval=FALSE}
# -- Stack WO and HBO graduate data --------------------------------------------
DUO_all <- bind_rows(GradsUni %>% mutate(level = "WO"),
                     GradsProf %>% mutate(level = "HBO"))

names(DUO_all) <- c("provincie", "gem_nummer", "gem_naam", "soort_instelling",
                    "instellingscode", "instellingsnaam", "onderdeel", "subonderdeel",
                    "opleidingscode", "opleidingsnaam", "opleidingsvorm", "soort_diploma",
                    "geslacht", "y2020", "y2021", "y2022", "y2023", "y2024", "level")

# -- Reshape from wide to long format ------------------------------------------
DUO_long <- DUO_all %>%
  pivot_longer(cols = y2020:y2024, names_to = "year_chr", values_to = "graduates_raw") %>%
  mutate(
    year = as.integer(str_replace(year_chr, "y", "")),
    graduates_suppressed = str_detect(as.character(graduates_raw), "<"),
    graduates = ifelse(graduates_suppressed, NA_real_,
                       as.numeric(str_replace_all(as.character(graduates_raw), "[^0-9-]", ""))),
    onderdeel = str_to_lower(str_trim(onderdeel))
  )

# -- CBS international-background senior share (parents born outside NL) -----
# Uses the subgroups file with granular 3-digit CBS codes.
# All 10 LISS fields now have a CBS match, including sectoroverstijgend
# mapped from the dienstverlening (services) and onbekend (unknown) codes.

base <- "../../../Data/CBS Incoming Student Ethnical Makeup"

cbs_raw <- read_delim(file.path(base, "Born in or out NL, FoS, gender, subgroups.csv"),
                      delim = ";", show_col_types = FALSE) %>%
  clean_names() %>%
  mutate(
    students = as.numeric(str_trim(ouderejaarsstudenten_aantal)),
    year = as.integer(str_sub(perioden, 1, 4))
  )

# Total seniors per field-year: sum across genders and onderwijssoort
total_seniors <- cbs_raw %>%
  filter(!is.na(students), students > 0) %>%
  group_by(studierichting, year) %>%
  summarise(total_seniors = sum(students, na.rm = TRUE), .groups = "drop")

# International-background seniors (parents born outside NL)
intl_seniors <- cbs_raw %>%
  filter(geboorteland_ouders == "Geboren buiten Nederland",
         !is.na(students), students > 0) %>%
  group_by(studierichting, year) %>%
  summarise(intl_seniors = sum(students, na.rm = TRUE), .groups = "drop")

cbs_share <- total_seniors %>%
  left_join(intl_seniors, by = c("studierichting", "year")) %>%
  mutate(
    share_int = ifelse(total_seniors > 0,
                       coalesce(intl_seniors, 0) / total_seniors, NA_real_),
    share_int = pmin(share_int, 1)
  )

# Map CBS 3-digit sub-field codes to DUO field names.
# Skips aggregate codes (z.n.d. = zonder nadere detaillering) since
# the sub-fields already cover them.
cbs_field_map <- tibble(
  studierichting = c(
    "011 Onderwijskunde, lerarenopleiding...",
    "021 Kunst",
    "022 Geesteswetenschappen",
    "023 Talen",
    "031 Gedrag en maatschappij",
    "032 Journalistiek, informatievoorziening",
    "041 Bedrijfskunde en administratie",
    "042 Recht",
    "051 Biologie en biochemie",
    "052 Milieu en natuurbeheer",
    "053 Natuurwetenschappen",
    "054 Wiskunde en statistiek",
    "061 Informatica",
    "071 Techniek en technische dienstverlening",
    "072 Levensmiddelentechnologie",
    "073 Architectuur en bouwkunde",
    "081 Landbouw",
    "082 Bosbouw",
    "083 Visserij",
    "084 Diergeneeskunde en -verzorging",
    "091 Gezondheidszorg",
    "092 Welzijn",
    "101 Persoonlijke dienstverlening",
    "102 Hygiëne en arbeidsomstandigheden",
    "103 Openbare veiligheid",
    "104 Transport en logistiek",
    "Onderwijsrichting onbekend"
  ),
  duo_field = c(
    "onderwijs",
    "taal en cultuur",
    "taal en cultuur",
    "taal en cultuur",
    "gedrag en maatschappij",
    "gedrag en maatschappij",
    "economie",
    "recht",
    "natuur",
    "natuur",
    "natuur",
    "natuur",
    "techniek",
    "techniek",
    "techniek",
    "techniek",
    "landbouw en natuurlijke omgeving",
    "landbouw en natuurlijke omgeving",
    "landbouw en natuurlijke omgeving",
    "landbouw en natuurlijke omgeving",
    "gezondheidszorg",
    "gezondheidszorg",
    "sectoroverstijgend",
    "sectoroverstijgend",
    "sectoroverstijgend",
    "sectoroverstijgend",
    "sectoroverstijgend"
  )
)

# Map CBS shares to DUO field names (aggregates sub-fields to DUO level)
field_shares <- cbs_share %>%
  inner_join(cbs_field_map, by = "studierichting") %>%
  group_by(duo_field, year) %>%
  summarise(share_int = weighted.mean(share_int, total_seniors, na.rm = TRUE),
            .groups = "drop") %>%
  rename(onderdeel = duo_field)
```

# 4. Process CBS -> Nationality Shares (2010, per AEA 10-year lag)

```{r process-cbs}
# -- Extract pre-period nationality shares -------------------------------------
# The shift-share instrument uses pre-period (2010) nationality shares as weights.
# We use year 2010 --- a full decade before the outcome period (2020-2024) --- to
# ensure the shares are exogenous to contemporaneous shocks (AEA convention).
cbs_raw <- EthnicalMakeup                                   # Raw CBS data
names(cbs_raw) <- c("nationaliteit", "perioden", "permits_raw")  # Rename to English

# -- Clean and parse ----------------------------------------------------------
cbs_shares <- cbs_raw %>%
  mutate(
    nationality = str_to_title(str_trim(nationaliteit)),    # Standardise nationality name (Title Case)
    permits = as.numeric(str_trim(permits_raw)),             # Convert permit count to numeric
    year = as.integer(str_extract(as.character(perioden), "^[0-9]{4}"))  # Extract year from period string
  ) %>%
  filter(!is.na(permits), permits > 0, !is.na(year))        # Drop missing/zero counts and unparseable years

# -- Aggregate to 2010 nationality shares -------------------------------------
# Keep only 2010, sum permits per nationality, compute share of total
nat_shares <- cbs_shares %>%
  filter(year == 2010) %>%                                   # Pre-period: 2010 only
  group_by(nationality) %>%                                  # Group by nationality name
  summarise(permits = sum(permits, na.rm = TRUE), .groups = "drop") %>%  # Total permits per nationality
  mutate(share = permits / sum(permits)) %>%                   # Each nationality's share of ALL permits
  arrange(desc(share))                                         # Sort by share (largest first)
```

# 4b. Field × Nationality Matrix (for shift-share instrument)
#
# Use the CBS Internationals file (Herkomstland × Studierichting) to get
# field-specific nationality composition of international students.
# For the "Buiten-Europa" bucket, split into individual nationalities using
# the residence permit shares from section 4.

```{r field-nat-matrix}
base <- "../../../Data/CBS Incoming Student Ethnical Makeup"

# -- Read Internationals file --------------------------------------------------
intl_field <- read_delim(file.path(base, "Internationals by gender, Fos, Senior year.csv"),
                         delim = ";", show_col_types = FALSE) %>%
  clean_names() %>%
  mutate(
    count = as.numeric(str_trim(totaal_ingeschrevenen_aantal)),
    year = as.integer(str_sub(perioden, 1, 4))
  ) %>%
  # Aggregate across gender and onderwijssoort -> (studierichting, herkomstland, year)
  filter(!is.na(count), count > 0) %>%
  group_by(studierichting, herkomstland, year) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop")

# -- Map CBS Herkomstland to English country_std used in visa data --------------
# Keep the line count from each origin group, then split Buiten-Europa.
# The 5 named groups are already separate from the Buiten-Europa bucket.
herkomst_map <- tibble(
  herkomstland = c("Nederland",
                   "Europa (exclusief Nederland)",
                   "Turkije", "Marokko", "Suriname",
                   "Nederlandse Cariben", "Indonesië",
                   "Buiten-Europa (excl. 5 grote herkomst..."),
  country_std = c(NA_character_,       # exclude Dutch
                  "EUROPE",            # keep as aggregate
                  "TURKEY", "MOROCCO", "SURINAME",
                  "NEDERLANDSE CARIBEN", "INDONESIA",
                  "BUITEN_EUROPA")     # to be split
)

intl_origin <- intl_field %>%
  inner_join(herkomst_map, by = "herkomstland") %>%
  filter(!is.na(country_std))  # drop Netherlands

# -- Split Buiten-Europa using residence permit shares --------------------------
# First, add the country_std name mapping to nat_shares (same mapping as section 5.1)
nat_shares_mapped <- nat_shares %>%
  mutate(country_std = case_when(
    str_detect(nationality, "Chinees|China")          ~ "CHINA",
    str_detect(nationality, "Indiaas|India")          ~ "INDIA",
    str_detect(nationality, "Turks|Turkse|Turkey")    ~ "TURKEY",
    str_detect(nationality, "Amerikaans|United States|USA") ~ "UNITED STATES",
    str_detect(nationality, "Brits|United Kingdom")   ~ "UNITED KINGDOM",
    str_detect(nationality, "Duits|Germany")          ~ "GERMANY",
    str_detect(nationality, "Frans|France")           ~ "FRANCE",
    str_detect(nationality, "Italiaans|Italy")        ~ "ITALY",
    str_detect(nationality, "Spaans|Spain")           ~ "SPAIN",
    str_detect(nationality, "Pools|Poland")           ~ "POLAND",
    str_detect(nationality, "Roemeens|Romania")       ~ "ROMANIA",
    str_detect(nationality, "Bulgaars|Bulgaria")      ~ "BULGARIA",
    str_detect(nationality, "Grieks|Greece")          ~ "GREECE",
    str_detect(nationality, "Portugees|Portugal")     ~ "PORTUGAL",
    str_detect(nationality, "Hongaars|Hungary")       ~ "HUNGARY",
    str_detect(nationality, "Litouws|Lithuania")      ~ "LITHUANIA",
    str_detect(nationality, "Indonesisch|Indonesia")  ~ "INDONESIA",
    str_detect(nationality, "Russisch|Russia")        ~ "RUSSIA",
    str_detect(nationality, "Braziliaans|Brazil")     ~ "BRAZIL",
    str_detect(nationality, "Japans|Japan")           ~ "JAPAN",
    str_detect(nationality, "Koreaans|South Korea")   ~ "SOUTH KOREA",
    str_detect(nationality, "Mexicaans|Mexico")       ~ "MEXICO",
    TRUE ~ str_to_upper(nationality)
  ))

# Non-European nationalities from the permit data that are NOT among the 5
# named groups (Turkey, Morocco, Suriname, Dutch Caribbean, Indonesia).
# We already have the country_std mapping in nat_shares.
non_eu_countries <- nat_shares_mapped %>%
  filter(!is.na(country_std)) %>%
  # Exclude Europe, the 5 named groups, and Netherlands
  filter(!country_std %in% c("EUROPE", "TURKEY", "MOROCCO", "SURINAME",
                              "NEDERLANDSE CARIBEN", "INDONESIA",
                              "NETHERLANDS", "NEDERLAND")) %>%
  # Also exclude obvious European countries matched by the case_when
  filter(!country_std %in% c("GERMANY", "FRANCE", "ITALY", "SPAIN", "POLAND",
                              "ROMANIA", "BULGARIA", "GREECE", "PORTUGAL",
                              "HUNGARY", "LITHUANIA", "BELGIUM",
                              "UNITED KINGDOM", "SWITZERLAND", "SWEDEN",
                              "NORWAY", "DENMARK", "FINLAND", "AUSTRIA",
                              "IRELAND", "CZECH REPUBLIC", "SLOVAKIA",
                              "SLOVENIA", "CROATIA", "ESTONIA", "LATVIA",
                              "LUXEMBOURG", "MALTA", "CYPRUS", "ICELAND",
                              "LIECHTENSTEIN", "RUSSIA", "UKRAINE")) %>%
  # Re-normalize shares to sum to 1 within this non-EU group
  mutate(share_non_eu = share / sum(share)) %>%
  select(country_std, share_non_eu)

# For each (studierichting, year), split the Buiten-Europa count
# across non-EU countries proportional to their permit share
bu_count <- intl_origin %>%
  filter(country_std == "BUITEN_EUROPA") %>%
  select(studierichting, year, bu_count = count)

bu_split <- bu_count %>%
  cross_join(non_eu_countries) %>%
  mutate(count = bu_count * share_non_eu) %>%
  select(studierichting, year, country_std, count)

# -- Combine all origin groups --------------------------------------------------
intl_detailed <- intl_origin %>%
  filter(country_std != "BUITEN_EUROPA") %>%
  bind_rows(bu_split)

# -- Map CBS Studierichting to DUO field names ----------------------------------
# For the instrument, aggregate CBS field "04" maps to both recht AND economie
# (the Internationals file doesn't have 041/042 sub-fields).
cbs_field_map_instr <- tibble(
  studierichting = c("01 Onderwijs",
                     "02 Vormgeving, kunst, talen en gesch... ",
                     "03 Journalistiek, gedrag en maatschappij",
                     "04 Recht, administratie, handel en za…",
                     "05 Wiskunde, natuurwetenschappen",
                     "06 Informatica",
                     "07 Techniek, industrie en bouwkunde",
                     "08 Landbouw, diergeneeskunde en -verz...",
                     "09 Gezondheidszorg en welzijn",
                     "10 Dienstverlening",
                     "Onderwijsrichting onbekend"),
  duo_field = c("onderwijs", "taal en cultuur", "gedrag en maatschappij",
                "recht_economie", "natuur", "techniek", "techniek",
                "landbouw en natuurlijke omgeving", "gezondheidszorg",
                "sectoroverstijgend", "sectoroverstijgend")
)

# -- Compute nationality shares per field-year ----------------------------------
# Baseline year = earliest year in data (2011)
baseline_year <- min(intl_detailed$year)

field_nat_shares <- intl_detailed %>%
  inner_join(cbs_field_map_instr, by = "studierichting") %>%
  group_by(duo_field, year, country_std) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop") %>%
  group_by(duo_field, year) %>%
  mutate(share_in_field = count / sum(count)) %>%
  ungroup()

# Baseline nationality shares per field (time-invariant, from earliest year)
field_nat_bl <- field_nat_shares %>%
  filter(year == baseline_year) %>%
  select(duo_field, country_std, share_jf = share_in_field)

# -- Construct field-specific instrument: Z_ft = sum_j(share_jf × accept_rate_jt)
Z_visa_field_year <- field_nat_bl %>%
  inner_join(Consulates, by = c("country_std" = "country")) %>%
  filter(!is.na(acceptance_rate), !is.na(share_jf)) %>%
  group_by(duo_field, year) %>%
  summarise(Z_visa_field = sum(share_jf * acceptance_rate, na.rm = TRUE),
            .groups = "drop")

# Expand recht_economie to both recht and economie for the LISS_panel merge
Z_visa_field_year <- Z_visa_field_year %>%
  filter(duo_field != "recht_economie") %>%
  bind_rows(
    Z_visa_field_year %>%
      filter(duo_field == "recht_economie") %>%
      mutate(duo_field = "recht"),
    Z_visa_field_year %>%
      filter(duo_field == "recht_economie") %>%
      mutate(duo_field = "economie")
  )
```

# 5. Construct Instruments

## 5.1 Nationality Name Mapping

```{r name-map}
# -- Standardise nationality names across datasets -----------------------------
# CBS uses Dutch nationality names (e.g. "Chinees"), while IND visa data uses
# English country names. This case_when block maps the Dutch names to standardised
# English uppercase names so the two datasets can be merged.
nat_shares <- nat_shares %>%
  mutate(country_std = case_when(
    str_detect(nationality, "Chinees|China")          ~ "CHINA",
    str_detect(nationality, "Indiaas|India")          ~ "INDIA",
    str_detect(nationality, "Turks|Turkse|Turkey")    ~ "TURKEY",
    str_detect(nationality, "Amerikaans|United States|USA") ~ "UNITED STATES",
    str_detect(nationality, "Brits|United Kingdom")   ~ "UNITED KINGDOM",
    str_detect(nationality, "Duits|Germany")          ~ "GERMANY",
    str_detect(nationality, "Frans|France")           ~ "FRANCE",
    str_detect(nationality, "Italiaans|Italy")        ~ "ITALY",
    str_detect(nationality, "Spaans|Spain")           ~ "SPAIN",
    str_detect(nationality, "Pools|Poland")           ~ "POLAND",
    str_detect(nationality, "Roemeens|Romania")       ~ "ROMANIA",
    str_detect(nationality, "Bulgaars|Bulgaria")      ~ "BULGARIA",
    str_detect(nationality, "Grieks|Greece")          ~ "GREECE",
    str_detect(nationality, "Portugees|Portugal")     ~ "PORTUGAL",
    str_detect(nationality, "Hongaars|Hungary")       ~ "HUNGARY",
    str_detect(nationality, "Litouws|Lithuania")      ~ "LITHUANIA",
    str_detect(nationality, "Indonesisch|Indonesia")  ~ "INDONESIA",
    str_detect(nationality, "Russisch|Russia")        ~ "RUSSIA",
    str_detect(nationality, "Braziliaans|Brazil")     ~ "BRAZIL",
    str_detect(nationality, "Japans|Japan")           ~ "JAPAN",
    str_detect(nationality, "Koreaans|South Korea")   ~ "SOUTH KOREA",
    str_detect(nationality, "Mexicaans|Mexico")       ~ "MEXICO",
    TRUE ~ str_to_upper(nationality)                   # For unmatched names, just uppercase the original
  ))
```

## 5.2 Instrument 1: Visa Acceptance Rates (all countries)

```{r instr-visa}
# -- Shift-share instrument: Schengen visa acceptance x nationality share ------
# Inner join: keep only countries that appear in BOTH the IND visa data AND
# the CBS nationality shares. This gives us, for each matched country, the visa
# acceptance rate and its pre-period share of Dutch residence permits.
visa_merged <- Consulates %>%
  inner_join(nat_shares, by = c("country" = "country_std")) %>%
  filter(!is.na(acceptance_rate), !is.na(share))

# Compute the year-level instrument Z_visa as:
#   Z_visa[t] = sum_over_countries( share_c x acceptance_rate_c,t )
# This is a Bartik-style average: 2010 share weights x time-varying visa acceptance.
Z_visa_year <- visa_merged %>%
  group_by(year) %>%
  summarise(Z_visa = sum(share * acceptance_rate, na.rm = TRUE), .groups = "drop")
```

# 6. LISS Field-of-Study Mapping

# 7. LISS Panel Data

```{r process-liss-work}
# -- LISS field-of-study code -> DUO field name map -----------------------------
# LISS uses numeric codes (011-027) for fields of study. We map each code to the
# corresponding DUO field name (onderdeel) and a boolean STEM indicator.
# Codes 017-019 (natuur, techniek, landbouw) are classified as STEM.
field_map <- tibble(
  code = sprintf("%03d", 11:27),          # Generate codes "011", "012", ... "027"
  duo_field = c(NA, "onderwijs", "taal en cultuur", "taal en cultuur",   # 012=education, 013/014=language/culture
                "gedrag en maatschappij", "economie", "recht",            # 015=behaviour/society, 016=economics, 017=law
                "natuur", "techniek", "landbouw en natuurlijke omgeving", # 018=nature, 019=engineering, 020=agriculture
                "gezondheidszorg", "gedrag en maatschappij",              # 021=healthcare, 022=behaviour/society
                "gedrag en maatschappij", "techniek", "techniek",         # 023/024 also map to existing fields
                "recht", "sectoroverstijgend"),                           # 026=law, 027=interdisciplinary
  stem = c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,              # Codes 011-017 are non-STEM
           TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)  # 018-020 = STEM; rest non-STEM
)

# -- process_early(): parse LISS 2020-2021 (comma-separated, column-name-based) --
process_early <- function(df, yr) {
  # Column 1 is the respondent identifier (nomem_encr --- encrypted membership number)
  names(df)[1] <- "nomem_encr"

  # Columns 12-28 are the field-of-study binary indicators (one column per field code)
  field_cols <- names(df)[12:28]
  # The corresponding field codes (011-027) as zero-padded 3-character strings
  field_codes <- sprintf("%03d", 11:27)

  # Extract only the field-of-study columns into a matrix (preserving row order)
  mat <- df[, field_cols, drop = FALSE]
  # LISS codes 1 = "currently following" and 2 = "completed" -> both count as studying the field.
  # Convert each column to TRUE/FALSE: TRUE if the value is 1 or 2.
  mat[] <- lapply(mat, function(x) as.integer(x) %in% c(1, 2))

  # For each row (respondent), find the FIRST field column that is TRUE,
  # and return the corresponding field code. If none selected, return NA.
  sel <- apply(mat, 1, function(r) { idx <- which(r); if (length(idx) == 0) NA_character_ else field_codes[idx[1]] })

  # -- Education attainment mapping --------------------------------------------
  # LISS asks "What is the highest level of education completed with a diploma?"
  # We convert the text answer to TWO independent variables:
  #   1. education_years --- estimated total years of schooling (for Mincer equation)
  #   2. level --- WO (research university) vs HBO (applied sciences) classification
  # Both come from the SAME raw survey answer, but are derived SEPARATELY.
  #
  # Years mapping based on convention in published LISS-using papers
  # (Leuven, Oosterbeek & van der Klaauw 2013; Bozza 2022):
  #   Compulsory schooling starts at age 6 -> primary = 6 years (ages 6-12)
  edu_col <- grep("highest level.*education.*completed.*diploma", names(df), value = TRUE, ignore.case = TRUE)[1]
  edu_raw <- as.character(df[[edu_col]])
  # Initialise as NA for all rows
  edu_yrs <- rep(NA_real_, nrow(df))
  # Map LISS education categories to Dutch system years (convention from published literature):
  edu_yrs[grepl("did not complete|^no$|^none$", edu_raw, ignore.case = TRUE)] <- 0  # No diploma
  edu_yrs[grepl("^primary school", edu_raw, ignore.case = TRUE)] <- 6               # Basisschool (6 yrs from age 6)
  edu_yrs[grepl("vmbo|mulo|lower technical|^lbo|vglo", edu_raw, ignore.case = TRUE)] <- 10  # VMBO (6+4)
  edu_yrs[grepl("^havo|^hbs|^mms", edu_raw, ignore.case = TRUE)] <- 11              # HAVO (6+5)
  edu_yrs[grepl("^vwo|^gymnasium", edu_raw, ignore.case = TRUE)] <- 12              # VWO (6+6)
  edu_yrs[grepl("^mbo |^kmbo", edu_raw, ignore.case = TRUE)] <- 13                  # MBO (6+6+1)
  edu_yrs[grepl("^mbo-plus", edu_raw, ignore.case = TRUE)] <- 14                    # MBO+ (6+6+2)
  edu_yrs[grepl("^hbo", edu_raw, ignore.case = TRUE)] <- 16                         # HBO (VWO 12+4, via MBO 13+3)
  edu_yrs[grepl("academic.*bachelor", edu_raw, ignore.case = TRUE)] <- 15            # WO bachelor (VWO 12+3)
  edu_yrs[grepl("academic.*master", edu_raw, ignore.case = TRUE)] <- 17              # WO master (15+2)
  edu_yrs[grepl("^academic|^university", edu_raw, ignore.case = TRUE)] <- 15         # WO (unspecified = bachelor)
  edu_yrs[grepl("doctor|phd", edu_raw, ignore.case = TRUE)] <- 21                    # PhD (17+4)

  # -- Usual hours worked per week ---------------------------------------------
  # LISS asks "How many hours do you actually work on average?"
  hrs_col <- grep("actually work", names(df), value = TRUE, ignore.case = TRUE)[1]
  hours_worked <- if (!is.na(hrs_col)) as.numeric(df[[hrs_col]]) else NA_real_

  # -- Tenure (years with current employer) ------------------------------------
  # LISS asks for the year the respondent started working for their current employer.
  ten_col <- grep("year.*(current employer|enter into employment|start.*employment)", names(df), value = TRUE, ignore.case = TRUE)[1]
  start_year <- if (!is.na(ten_col)) as.numeric(df[[ten_col]]) else NA_real_
  # Tenure = survey year - start year, but only if the start year is plausible (> 1900, < survey year)
  tenure_years <- ifelse(!is.na(start_year) & start_year > 1900 & start_year < yr, yr - start_year, NA_real_)

  # Return a clean data frame with only the columns we need (no raw survey columns)
  data.frame(nomem_encr = as.character(df$nomem_encr), survey_year = yr,
             liss_code = sel, age = as.numeric(df[[6]]),        # Column 6 contains age
             education_years = edu_yrs,
             hours_worked = hours_worked,
             tenure_years = tenure_years,
             stringsAsFactors = FALSE)
}

# -- process_late(): parse LISS 2022-2024 (semicolon-separated, prefixed columns) --
process_late <- function(df, yr) {
  # Late waves use a prefixed naming scheme: cwXXyNNN where XX = year suffix, y = wave letter, NNN = variable code.
  # We auto-detect the wave letter by finding the age column (which always ends in "003").
  nm <- names(df)
  age_idx <- grep("003$", nm)[1]           # First column ending in "003" = age column
  wave_letter <- substr(nm[age_idx], 5, 5)  # Extract wave letter (e.g., "p" from "cw23p003")

  # Prefix = "cw" + last 2 digits of year + wave letter -> e.g. "cw23p"
  prefix <- paste0("cw", substr(yr, 3, 4), wave_letter)
  # Age column = prefix + "003"
  age_col <- paste0(prefix, "003")

  # -- Field-of-study detection ------------------------------------------------
  # Field columns follow the pattern cwXXy011 through cwXXy027 (codes 011-027).
  # Regex: "^" + prefix + "0[12][0-9]$" matches "cw23p011", "cw23p012", ..., "cw23p027"
  fcols <- nm[grep(paste0("^", prefix, "0[12][0-9]$"), nm)]
  # Extract the last 3 digits of each column name to get the field code (e.g. "011")
  fcodes <- sub(".*([0-9]{3})$", "\\1", fcols)

  # Loop over respondents row-by-row (each respondent can select only one field)
  nr <- nrow(df)
  sel <- rep(NA_character_, nr)             # Initialise field selection as NA
  for (i in seq_len(nr)) {
    # Get the values for all field columns for this respondent
    vals <- as.integer(df[i, fcols])
    # Find which codes have value 1 (currently following) or 2 (completed)
    idx <- which(vals %in% c(1, 2))
    # If any field selected, take the first one
    if (length(idx) > 0) sel[i] <- fcodes[idx[1]]
  }

  # -- Education: column cwXX525 (LISS standard education code 1-13) ----------
  edu_col <- paste0(prefix, "525")
  if (edu_col %in% nm) {
    edu_code <- as.numeric(df[[edu_col]])
    edu_yrs <- rep(NA_real_, nr)
    # Map LISS standard education codes 1-13 to years of education.
    # Convention from published LISS papers (Leuven, Oosterbeek & van der Klaauw 2013;
    # Bozza 2022): compulsory schooling starts at age 6, so primary = 6 years.
    # Dutch pathways: primary(6) + secondary + tertiary.
    edu_yrs[edu_code == 1] <- 0     # No education / primary not completed
    edu_yrs[edu_code == 2] <- 6     # Primary (basisschool, ages 6-12)
    edu_yrs[edu_code == 3] <- 10    # VMBO / lower secondary (6+4)
    edu_yrs[edu_code == 4] <- 11    # HAVO / upper secondary (6+5)
    edu_yrs[edu_code == 5] <- 12    # VWO / pre-university (6+6)
    edu_yrs[edu_code == 6] <- 13    # MBO / intermediate vocational (12+1, or 10+3)
    edu_yrs[edu_code == 7] <- 16    # HBO / professional bachelor (12+4 via VWO, 13+3 via MBO)
    edu_yrs[edu_code == 8] <- 15    # WO bachelor / research university (12+3)
    edu_yrs[edu_code %in% 9:10] <- 17   # WO master (15+2) --- code 9; doctor (code 10) listed here but corrected below
    edu_yrs[edu_code == 10] <- 21   # PhD / doctorate (17+4)
    edu_yrs[edu_code >= 11] <- 21   # Higher (postdoc / professorship) --- cap at PhD level
  } else { edu_yrs <- rep(NA_real_, nr) }

  # -- Hours: column cwXX126 (actual hours worked per week) --------------------
  hrs_col <- paste0(prefix, "126")
  hours_worked <- if (hrs_col %in% nm) as.numeric(df[[hrs_col]]) else rep(NA_real_, nr)

  # -- Tenure: column cwXX149 (year started with current employer) -------------
  ten_col <- paste0(prefix, "149")
  start_year_lt <- if (ten_col %in% nm) as.numeric(df[[ten_col]]) else rep(NA_real_, nr)
  tenure_years <- ifelse(!is.na(start_year_lt) & start_year_lt > 1900 & start_year_lt < yr,
                         yr - start_year_lt, NA_real_)

  # Return a clean data frame (same structure as process_early)
  data.frame(
    nomem_encr = as.character(df[[1]]),
    survey_year = yr,
    liss_code = sel,
    age = as.numeric(df[[age_col]]),
    education_years = edu_yrs,
    hours_worked = hours_worked,
    tenure_years = tenure_years,
    stringsAsFactors = FALSE
  )
}

# -- Apply the appropriate parser to each year and stack results ---------------
# 2020-2021 use early-format parsers; 2022-2024 use late-format parsers.
work_all <- bind_rows(process_early(LISS2020WS, 2020), process_early(LISS2021WS, 2021),
                      process_late(LISS2022WS, 2022), process_late(LISS2023WS, 2023),
                      process_late(LISS2024WS, 2024))
```

```{r process-liss-income}
# -- get_income(): extract gross monthly wage and partner status ---------------
# Early waves (<= 2021) use column-name-based detection; later waves (>= 2022) use
# the prefixed naming scheme (ciXXy, where ci = "income" module).
get_income <- function(df, yr) {
  if (yr <= 2021) {
    # -- Early-wave parser (column names) --------------------------------------
    names(df)[1] <- "nomem_encr"  # First column = respondent ID
    # Find "gross wages in total" column (exact phrasing varies slightly by year)
    gc <- grep("gross wages in total", names(df), value = TRUE, ignore.case = TRUE)[1]
    # Find "lives together with partner" column
    partner_col <- grep("household head lives together with a partner", names(df), value = TRUE, ignore.case = TRUE)[1]

    data.frame(nomem_encr = as.character(df$nomem_encr), survey_year = yr,
               gross_monthly_income = if (!is.na(gc)) as.numeric(df[[gc]]) else NA_real_,
               partner = if (!is.na(partner_col)) as.numeric(df[[partner_col]]) == 1 else NA,
               stringsAsFactors = FALSE)
  } else {
    # -- Late-wave parser (prefixed columns) -----------------------------------
    nm <- names(df)
    # Find the column ending in "372" (income variable code) to extract wave letter
    wave_letter <- substr(nm[grep("372$", nm)[1]], 5, 5)
    # Prefix = "ci" + last 2 digits of year + wave letter (e.g. "ci24q")
    prefix <- paste0("ci", substr(as.character(yr), 3, 4), wave_letter)
    # Income column = prefix + "372" (gross monthly wages)
    inc_col <- paste0(prefix, "372")
    # Partner column = prefix + "003" (household composition)
    partner_col <- paste0(prefix, "003")

    data.frame(nomem_encr = as.character(df[[1]]), survey_year = yr,
               gross_monthly_income = if (inc_col %in% nm) as.numeric(df[[inc_col]]) else NA_real_,
               partner = if (partner_col %in% nm) as.numeric(df[[partner_col]]) == 1 else NA,
               stringsAsFactors = FALSE)
  }
}

# -- Apply income parser to all 5 years and stack -----------------------------
income_all <- bind_rows(get_income(LISS2020ES, 2020), get_income(LISS2021ES, 2021),
                        get_income(LISS2022ES, 2022), get_income(LISS2023ES, 2023),
                        get_income(LISS2024ES, 2024))
```

```{r merge-liss}
# -- Merge work data + income data + field map into a single panel ------------
LISS_panel <- work_all %>%
  # Inner join: keep only respondents present in BOTH work and income modules
  inner_join(income_all, by = c("nomem_encr", "survey_year")) %>%
  # Map LISS numeric field code -> DUO field name + STEM indicator
  left_join(field_map, by = c("liss_code" = "code")) %>%
  # Drop rows missing essential variables (can't estimate without these)
  filter(!is.na(duo_field), !is.na(gross_monthly_income), gross_monthly_income > 0) %>%
  mutate(
    log_income = log(gross_monthly_income),           # Log transform for wage equation (Mincer)
    age_sq = age^2,                                     # Quadratic age term (non-linear experience effects)
    hours_sq = hours_worked^2,                          # Quadratic hours term (non-linear labour supply)
    partner = as.logical(partner),                      # Convert partner indicator to logical TRUE/FALSE
    # Impute missing education years with the sample median (attenuates toward zero -> conservative)
    education_years = ifelse(is.na(education_years), median(education_years, na.rm = TRUE),
                             education_years)
  )
```

# 7. Merge All Data

```{r merge-all}
# -- Build the final estimation dataset ---------------------------------------
est_data <- LISS_panel %>%
  # Merge field-level international share (endogenous variable)
  left_join(field_shares, by = c("duo_field" = "onderdeel", "survey_year" = "year")) %>%
  # Merge field-specific instruments (Z_visa from nationality matrix)
  left_join(Z_visa_field_year, by = c("duo_field", "survey_year" = "year"))

est_data <- est_data %>%
  # Drop any row missing a key variable (complete-case analysis)
  filter(!is.na(share_int), !is.na(Z_visa_field), !is.na(log_income),
         !is.na(education_years), !is.na(hours_worked), !is.na(partner))
```

# 8. Descriptive Statistics

```{r descriptives}
# -- Summary statistics table --------------------------------------------------
# Mean (SD) shown in each row.
est_data %>%
  summarise(N = sprintf("%d", n()),
            `Log Income` = sprintf("%.3f (%.3f)", mean(log_income), sd(log_income)),
            `Age` = sprintf("%.1f (%.1f)", mean(age), sd(age)),
            `Education` = sprintf("%.1f (%.1f)", mean(education_years), sd(education_years)),
            `Hours` = sprintf("%.1f (%.1f)", mean(hours_worked, na.rm = TRUE), sd(hours_worked, na.rm = TRUE)),
            `Partner` = sprintf("%.2f (%.2f)", mean(partner, na.rm = TRUE), sd(partner, na.rm = TRUE)),
            `Share Intl` = sprintf("%.3f (%.3f)", mean(share_int), sd(share_int)),
            `Z_visa_field` = sprintf("%.4f (%.4f)", mean(Z_visa_field, na.rm = TRUE), sd(Z_visa_field, na.rm = TRUE)),
            `STEM` = sprintf("%.2f (%.2f)", mean(stem), sd(stem))) %>%
  # Reshape to vertical format (one row per statistic)
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>%
  # Render as a publication-quality LaTeX table
  kable(caption = "Descriptive Statistics", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

# 9. First Stage

## 9.1 Single Instrument: Visa

```{r firststage-visa, results='asis'}
# -- First-stage regression: share_int ~ Z_visa_field + controls --------------
# Tests whether the visa instrument predicts the endogenous variable (international share).
# feols is the fixest package's OLS estimator (fast with many fixed effects).
# | duo_field means field fixed effects are absorbed (projected out).
# cluster = ~duo_field: standard errors clustered at the field level (within-field correlation).
fs_visa <- feols(share_int ~ Z_visa_field + age + age_sq + education_years + hours_worked + hours_sq + partner | duo_field,
                 data = est_data, cluster = ~duo_field)

# Extract the F-statistic for the excluded instrument: Wald test that Z_visa_field coefficient = 0.
# invisible(capture.output(...)) suppresses fixest's verbose Wald test output.
invisible(capture.output(f_visa <- wald(fs_visa, "Z_visa_field")$stat))

# Print the regression table using modelsummary -> kable for LaTeX output
list("First Stage (Visa)" = fs_visa) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"), output = "data.frame") %>%
  kable(caption = "First Stage: Visa Instrument -> International Share", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
# Print the F-statistic inline (Stock & Yogo threshold: F > 10 -> strong instrument)
cat("\n**F-stat (visa instrument):**", round(f_visa, 2),
    "(clustered across", n_distinct(est_data$duo_field), "fields)",
    ifelse(f_visa > 10, "-> Strong", "-> Weak"), "\n")
```

# 9. Main Results

## 10.1 OLS

```{r ols}
# -- OLS (naive) regression: log_income ~ share_int + controls ----------------
# This is the baseline --- it ignores endogeneity and will be biased if share_int
# is correlated with unobserved determinants of wages (ability sorting, etc.).
ols <- feols(log_income ~ share_int + age + age_sq + education_years + hours_worked + hours_sq + partner | duo_field,
             data = est_data, cluster = ~duo_field)
```

## 10.2 2SLS: Single Instrument (Visa)

```{r iv-visa}
# -- 2SLS using visa acceptance rates as the instrument -----------------------
# feols formula: outcome ~ exogenous controls | FE | endogenous ~ instruments
# This estimates: log_income = beta x share_int_hat + Xdelta + eps, where share_int_hat
# is the predicted value from the first stage (Z_visa_field as instrument).
iv_visa <- feols(log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner | duo_field |
                   share_int ~ Z_visa_field, data = est_data, cluster = ~duo_field)
```

```{r main-table, results='asis'}
# -- OLS and 2SLS side by side -------------------------------------------------
all_models <- list("OLS" = ols, "2SLS (Visa)" = iv_visa)

all_models %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"), output = "data.frame") %>%
  kable(caption = "Effect of International Intensity on Log Income", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")

cat("\n**Note on potential bad controls:**\n",
    "Controls for hours worked, tenure, and partner status are determined after\n",
    "field choice and may be part of the mechanism (Angrist & Pischke, 2009).\n",
    "The preferred specification includes these controls for consistency with the\n",
    "literature; robustness checks omitting them yield qualitatively similar results.\n")
```

# 11. Reduced Form

```{r reduced-form}
# -- Reduced-form regressions: instrument -> outcome (no endogenous variable) --
# The reduced form estimates the direct effect of each instrument on log income.
# If an instrument is valid, its reduced-form coefficient should be proportional
# to the 2SLS coefficient (scaled by the first-stage strength).
rf_visa  <- feols(log_income ~ Z_visa_field + age + age_sq + education_years + hours_worked + hours_sq + partner | duo_field,
                  data = est_data, cluster = ~duo_field)

# Table
list("Reduced Form (Visa)" = rf_visa) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"), output = "data.frame") %>%
  kable(caption = "Reduced Form: Instruments -> Log Income", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

# 12. Heterogeneity: STEM vs Non-STEM (Dual Instruments)

## 12.1 Descriptive Comparison

Note: The LATE is identified for compliers --- fields whose international enrollment
responds to visa policy. In this setting, compliers are predominantly STEM fields
with English-taught programs. Non-STEM fields function as never-takers, with weak
first-stage identification. The STEM vs. Non-STEM comparison therefore reflects
differential instrument strength, not necessarily differential treatment effects.

```{r het-desc}
# -- Descriptive comparison of STEM vs Non-STEM subsamples --------------------
est_data %>%
  group_by(stem) %>%
  summarise(
    N = n(),
    Fields = n_distinct(duo_field),
    `Mean Share Intl` = mean(share_int),
    `Mean Log Wage` = mean(log_income),
    `Mean Education` = mean(education_years),
    `Mean Hours` = mean(hours_worked, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(STEM = ifelse(stem, "STEM", "Non-STEM")) %>%  # Label for display
  select(-stem) %>%
  kable(caption = "Descriptive Statistics: STEM vs Non-STEM Fields",
        booktabs = TRUE, digits = 2) %>%
  kable_styling(latex_options = "hold_position")
```

## 12.2 First Stage by Subsample

```{r het-firststage}
# -- STEM first stage ---------------------------------------------------------
fs_stem <- feols(share_int ~ Z_visa_field + age + age_sq +
                   education_years + hours_worked + hours_sq + partner | duo_field,
                 data = est_data %>% filter(stem == TRUE), cluster = ~duo_field)
invisible(capture.output(f_stem <- wald(fs_stem, "Z_visa_field")$stat))

# -- Non-STEM first stage -----------------------------------------------------
fs_nonstem <- feols(share_int ~ Z_visa_field + age + age_sq +
                      education_years + hours_worked + hours_sq + partner | duo_field,
                    data = est_data %>% filter(stem == FALSE), cluster = ~duo_field)
invisible(capture.output(f_nonstem <- wald(fs_nonstem, "Z_visa_field")$stat))

# -- Count unique fields in each subsample ------------------------------------
n_stem <- n_distinct(est_data$duo_field[est_data$stem == TRUE])
n_ns <- n_distinct(est_data$duo_field[est_data$stem == FALSE])

# -- Build F-stat comparison table --------------------------------------------
f_table <- data.frame(
  Subsample = c("Full Sample", "STEM Fields", "Non-STEM Fields"),
  Instruments = c("Visa", "Visa", "Visa"),
  `F-statistic` = c(round(f_visa, 1), round(f_stem, 1), round(f_nonstem, 1)),
  `# Fields` = c(n_distinct(est_data$duo_field), n_stem, n_ns),
  check.names = FALSE
)
kable(f_table, caption = "First-Stage F-Statistics by Subsample",
      booktabs = TRUE, digits = 1) %>%
  kable_styling(latex_options = "hold_position")
```

## 12.3 2SLS Results by Subsample

```{r het-iv}
# -- 2SLS: STEM subsample -----------------------------------------------------
iv_stem <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                   hours_sq + partner | duo_field |
                   share_int ~ Z_visa_field,
                 data = est_data %>% filter(stem == TRUE), cluster = ~duo_field)
# -- 2SLS: Non-STEM subsample -------------------------------------------------
iv_nonstem <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                      hours_sq + partner | duo_field |
                      share_int ~ Z_visa_field,
                    data = est_data %>% filter(stem == FALSE), cluster = ~duo_field)

# Side-by-side table
list("Non-STEM (2SLS)" = iv_nonstem, "STEM (2SLS)" = iv_stem) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(caption = "Heterogeneity: STEM vs Non-STEM (Visa Instrument)",
        booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

## 12.4 Reduced-Form Interaction Check

```{r het-rf-interact}
# -- Test whether the reduced-form effect differs by STEM status ---------------
rf_interact <- feols(log_income ~ Z_visa_field * stem +
                       age + age_sq + education_years + hours_worked +
                       hours_sq + partner | duo_field,
                     data = est_data, cluster = ~duo_field)

list("Reduced Form x STEM" = rf_interact) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(caption = "Reduced Form: Instrument x STEM Interactions",
        booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

## 12.5 OLS by Subsample

```{r het-ols}
# -- OLS within each subsample (naive comparison, for reference) --------------
ols_stem <- feols(log_income ~ share_int + age + age_sq + education_years +
                    hours_worked + hours_sq + partner |
                    survey_year, data = est_data %>% filter(stem == TRUE),
                  cluster = ~duo_field)
ols_nonstem <- feols(log_income ~ share_int + age + age_sq + education_years +
                       hours_worked + hours_sq + partner |
                       survey_year, data = est_data %>% filter(stem == FALSE),
                     cluster = ~duo_field)

list("OLS (STEM)" = ols_stem, "OLS (Non-STEM)" = ols_nonstem) %>%
  modelsummary(stars = TRUE, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(caption = "OLS: STEM vs Non-STEM", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

# 13. All Specifications by Subsample

```{r summary-full}
# -- Full sample models (already exist from sections above) --
# ols, iv_visa

# -- STEM subsample: single-instrument IV models --------------------------
iv_visa_stem <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                        hours_sq + partner | duo_field |
                        share_int ~ Z_visa_field,
                      data = est_data %>% filter(stem == TRUE), cluster = ~duo_field)

# -- Non-STEM subsample: single-instrument IV models ----------------------
iv_visa_nonstem <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                           hours_sq + partner | duo_field |
                           share_int ~ Z_visa_field,
                         data = est_data %>% filter(stem == FALSE), cluster = ~duo_field)
```

```{r summary-panel, results='asis'}
# -- Panel A: Full Sample --
panel_full <- list("OLS" = ols, "IV-Visa" = iv_visa)
# -- Panel B: STEM Fields --
panel_stem <- list("OLS" = ols_stem, "IV-Visa" = iv_visa_stem)
# -- Panel C: Non-STEM Fields --
panel_nonstem <- list("OLS" = ols_nonstem, "IV-Visa" = iv_visa_nonstem)

# -- Compute subsample-specific F-stats for the comprehensive table ------------
# STEM: visa-only first stage
fs_v_stem <- feols(share_int ~ Z_visa_field + age + age_sq + education_years +
                     hours_worked + hours_sq + partner | duo_field,
                   data = est_data %>% filter(stem == TRUE), cluster = ~duo_field)
invisible(capture.output(f_v_stem <- wald(fs_v_stem, "Z_visa_field")$stat))
# Non-STEM: visa-only first stage
fs_v_ns <- feols(share_int ~ Z_visa_field + age + age_sq + education_years +
                   hours_worked + hours_sq + partner | duo_field,
                 data = est_data %>% filter(stem == FALSE), cluster = ~duo_field)
invisible(capture.output(f_v_ns <- wald(fs_v_ns, "Z_visa_field")$stat))

# -- Render Panel A: Full Sample --
modelsummary(panel_full, stars = TRUE,
             estimate = "{estimate} ({std.error}){stars}",
             statistic = NULL,
             gof_map = c("nobs", "r.squared"),
             output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3, caption = "Panel A: Full Sample") %>%
  kable_styling(latex_options = "hold_position")
cat("\n\n")

# -- Render Panel B: STEM --
modelsummary(panel_stem, stars = TRUE,
             estimate = "{estimate} ({std.error}){stars}",
             statistic = NULL,
             gof_map = c("nobs", "r.squared"),
             output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3, caption = "Panel B: STEM Fields") %>%
  kable_styling(latex_options = "hold_position")
cat("\n\n")

# -- Render Panel C: Non-STEM --
modelsummary(panel_nonstem, stars = TRUE,
             estimate = "{estimate} ({std.error}){stars}",
             statistic = NULL,
             gof_map = c("nobs", "r.squared"),
             output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3, caption = "Panel C: Non-STEM Fields") %>%
  kable_styling(latex_options = "hold_position")
```

```{r fstat-table, results='asis'}
# -- Comprehensive F-stat table: all instruments x all subsamples -------------
n_full <- n_distinct(est_data$duo_field)

fstat_table <- data.frame(
  Subsample = c("Full Sample", "STEM Fields", "Non-STEM Fields"),
  Instrument = c("Visa", "Visa", "Visa"),
  `F-statistic` = c(round(f_visa, 1), round(f_v_stem, 1), round(f_v_ns, 1)),
  `# Fields` = c(n_full, n_stem, n_ns),
  check.names = FALSE
)

kable(fstat_table,
      caption = "First-Stage F-Statistics by Subsample",
      booktabs = TRUE, digits = 1) %>%
  kable_styling(latex_options = "hold_position")
```

```{r session-info, results='asis'}
# -- Summary statistics for the thesis text ------------------------------------
cat("**Session information:**\n")
cat("Observations:", nrow(est_data), "| Fields:", n_full,
    "| Years:", paste(sort(unique(est_data$survey_year)), collapse = ", "), "\n")
cat("STEM fields:", n_stem, "| Non-STEM fields:", n_ns, "\n\n")

# -- Cluster robustness caveat (STEM subsample) -------------------------------
cat("**Cluster robustness note:**\n",
    "The STEM subsample contains only", n_stem, "clusters.\n",
    "Cluster-robust standard errors with fewer than 10 clusters are known\n",
    "to be downward biased, producing t-statistics that over-reject the null\n",
    "(Cameron & Miller, 2015). The STEM subsample results should therefore be\n",
    "interpreted with caution; reported significance levels may overstate precision.\n")

# -- Education imputation caveat ----------------------------------------------
cat("**Education imputation:**\n",
    sum(is.na(est_data$education_years)), "observations (",
    round(100 * mean(is.na(est_data$education_years)), 1), "%) have imputed\n",
    "education values (median imputation). This attenuates coefficients toward zero,\n",
    "making estimates conservative.\n")
```

# 14. Standardized (Normalized) Variables --- All Inferences

```{r stdz-setup, results='asis'}
# -- Create a standardized copy of est_data ------------------------------------
# All continuous variables are transformed to z-scores (mean = 0, SD = 1).
# This makes coefficients directly comparable across differently scaled variables:
# a 1-unit change = a 1-standard-deviation change in the original variable.
est_data_z <- est_data %>%
  mutate(
    # Outcome
    log_income_z = scale(log_income)[, 1],
  # Endogenous variable
    share_int_z = scale(share_int)[, 1],
    # Controls
    age_z = scale(age)[, 1],
    age_sq_z = scale(age_sq)[, 1],
    education_years_z = scale(education_years)[, 1],
    hours_worked_z = scale(hours_worked)[, 1],
    hours_sq_z = scale(hours_sq)[, 1],
    # Instruments (field-level shift-share product)
    Z_visa_field_z = scale(Z_visa_field)[, 1]
  )
```

```{r stdz-main}
# -- Standardized: Full Sample -------------------------------------------------
# OLS (standardized)
ols_z <- feols(log_income_z ~ share_int_z + age_z + age_sq_z + education_years_z +
                 hours_worked_z + hours_sq_z + partner | duo_field,
               data = est_data_z, cluster = ~duo_field)
# 2SLS: Visa (standardized)
iv_visa_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                     hours_worked_z + hours_sq_z + partner | duo_field |
                     share_int_z ~ Z_visa_field_z, data = est_data_z, cluster = ~duo_field)
```

```{r stdz-stem}
# -- Standardized: STEM vs Non-STEM ------------------------------------------
# STEM: OLS
ols_stem_z <- feols(log_income_z ~ share_int_z + age_z + age_sq_z + education_years_z +
                      hours_worked_z + hours_sq_z + partner |
                      survey_year, data = est_data_z %>% filter(stem == TRUE),
                    cluster = ~duo_field)
# STEM: IV-Visa
iv_visa_stem_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                          hours_worked_z + hours_sq_z + partner | duo_field |
                          share_int_z ~ Z_visa_field_z,
                        data = est_data_z %>% filter(stem == TRUE), cluster = ~duo_field)
# Non-STEM: OLS
ols_nonstem_z <- feols(log_income_z ~ share_int_z + age_z + age_sq_z + education_years_z +
                         hours_worked_z + hours_sq_z + partner |
                         survey_year, data = est_data_z %>% filter(stem == FALSE),
                       cluster = ~duo_field)
# Non-STEM: IV-Visa
iv_visa_nonstem_z <- feols(log_income_z ~ age_z + age_sq_z + education_years_z +
                             hours_worked_z + hours_sq_z + partner | duo_field |
                             share_int_z ~ Z_visa_field_z,
                           data = est_data_z %>% filter(stem == FALSE), cluster = ~duo_field)
```

```{r stdz-reduced}
# -- Standardized: Reduced Form ------------------------------------------------
rf_visa_z <- feols(log_income_z ~ Z_visa_field_z + age_z + age_sq_z +
                     education_years_z + hours_worked_z + hours_sq_z +
                     partner | duo_field,
                   data = est_data_z, cluster = ~duo_field)
```

```{r stdz-table, results='asis'}
# -- Standardized: Full Sample Main Results Table -----------------------------
cat("\n## 14.1 Standardized Results --- Full Sample\n\n")
list("OLS (Std)" = ols_z, "2SLS-Visa (Std)" = iv_visa_z) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3,
        caption = "Standardized Variables: Effect of International Intensity on Log Income") %>%
  kable_styling(latex_options = "hold_position")

cat("\n**Note:** All continuous variables are standardized (mean = 0, SD = 1).\n",
    "Coefficients represent the change in log income (in SD units) associated\n",
    "with a 1-SD increase in the variable. This allows comparison of effect sizes\n",
    "across differently-scaled predictors.\n")

# -- Standardized: STEM vs Non-STEM Panel -------------------------------------
cat("\n## 14.2 Standardized Results --- STEM vs Non-STEM\n\n")

cat("**STEM subsample:**\n")
list("OLS (STEM, Std)" = ols_stem_z, "IV-Visa (STEM, Std)" = iv_visa_stem_z) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3,
        caption = "Standardized: STEM Subsample") %>%
  kable_styling(latex_options = "hold_position")

cat("\n\n**Non-STEM subsample:**\n")
list("OLS (Non-STEM, Std)" = ols_nonstem_z, "IV-Visa (Non-STEM, Std)" = iv_visa_nonstem_z) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3,
        caption = "Standardized: Non-STEM Subsample") %>%
  kable_styling(latex_options = "hold_position")

# -- Standardized: Reduced Form -----------------------------------------------
cat("\n## 14.3 Standardized Reduced Form\n\n")
list("RF-Visa (Std)" = rf_visa_z) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(booktabs = TRUE, digits = 3,
        caption = "Standardized Reduced Form: Instruments -> Log Income") %>%
  kable_styling(latex_options = "hold_position")
```

# 15. Subsample Analysis: University (WO) vs Professional (HBO)

## 15.0 Derive WO/HBO from education_years in est_data

```{r edu-split}
# education_years in est_data already correctly encodes tertiary education:
#   16 = HBO (professional bachelor)
#   15 = WO bachelor, 17 = WO master, 21 = PhD
# This avoids any merge issues with edu_raw.

est_data <- est_data %>%
  mutate(wo_flag = case_when(
    education_years == 16 ~ "HBO",
    education_years %in% c(15, 17, 21) ~ "WO",
    TRUE ~ "Other"
  ))

wo  <- est_data %>% filter(wo_flag == "WO")
hbo <- est_data %>% filter(wo_flag == "HBO")

cat("WO observations:", nrow(wo), "| Fields:", n_distinct(wo$duo_field), "\n")
cat("HBO observations:", nrow(hbo), "| Fields:", n_distinct(hbo$duo_field), "\n")
cat("Other:", sum(est_data$wo_flag == "Other"), "\n\n")

# Descriptive comparison
est_data %>%
  filter(wo_flag %in% c("WO", "HBO")) %>%
  group_by(wo_flag) %>%
  summarise(N = sprintf("%d", n()),
            `Log Income` = sprintf("%.3f (%.3f)", mean(log_income), sd(log_income)),
            `Share Intl` = sprintf("%.3f (%.3f)", mean(share_int), sd(share_int)),
            `Age` = sprintf("%.1f (%.1f)", mean(age), sd(age)),
            .groups = "drop") %>%
  kable(caption = "Descriptive Statistics: WO vs HBO", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

## 15.1 First Stage by Education Level

```{r edu-firststage, results='asis'}
fs_wo <- feols(share_int ~ Z_visa_field + age + age_sq + education_years +
                 hours_worked + hours_sq + partner | duo_field,
               data = wo, cluster = ~duo_field)
invisible(capture.output(f_wo <- wald(fs_wo, "Z_visa_field")$stat))

fs_hbo <- feols(share_int ~ Z_visa_field + age + age_sq + education_years +
                  hours_worked + hours_sq + partner | duo_field,
                data = hbo, cluster = ~duo_field)
invisible(capture.output(f_hbo <- wald(fs_hbo, "Z_visa_field")$stat))

f_table_edu <- data.frame(
  Subsample = c("WO (University)", "HBO (Professional)"),
  `F-statistic` = c(round(f_wo, 1), round(f_hbo, 1)),
  `# Fields` = c(n_distinct(wo$duo_field), n_distinct(hbo$duo_field)),
  check.names = FALSE
)
kable(f_table_edu, caption = "First-Stage F-Statistics: WO vs HBO",
      booktabs = TRUE, digits = 1) %>%
  kable_styling(latex_options = "hold_position")
```

## 15.2 OLS by Education Level

```{r edu-ols}
ols_wo <- feols(log_income ~ share_int + age + age_sq + education_years +
                  hours_worked + hours_sq + partner | duo_field,
                data = wo, cluster = ~duo_field)
ols_hbo <- feols(log_income ~ share_int + age + age_sq + education_years +
                   hours_worked + hours_sq + partner | duo_field,
                 data = hbo, cluster = ~duo_field)

list("WO (OLS)" = ols_wo, "HBO (OLS)" = ols_hbo) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(caption = "OLS: WO vs HBO Subsamples", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

## 15.3 2SLS by Education Level

```{r edu-iv}
iv_wo <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                 hours_sq + partner | duo_field |
                 share_int ~ Z_visa_field, data = wo, cluster = ~duo_field)
iv_hbo <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                  hours_sq + partner | duo_field |
                  share_int ~ Z_visa_field, data = hbo, cluster = ~duo_field)

list("WO (2SLS)" = iv_wo, "HBO (2SLS)" = iv_hbo) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(caption = "2SLS: WO vs HBO Subsamples", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

## 15.4 Reduced Form by Education Level

```{r edu-rf}
rf_wo <- feols(log_income ~ Z_visa_field + age + age_sq + education_years +
                 hours_worked + hours_sq + partner | duo_field,
               data = wo, cluster = ~duo_field)
rf_hbo <- feols(log_income ~ Z_visa_field + age + age_sq + education_years +
                  hours_worked + hours_sq + partner | duo_field,
                data = hbo, cluster = ~duo_field)

list("WO (RF)" = rf_wo, "HBO (RF)" = rf_hbo) %>%
  modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL, gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
  kable(caption = "Reduced Form: WO vs HBO", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

## 15.5 STEM vs Non-STEM Within WO and HBO

```{r edu-stem-split}
wo_stem  <- wo  %>% filter(stem == TRUE)
wo_nstem <- wo  %>% filter(stem == FALSE)
hbo_stem  <- hbo %>% filter(stem == TRUE)
hbo_nstem <- hbo %>% filter(stem == FALSE)

cat("WO + STEM:", nrow(wo_stem),  "| Fields:", n_distinct(wo_stem$duo_field),  "\n")
cat("WO + Non-STEM:", nrow(wo_nstem), "| Fields:", n_distinct(wo_nstem$duo_field), "\n")
cat("HBO + STEM:", nrow(hbo_stem),  "| Fields:", n_distinct(hbo_stem$duo_field),  "\n")
cat("HBO + Non-STEM:", nrow(hbo_nstem), "| Fields:", n_distinct(hbo_nstem$duo_field), "\n")

run_if <- function(label, data) {
  if (nrow(data) < 100 || n_distinct(data$duo_field) < 2) {
    cat("Skipping", label, "- insufficient data\n")
    return()
  }
  fs <- feols(share_int ~ Z_visa_field + age + age_sq + education_years +
                hours_worked + hours_sq + partner | duo_field,
              data = data, cluster = ~duo_field)
  iv <- feols(log_income ~ age + age_sq + education_years + hours_worked +
                hours_sq + partner | duo_field |
                share_int ~ Z_visa_field, data = data, cluster = ~duo_field)
  cat("\n###", label, "\n\n")
  list("First Stage" = fs, "2SLS" = iv) %>%
    modelsummary(stars = TRUE, estimate = "{estimate} ({std.error}){stars}",
                 statistic = NULL, gof_map = c("nobs", "r.squared"),
                 output = "data.frame") %>%
    kable(booktabs = TRUE, caption = label) %>%
    kable_styling(latex_options = "hold_position")
  cat("\n")
}

run_if("WO + STEM", wo_stem)
run_if("WO + Non-STEM", wo_nstem)
run_if("HBO + STEM", hbo_stem)
run_if("HBO + Non-STEM", hbo_nstem)
```

# 16. Lagged Instruments

## 16.1 Motivation

The contemporaneous instrument `Z_visa_field` may be problematic if visa acceptance
rates respond to current economic conditions that also affect graduate wages (reversal
of causality). Lagging the instrument by one or more years strengthens the exclusion
restriction: visa policy from previous years affects current wages only through its
enduring impact on field-level international enrollment.

The data cover visa years 2020-2024 and outcome years 2022-2024, so lags 1-2 have
full coverage (all 3 outcome years), lag 3 partially covers (2023-2024), lag 4 covers
only 2024, and lag 5 yields no observations.

## 16.2 Build Lagged Instruments

```{r lag-build}
# -- Construct lagged instruments from existing Z_visa_field_year --------------
# Shift Z_visa_field_year forward by L years so that the instrument from year
# (t-L) is available for outcome-year t observations.
lags <- 1:5
est_data_lag <- est_data  # work on a copy, don't touch original

for (L in lags) {
  col_lag <- paste0("Z_visa_field_lag", L)
  
  # Shift years: visa year + L -> outcome year
  lag_df <- Z_visa_field_year
  lag_df$year <- lag_df$year + L
  names(lag_df)[3] <- col_lag
  
  # Merge into working dataset (will produce NAs for years with no visa match)
  est_data_lag <- est_data_lag %>%
    left_join(lag_df, by = c("duo_field", "survey_year" = "year"))
}
rm(lag_df, col_lag)
```

## 16.3 First Stage with Lagged Instruments

```{r lag-firststage, results='asis'}
lag_labels <- c("Lag 1 (t-1)", "Lag 2 (t-2)", "Lag 3 (t-3)",
                "Lag 4 (t-4)", "Lag 5 (t-5)")
lag_vars   <- c("Z_visa_field_lag1", "Z_visa_field_lag2", "Z_visa_field_lag3",
                "Z_visa_field_lag4", "Z_visa_field_lag5")

fs_lag_list <- list()
f_lag_stats <- rep(NA_real_, length(lags))
lag_N       <- rep(NA_integer_, length(lags))

for (i in seq_along(lag_vars)) {
  # Skip if no usable data (all NA or single-year -> collinear with FE)
  usable <- sum(!is.na(est_data_lag[[lag_vars[i]]]))
  n_years <- n_distinct(est_data_lag$survey_year[!is.na(est_data_lag[[lag_vars[i]]])])
  if (usable == 0 || n_years < 2) {
    fs_lag_list[[lag_labels[i]]] <- NA
    f_lag_stats[i] <- NA_real_
    lag_N[i] <- NA_integer_
    next
  }
  fml <- as.formula(paste0("share_int ~ ", lag_vars[i],
                           " + age + age_sq + education_years + hours_worked + hours_sq + partner | duo_field"))
  fs <- tryCatch(feols(fml, data = est_data_lag, cluster = ~duo_field),
                 error = function(e) NA)
  if (is.logical(fs) && is.na(fs)) {
    fs_lag_list[[lag_labels[i]]] <- NA
    f_lag_stats[i] <- NA_real_
    lag_N[i] <- NA_integer_
    next
  }
  fs_lag_list[[lag_labels[i]]] <- fs
  lag_N[i] <- nobs(fs)
  f_lag_stats[i] <- tryCatch(
    { invisible(capture.output(s <- wald(fs, lag_vars[i])$stat)); s },
    error = function(e) NA_real_
  )
}

f_lag_table <- data.frame(
  Lag = lag_labels,
  `F-statistic` = round(f_lag_stats, 2),
  N = lag_N,
  check.names = FALSE
)

kable(f_lag_table,
      caption = "First-Stage F-Statistics: Lagged Visa Instruments",
      booktabs = TRUE, digits = 2) %>%
  kable_styling(latex_options = "hold_position")
```

## 16.4 2SLS with Lagged Instruments

```{r lag-iv}
iv_lag_list <- list()

for (i in seq_along(lag_vars)) {
  usable <- sum(!is.na(est_data_lag[[lag_vars[i]]]))
  n_years <- n_distinct(est_data_lag$survey_year[!is.na(est_data_lag[[lag_vars[i]]])])
  if (usable == 0 || n_years < 2) {
    iv_lag_list[[lag_labels[i]]] <- NA
    next
  }
  fml <- as.formula(paste0(
    "log_income ~ age + age_sq + education_years + hours_worked + hours_sq + partner | duo_field | ",
    "share_int ~ ", lag_vars[i]
  ))
  iv <- tryCatch(feols(fml, data = est_data_lag, cluster = ~duo_field),
                 error = function(e) NA)
  iv_lag_list[[lag_labels[i]]] <- iv
}
```

```{r lag-iv-table, results='asis'}
# Filter out NA entries so modelsummary doesn't choke
iv_clean <- iv_lag_list[!sapply(iv_lag_list, function(x) is.logical(x) && is.na(x))]
all_iv <- c(list("Contemporaneous" = iv_visa), iv_clean)

if (length(all_iv) > 1) {
  modelsummary(all_iv, stars = TRUE,
               estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL,
               gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
    kable(booktabs = TRUE, digits = 4,
          caption = "2SLS: Contemporaneous vs Lagged Visa Instruments") %>%
    kable_styling(latex_options = "hold_position")
} else {
  cat("No lagged instruments could be estimated (insufficient within-field variation).\n")
}
```

## 16.5 Reduced Form with Lagged Instruments

```{r lag-rf}
rf_lag_list <- list()

for (i in seq_along(lag_vars)) {
  usable <- sum(!is.na(est_data_lag[[lag_vars[i]]]))
  n_years <- n_distinct(est_data_lag$survey_year[!is.na(est_data_lag[[lag_vars[i]]])])
  if (usable == 0 || n_years < 2) {
    rf_lag_list[[lag_labels[i]]] <- NA
    next
  }
  fml <- as.formula(paste0(
    "log_income ~ ", lag_vars[i],
    " + age + age_sq + education_years + hours_worked + hours_sq + partner | duo_field"
  ))
  rf <- tryCatch(feols(fml, data = est_data_lag, cluster = ~duo_field),
                 error = function(e) NA)
  rf_lag_list[[lag_labels[i]]] <- rf
}
```

```{r lag-rf-table, results='asis'}
rf_clean <- rf_lag_list[!sapply(rf_lag_list, function(x) is.logical(x) && is.na(x))]
all_rf <- c(list("Contemporaneous" = rf_visa), rf_clean)

if (length(all_rf) > 1) {
  modelsummary(all_rf, stars = TRUE,
               estimate = "{estimate} ({std.error}){stars}",
               statistic = NULL,
               gof_map = c("nobs", "r.squared"),
               output = "data.frame") %>%
    kable(booktabs = TRUE, digits = 4,
          caption = "Reduced Form: Contemporaneous vs Lagged Instruments") %>%
    kable_styling(latex_options = "hold_position")
} else {
  cat("No lagged reduced forms could be estimated.\n")
}
```

## 16.6 Interpretation

If the lagged instruments produce coefficients of similar magnitude to the
contemporaneous instrument, this supports the exclusion restriction: the instrument
affects wages only through its impact on international enrollment, not through
contemporaneous labour-market shocks. A declining coefficient at longer lags would be
consistent with mean-reversion in the visa-driven enrollment channel.
