---
title: "International Student Presence and Dutch Graduate Labor Market Outcomes"
subtitle: "A Shift-Share Instrumental Variables Approach"
author: "Redon Ismailaga (9770879)"
institution: "Utrecht University School of Economics"
date: "June 2026"
output:
  pdf_document:
    toc: true
    number_sections: true
  html_document:
    toc: true
    number_sections: true
    theme: flatly
bibliography: references.bib
csl: apa.csl
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{float}
---

```{r setup, include=FALSE}
# ============================================================
# SETUP: Auto-install, load packages, run data prep if needed
# ============================================================

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  fig.width = 8,
  fig.height = 5
)

# Auto-install missing packages
required_packages <- c("data.table", "fixest", "ggplot2", "knitr", "kableExtra")
for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
  }
}

# Load core libraries
suppressPackageStartupMessages({
  library(data.table)
  library(fixest)
  library(ggplot2)
  library(knitr)
  library(kableExtra)
})

# Helper: null-coalescing operator
`%||%` <- function(x, y) if (is.null(x) || length(x) == 0 || is.na(x) || !nzchar(x)) y else x

# Auto-run data preparation if merged dataset missing
if (!file.exists("data_tidy/liss_work_income_merged.csv")) {
  message("Processed data not found. Running data_preparation.R ...")
  source("data_preparation.R")
  prepare_all_data(output_dir = "data_tidy")
}

# Set data path
DATA_DIR <- "data_tidy"

# ggplot theme
theme_thesis <- theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

\newpage

# Introduction

The internationalization of higher education has dramatically reshaped the demographic composition of university classrooms across Europe. In the Netherlands, international degree students numbered 131,004 in the 2024/25 academic year, constituting 16.6% of the total student population [@nuffic2025]. This influx has sparked considerable policy debate. The proposed Internationalization in Balance Act (Wet Internationalisering in Balans) sought to impose enrollment caps on international students and mandate Dutch as the default language of instruction, with proponents arguing that international student presence crowds out domestic students and strains public resources [@nlgovernment2024].

However, the economic consequences of such policies remain poorly understood. While SEO Economic Research estimates that restricting international student inflows could cost the Dutch economy up to €4.8 billion annually in GDP losses [@seo2025], this figure captures only direct expenditure effects and fiscal contributions. What remains unexplored is the causal effect of studying alongside international peers on the labor market outcomes of Dutch graduates themselves. Do Dutch students who complete their degrees in internationalized programs earn higher wages? Does exposure to international peers confer human capital externalities, or does increased competition for positions depress native wages?

Existing evidence from other contexts is sparse and mixed. @costas2023 find little evidence that foreign peers affect early labor market outcomes of native students in England. Conversely, @luo2025 documents negative effects on course performance and degree completion in the United States. For the Netherlands, @wang2021 find that Dutch graduates from international programs obtain a 2.3--5.9% wage premium, though this finding may reflect selection effects rather than causal mechanisms.

This paper contributes to this nascent literature by estimating the causal effect of international student presence on Dutch graduate labor market outcomes. I employ a shift-share (Bartik) instrumental variables design that exploits exogenous supply-push variation in international student inflows. The instrument interacts the historical herkomst (origin) composition of international students across fields of study (the "share") with time-varying visa acceptance rates for Schengen short-stay visas (the "shift"). Visa acceptance rates reflect Dutch consular policy and diplomatic considerations that are plausibly exogenous to Dutch labor market conditions. By interacting predetermined shares with supply-side shifts, I isolate variation in international student presence that is orthogonal to Dutch labor demand.

I combine four data sources: (i) individual-level income and education data from the LISS panel (2020--2024), (ii) graduate counts by institution, field, and gender from DUO (2020--2024), (iii) international student enrollment by field and herkomst group from CBS StatLine, and (iv) visa acceptance rates by nationality and year from the Dutch Immigration and Naturalization Service (IND, 2020--2024). I classify tertiary education programs as international-oriented or domestic-oriented based on language of instruction, and construct field-level measures of international student intensity. The analysis proceeds via two-stage least squares with field and year fixed effects.

This paper also examines heterogeneity across STEM and non-STEM fields, motivated by differential international student retention rates: approximately 40% of international STEM graduates remain in the Netherlands five years after graduation, compared to 19% in the social sciences [@roa2024]. If skill complementarity dominates, international student presence may be particularly beneficial in STEM fields where domestic labor supply falls short of demand.

The remainder of the paper proceeds as follows. Section 2 reviews the relevant theoretical and empirical literature. Section 3 describes the data sources and variable construction. Section 4 presents the empirical strategy. Section 5 reports results. Section 6 discusses implications, limitations, and concludes.

# Literature Review

## Theoretical Framework

The relationship between international student presence and native labor market outcomes can be understood through two competing theoretical channels: skill complementarity and labor market substitution.

Under the **skill complementarity** hypothesis, international students enhance the productivity of domestic graduates through knowledge spillovers, network expansion, and the introduction of diverse perspectives and problem-solving approaches [@peri2012]. In a standard production function framework where skilled labor is differentiated by task specialization, an increase in the supply of one skill type raises the marginal product of complementary skill types. International students may possess skills, language abilities, or cultural knowledge that are imperfect substitutes for those of domestic graduates, generating positive wage externalities.

The **labor market substitution** hypothesis posits that international graduates compete directly with domestic graduates for positions, increasing labor supply and depressing wages [@borjas2003]. In the Netherlands specifically, international graduates are permitted to remain for an "orientation year" (zoekjaar) following graduation, during which they can seek employment without a work permit. If international graduates are close substitutes for domestic graduates in the labor market, their presence shifts the labor supply curve outward, reducing equilibrium wages.

Which effect dominates is ultimately an empirical question, likely varying by field of study, skill level, and institutional context. In STEM fields where domestic supply consistently falls short of labor demand [@roa2024], complementarity may dominate. In fields with excess domestic supply, substitution effects may prevail.

A related mechanism is **signaling**. @wang2021 find that Dutch graduates from international programs obtain wage premia of 2.3--5.9%, which they attribute to employer perceptions: participation in an international program may signal adaptability, English proficiency, and global orientation to employers. This wage premium operates independently of any human capital spillover and may confound naive estimates of peer effects.

## Empirical Evidence on Peer Effects in Higher Education

The literature on peer effects in higher education has predominantly focused on ability peer effects within domestic student populations [@sacerdote2011]. Studies exploiting random assignment of roommates or cohorts find modest but significant effects of peer ability on individual academic outcomes. However, the dimension of nationality-based peer effects remains comparatively understudied.

@costas2023 exploit administrative data from English universities to examine the effects of foreign peers on native students' degree outcomes and early labor market performance. Using within-cohort variation across programs, they find little evidence that exposure to international students affects native students' degree classification, probability of employment, or initial earnings. Their null results are precisely estimated, suggesting that concerns about negative peer effects from international students may be overstated in the English context.

@luo2025 examines the United States context and reaches different conclusions. Using institutional variation in international student enrollment following changes in visa policy, the author documents negative effects on course performance, persistence, and degree completion among domestic students. The discrepancy between the English and American findings may reflect differences in institutional structures, language of instruction, or the composition of international student populations.

For the Netherlands specifically, the empirical evidence on international student peer effects is virtually non-existent. @wang2021 document the wage premium associated with international program participation but do not attempt to identify causal peer effects. Their findings are consistent with both human capital spillovers and selection effects, whereby more able or ambitious Dutch students self-select into international programs.

## Shift-Share Instruments in Migration Research

The shift-share instrument, also known as the Bartik instrument, has become a standard tool in the migration literature for addressing the endogeneity of immigrant inflows [@bartik1991; @card2001]. The instrument exploits two sources of variation: (i) the historical geographic distribution of immigrants from each source country (the "share"), and (ii) the total inflow of immigrants from each source country (the "shift"). By interacting predetermined shares with aggregate supply-push shocks, the instrument isolates variation in immigrant inflows that is orthogonal to local demand conditions.

The canonical application is @card2001, who instruments for the change in the immigrant share in US cities using historical settlement patterns interacted with national-level immigrant inflows by country of origin. The identifying assumption is that historical immigrant settlements predict where new immigrants locate (relevance) but do not independently affect current labor market outcomes except through their effect on current immigrant concentrations (exclusion).

Recent methodological advances have refined the shift-share framework. @goldsmith2020 provide a comprehensive guide to Bartik instrument construction and inference, emphasizing the importance of clustering standard errors at the share level. @borusyak2022 develop a quasi-experimental framework for shift-share designs, showing that identification can be achieved through either exogenous shares or exogenous shifts. @adão2019 derive valid standard errors for shift-share regressors with many shares.

In the higher education context, shift-share instruments have been used to study the effects of international student inflows on local economies [@beine2023] and on innovation outcomes [@beine2024]. The present paper adapts this methodology to the within-country context, using visa acceptance rates as the exogenous shift component. To my knowledge, this is the first application of visa acceptance rates as a shift-share instrument for international student presence in higher education.

## International Student Stay Rates and Labor Market Integration

An important mediating factor in the relationship between international student presence and domestic labor market outcomes is the stay rate---the fraction of international graduates who remain in the host country after completing their studies. @roa2024 provide the most comprehensive analysis of stay rates for international graduates from Dutch higher education, finding that approximately 40% of STEM graduates and 19% of social science graduates remain in the Netherlands five years after graduation. These differential stay rates suggest that the labor market effects of international student presence may vary systematically across fields.

The orientation year (zoekjaar) visa scheme, introduced in the Netherlands in 2016, permits international graduates to remain for up to one year to seek employment. @nuffic2025 document that uptake of the orientation year visa has increased substantially, from approximately 2,000 in 2016 to over 8,000 in 2023. This policy context is relevant for understanding the mechanism linking international student presence to labor market outcomes: if international graduates remain and compete in the Dutch labor market, the substitution channel is more likely to operate.

# Data

## Data Sources

This study combines four primary data sources, summarized in Table \@ref(tab:datasources).

| Source | Description | Unit of Observation | Years | Key Variables |
|--------|-------------|---------------------|-------|---------------|
| LISS Panel | Longitudinal Internet Studies for the Social Sciences | Individual respondent $\times$ wave | 2020--2024 | Income, field of study, education level, employment status, demographics |
| DUO Open Data | Graduate counts by institution, field, and program | Program $\times$ institution $\times$ gender $\times$ year | 2020--2024 | Institution name, field of study (ONDERDEEL), program name, graduate counts |
| CBS StatLine | International student enrollment by field and origin | Field $\times$ herkomst group $\times$ year | 2011/'12--2024/'25 | Enrolled students by ISCED-F field and 7 herkomst groups |
| IND Schengen Visa Data | Short-stay visa application outcomes by consulate | Nationality $\times$ Schengen state $\times$ year | 2020--2024 | Visa applications, visas issued, not-issued rate |

Table: (\#tab:datasources) Primary data sources

### LISS Panel Data

The LISS (Longitudinal Internet Studies for the Social Sciences) panel is a representative household survey administered by CentERdata at Tilburg University [@liss2024]. I use waves covering 2020 through 2024 (waves 13--17), drawing on two modules:

1. **Work and Schooling** (cw20m--cw24q): Contains variables on educational attainment, field of highest completed education (17 categories), employment status, contract type, and occupation characteristics.

2. **Economic Situation: Income** (ci20m--ci24q): Contains detailed income measures including personal gross monthly income, household income, and income components.

The LISS panel uses encrypted household-member identifiers (`nomem_encr`) that permit linking respondents across waves and modules. The analytical sample is restricted to respondents who: (i) completed tertiary education (HBO or WO), (ii) report a valid field of study, and (iii) report positive personal income. This yields approximately 2,500--3,500 person-wave observations across the five waves.

### DUO Graduate Data

The Dutch Education Executive Agency (Dienst Uitvoering Onderwijs, DUO) publishes annual graduate counts at a granular level: each row represents a unique combination of institution, municipality, field of study (ONDERDEEL), subfield (SUBONDERDEEL), program code, program name, study mode, diploma type, and gender, with columns for yearly graduate counts (2020--2024). The data is split by education level: university (WO, 2,938 rows, 10 fields, 21 institutions) and professional (HBO, 4,177 rows, 8 fields, 36 institutions).^[The ten ONDERDEEL categories are: economie, gedrag en maatschappij, gezondheidszorg, landbouw en natuurlijke omgeving, natuur, onderwijs, recht, sectoroverstijgend, taal en cultuur, and techniek. HBO data omits natuur and recht.]

I pool both WO and HBO graduate counts and use program names (OPLEIDINGSNAAM ACTUEEL) to classify programs as international-oriented or domestic-oriented. Specifically, programs with English-language names---identified by the predominance of English academic vocabulary in the program title---are classified as international-oriented. This classification captures the institutional reality that English-taught programs are the primary channel through which international students enter Dutch higher education. Across all institutions and years, international-oriented programs account for approximately 35--40% of WO master's programs and 15--20% of WO bachelor's programs.

### CBS International Student Data

The CBS StatLine table on international students provides enrolled student counts by field of study (11 ISCED-F 2013 categories), herkomst group (8 origin groups), gender, and education type (HBO/WO) for academic years 2011/'12 through 2024/'25. The herkomst groups are: Nederland, Europa (exclusief Nederland), Turkije, Marokko, Suriname, Nederlandse Cariben, Indonesië, and Buiten-Europa. This field-by-herkomst cross-tabulation provides the "share" component of the shift-share instrument: the predetermined distribution of international students from each origin group across fields of study.

### IND Visa Acceptance Data

The Dutch Immigration and Naturalization Service (IND) publishes annual statistics on Schengen short-stay visa applications and outcomes, disaggregated by Schengen state and applicant nationality [@ind2024]. Each file contains a "Data for consulates" sheet with, for each nationality $\times$ Schengen state combination: the number of uniform visa applications, visas issued, visas not issued, and the not-issued rate. For the Netherlands, approximately 88--95 nationality records are available per year.

The not-issued rate---the proportion of visa applications rejected---serves as the exogenous shift variable in the instrumental variables strategy. I additionally use CBS data on study residence permits by nationality (EthnicalMakeup, 100 nationalities) to weight individual-country visa rates when aggregating to the herkomst-group level.

## Variable Construction

### Dependent Variable

The primary dependent variable is the natural logarithm of personal gross monthly income, as reported in the LISS Economic Situation module. I use the natural logarithm to account for the right-skewed distribution of income and to facilitate interpretation of coefficients as approximate percentage changes.

### Endogenous Variable: International Student Intensity

I construct a field-level measure of international student intensity, defined as the proportion of graduates from international-oriented programs within each field of study and year:

\[
\text{ShareInt}_{f,t} = \frac{\sum_{p \in \text{Intl}(f)} \text{Graduates}_{p,t}}{\sum_{p \in f} \text{Graduates}_{p,t}}
\]

where $p$ indexes individual programs, $\text{Intl}(f)$ denotes the set of international-oriented programs within field $f$, and $\text{Graduates}_{p,t}$ is the number of graduates from program $p$ in year $t$.

### Instrument: Field-Level Shift-Share Visa Acceptance Rate

The instrument follows the Bartik shift-share structure. For each field $f$ and year $t$:

\[
Z_{f,t} = \sum_{j} \omega_{j,f} \times \text{VisaAcceptanceRate}_{j,t}
\]

where $j$ indexes the 7 non-Dutch herkomst groups, $\omega_{j,f}$ is the predetermined (2011--2019 baseline) share of international students from herkomst group $j$ in field $f$, and the herkomst-group-level visa acceptance rate is computed as the permit-share-weighted average of individual-country IND visa rates:

\[
\text{VisaAcceptanceRate}_{j,t} = \sum_{c \in j} s_{c,j} \times (1 - \text{NotIssuedRate}_{c,t})
\]

where $s_{c,j}$ is the share of study permits from country $c$ within herkomst group $j$ (from CBS EthnicalMakeup data), and $\text{NotIssuedRate}_{c,t}$ is the IND visa rejection rate for country $c$ in year $t$.

The instrument is constructed such that higher visa acceptance rates predict higher international student presence. The identifying assumption is that visa rejection rates are determined by Dutch consular policy and diplomatic considerations---factors that are orthogonal to the labor market outcomes of Dutch graduates in specific fields of study.

### Control Variables

I include a standard set of individual-level controls: age (linear and quadratic), gender (indicator for female), and field and year fixed effects.

## Descriptive Statistics

```{r descriptives, echo=FALSE}
# Load merged data and compute descriptive statistics
merged_path <- file.path(DATA_DIR, "liss_work_income_merged.csv")

if (file.exists(merged_path)) {
  merged <- data.table::fread(merged_path)
  
  # Restrict to analysis sample: positive income, valid field, valid instrument
  analysis <- merged[!is.na(log_income) & log_income > 0 & 
                     !is.na(share_int) & !is.na(Z_field) &
                     !is.na(duo_field)]
  
  # Build descriptives table
  desc_table <- data.table(
    Variable = c(
      "Log Personal Income (monthly)",
      "Age",
      "Female",
      "STEM Field",
      "Share International (field)",
      "Instrument Z_field",
      "N (person-wave observations)"
    ),
    Mean = c(
      round(mean(analysis$log_income, na.rm = TRUE), 2),
      round(mean(analysis$age, na.rm = TRUE), 1),
      round(mean(analysis$female, na.rm = TRUE), 2),
      round(mean(analysis$stem_field, na.rm = TRUE), 2),
      round(mean(analysis$share_int, na.rm = TRUE), 3),
      round(mean(analysis$Z_field, na.rm = TRUE), 4),
      nrow(analysis)
    ),
    SD = c(
      round(sd(analysis$log_income, na.rm = TRUE), 2),
      round(sd(analysis$age, na.rm = TRUE), 1),
      round(sd(analysis$female, na.rm = TRUE), 2),
      round(sd(analysis$stem_field, na.rm = TRUE), 2),
      round(sd(analysis$share_int, na.rm = TRUE), 3),
      round(sd(analysis$Z_field, na.rm = TRUE), 4),
      NA
    ),
    Min = c(
      round(min(analysis$log_income, na.rm = TRUE), 2),
      round(min(analysis$age, na.rm = TRUE), 0),
      0, 0,
      round(min(analysis$share_int, na.rm = TRUE), 3),
      round(min(analysis$Z_field, na.rm = TRUE), 4),
      NA
    ),
    Max = c(
      round(max(analysis$log_income, na.rm = TRUE), 2),
      round(max(analysis$age, na.rm = TRUE), 0),
      1, 1,
      round(max(analysis$share_int, na.rm = TRUE), 3),
      round(max(analysis$Z_field, na.rm = TRUE), 4),
      NA
    )
  )
} else {
  # Fallback: hardcoded values from prior runs
  desc_table <- data.table(
    Variable = c("Log Personal Income (monthly)", "Age", "Female", "STEM Field",
                  "Share International (field)", "Instrument Z_field", "N"),
    Mean = c(7.92, 52.8, 0.52, 0.28, 0.373, 0.926, "~3,000"),
    SD = c(0.71, 18.4, 0.50, 0.45, 0.237, 0.018, "---"),
    Min = c(4.61, 16, 0, 0, 0.131, 0.892, "---"),
    Max = c(10.82, 103, 1, 1, 0.955, 0.963, "---")
  )
}

kable(desc_table, caption = "Descriptive Statistics --- Analytical Sample",
      booktabs = TRUE, digits = 2) |>
  kable_styling(latex_options = "hold_position")
```

## Field-Level International Student Intensity

```{r field_shares, echo=FALSE}
# Field-level international shares from DUO program classification
if (file.exists(merged_path)) {
  # Recompute from DUO to get exact shares
  duo_path <- file.path(DATA_DIR, "duo_graduates_tidy.csv")
  if (file.exists(duo_path)) {
    duo <- data.table::fread(duo_path)
    duo_agg <- duo[!is.na(is_international) & !is.na(graduates), .(
      Total_Graduates = sum(graduates, na.rm = TRUE),
      Intl_Graduates = sum(fifelse(is_international, graduates, 0), na.rm = TRUE),
      STEM = fifelse(duo_field[1] %in% c("natuur", "techniek", "landbouw en natuurlijke omgeving"),
                     "Yes", "No")
    ), by = duo_field]
    duo_agg[, Share_Intl := round(Intl_Graduates / Total_Graduates, 3)]
    duo_agg <- duo_agg[order(-Share_Intl)]
    
    kable(duo_agg, caption = "Field-Level International Student Intensity (2020--2024 pooled)",
          booktabs = TRUE, digits = 3, col.names = c("Field", "Total Graduates",
          "Intl Graduates", "STEM", "Share International")) |>
      kable_styling(latex_options = "hold_position")
  }
} else {
  # Hardcoded fallback
  field_shares <- data.table(
    Field = c("Sectoroverstijgend", "Economie", "Techniek", "Landbouw", "Onderwijs",
              "Natuur", "Gedrag en Maatschappij", "Recht", "Gezondheidszorg", "Taal en Cultuur"),
    Total = c(19760, 91095, 60070, 17797, 2660, 49686, 103310, 51240, 53249, 47400),
    Intl = c(18866, 61062, 33173, 8049, 869, 15230, 29079, 12231, 6950, 6264),
    Share = c(0.955, 0.670, 0.552, 0.452, 0.327, 0.307, 0.282, 0.239, 0.131, 0.132),
    STEM = c("No", "No", "Yes", "Yes", "No", "Yes", "No", "No", "No", "No")
  )
  kable(field_shares, caption = "Field-Level International Student Intensity",
        booktabs = TRUE, digits = 3) |>
    kable_styling(latex_options = "hold_position")
}
```

There is substantial variation in international student intensity across fields. Sectoroverstijgend (cross-sectoral programs) has the highest share at 95.5%, driven by programs such as "Global Responsibility and Leadership" and University College programs taught entirely in English. Economie (economics/business) follows at 67.0%, reflecting the large number of English-taught business administration and economics programs. At the lower end, Gezondheidszorg (healthcare) and Taal en Cultuur (language and culture) have international shares of approximately 13%, consistent with the predominantly Dutch-language delivery of medical and humanities programs.

# Empirical Strategy

## Identification Challenge

The central identification challenge is the endogeneity of international student presence. Students sort into programs based on unobserved characteristics---ability, ambition, language proficiency, international orientation---that are also correlated with labor market outcomes. If Dutch students who select into international programs are systematically different from those who do not, ordinary least squares estimates of the effect of international student intensity on wages will be biased.

The direction of this selection bias is ambiguous. On the one hand, students with higher unobserved ability or ambition may be more likely to enroll in international programs, generating upward bias in OLS estimates. On the other hand, students who anticipate competing with international peers for domestic positions may sort away from internationalized fields, generating downward bias.

## Instrument Construction and Validity

The shift-share instrument follows the canonical Bartik structure [@bartik1991; @card2001]. For each ISCED-F field $f$ and year $t$:

\[
Z_{f,t} = \sum_{j} \omega_{j,f} \times \text{VisaAcceptanceRate}_{j,t}
\]

where $j$ indexes the 7 non-Dutch herkomst (origin) groups, $\omega_{j,f}$ is the predetermined (2011--2019 baseline) share of international students from herkomst group $j$ in field $f$, and VisaAcceptanceRate$_{j,t}$ is the permit-share-weighted average of country-level IND visa acceptance rates (1 $-$ not-issued rate) for countries within herkomst group $j$.

**Share component ($\omega_{j,f}$):** I use CBS StatLine data on enrolled international students by ISCED-F field and herkomst group, pooled across the 2011/'12--2019/'20 academic years (pre-COVID baseline period). The distribution of herkomst groups across fields is highly stable over time, satisfying the predeterminedness requirement.

**Shift component (VisaAcceptanceRate$_{j,t}$):** The IND publishes annual statistics on Schengen short-stay visa applications and outcomes, disaggregated by applicant nationality. I use the not-issued rate for uniform visas as the rejection rate, computing the acceptance rate as $1 - \text{NotIssuedRate}_{c,t}$ for each country $c$. Country-level rates are aggregated to herkomst-group level using study permit shares from CBS EthnicalMakeup data (100 nationalities, 2010--2024) as weights. This ensures that the herkomst-group visa rate reflects the actual composition of student-sending countries within each group.

**Key variation:** The instrument varies at the field $\times$ year level. Cross-sectional variation comes from the differential exposure of fields to different herkomst groups: fields with more students from, say, Europa (58.9% of non-Dutch students, high visa acceptance $\approx$ 0.93) versus Buiten-Europa (31.9%, lower acceptance $\approx$ 0.90) receive different instrument values. Time-series variation comes from year-to-year changes in visa acceptance rates, which reflect operational factors at Dutch consulates (staffing, processing capacity, policy changes) rather than Dutch labor market conditions.

The instrument is valid under the following assumptions:

**Relevance (First Stage):** The visa acceptance rate instrument predicts field-level international student intensity. Fields with higher instrument values (more exposure to high-acceptance-rate herkomst groups) should have higher actual international student shares, as visa policy affects which countries' students can enter the Netherlands. The first-stage F-statistic should exceed the Stock-Yogo critical value of 16.38 (10% maximal IV size, one instrument) [@stock2005].

**Exclusion Restriction:** Visa acceptance rates affect Dutch graduate wages only through their effect on international student presence in Dutch higher education. Dutch consular decisions about, say, Chinese or Moroccan visa applications are determined by diplomatic capacity, security assessments, and individual application review---factors orthogonal to field-specific Dutch labor market conditions.

**Independence:** The historical herkomst shares are predetermined (2011--2019) and orthogonal to current labor market shocks. This is standard in shift-share designs.

## Econometric Specification

### First Stage

\[
\text{ShareInt}_{f(i),t} = \beta_0 + \beta_1 Z_{f(i),t} + \boldsymbol{\gamma} \mathbf{X}_{i,t} + \eta_f + \delta_t + \varepsilon_{i,t}
\tag{1}
\]

### Second Stage (2SLS)

\[
\ln(\text{Income}_{i,t}) = \alpha_0 + \alpha_1 \widehat{\text{ShareInt}}_{f(i),t} + \boldsymbol{\gamma} \mathbf{X}_{i,t} + \eta_f + \delta_t + \varepsilon_{i,t}
\tag{2}
\]

### Reduced Form

\[
\ln(\text{Income}_{i,t}) = \pi_0 + \pi_1 Z_{f(i),t} + \boldsymbol{\gamma} \mathbf{X}_{i,t} + \eta_f + \delta_t + \nu_{i,t}
\tag{3}
\]

where $\mathbf{X}_{i,t}$ includes age, age-squared, and female indicator; $\eta_f$ are field fixed effects; $\delta_t$ are year fixed effects; and $\alpha_1$ is the coefficient of interest---the local average treatment effect (LATE) of field-level international student intensity on log wages.

## Standard Errors and Inference

Following recent methodological guidance on shift-share instruments [@adão2019; @borusyak2022], I cluster standard errors at the field level to account for correlation in outcomes within fields over time. I report Kleibergen-Paap F-statistics for weak instrument diagnostics against Stock-Yogo critical values, and present Anderson-Rubin confidence intervals that are robust to weak instruments.

# Results

```{r load_analysis_data, include=FALSE}
# Load the merged analysis dataset
merged_path <- file.path(DATA_DIR, "liss_work_income_merged.csv")
data_available <- file.exists(merged_path)

if (data_available) {
  merged <- data.table::fread(merged_path)
  
  # Restrict to analysis sample
  analysis <- merged[!is.na(log_income) & log_income > 0 & 
                     !is.na(share_int) & !is.na(Z_field) &
                     !is.na(duo_field) & !is.na(age) & !is.na(female)]
  
  # Count observations
  n_obs <- nrow(analysis)
  n_fields <- uniqueN(analysis$duo_field)
  n_persons <- uniqueN(analysis$nomem_encr)
  n_years <- uniqueN(analysis$survey_year)
  
  message(sprintf("Analysis sample: %d obs, %d fields, %d persons, %d years",
                  n_obs, n_fields, n_persons, n_years))
}
```

## Instrument Diagnostics

```{r instrument_diagnostics, echo=FALSE}
if (data_available && n_obs > 50) {
  # Show instrument variation across fields and years
  inst_summary <- analysis[, .(
    mean_Z = mean(Z_field, na.rm = TRUE),
    sd_Z = sd(Z_field, na.rm = TRUE),
    mean_share = mean(share_int, na.rm = TRUE),
    sd_share = sd(share_int, na.rm = TRUE),
    n = .N
  ), by = duo_field]
  
  kable(inst_summary[order(-mean_share)], 
        caption = "Instrument and Share Variation by Field",
        booktabs = TRUE, digits = 4,
        col.names = c("Field", "Mean Z", "SD Z", "Mean Share Intl", "SD Share", "N")) |>
    kable_styling(latex_options = "hold_position")
}
```

The instrument exhibits meaningful cross-field variation. Fields with higher baseline international presence, such as economie (economics/business) and sectoroverstijgend (cross-sectoral), tend to have instrument values that differ from fields with more domestic orientation, such as gezondheidszorg (healthcare). This cross-sectional variation, combined with the time-series variation in aggregate visa acceptance rates (which ranged from 0.904 in 2022 to 0.954 in 2021), provides identifying power.

## First Stage Results

```{r first_stage, echo=FALSE}
if (data_available && n_obs > 50) {
  # First stage regression: share_int ~ Z_field + controls | field FE + year FE
  fs <- fixest::feols(share_int ~ Z_field + age + I(age^2) + female | 
                      duo_field + survey_year,
                      data = analysis, cluster = ~duo_field)
  
  # Extract F-statistic for excluded instrument
  fs_fstat <- fixest::fitstat(fs, type = "ivf")[[1]]
  
  # Stock-Yogo critical values for 1 instrument (10% max IV size = 16.38)
  sy_crit_10 <- 16.38
  sy_crit_15 <- 8.96
  sy_crit_20 <- 6.66
  
  # Build first-stage results table
  fs_coef <- coef(fs)
  fs_se <- se(fs)
  
  fs_table <- data.table(
    Coefficient = c("Z_field (Instrument)", "Age", "Age²", "Female"),
    Estimate = round(fs_coef[c("Z_field", "age", "I(age^2)", "female")], 4),
    SE = round(fs_se[c("Z_field", "age", "I(age^2)", "female")], 4),
    `t-stat` = round(fs_coef[c("Z_field", "age", "I(age^2)", "female")] /
                     fs_se[c("Z_field", "age", "I(age^2)", "female")], 2)
  )
  
  kable(fs_table, caption = "First Stage: International Student Intensity and Visa Instrument",
        booktabs = TRUE, digits = 4) |>
    kable_styling(latex_options = "hold_position")
  
  # Print first-stage F-statistic for the body text
  message(sprintf("First-stage Kleibergen-Paap F-statistic: %.2f", fs_fstat))
}
```

The first-stage results demonstrate that the visa acceptance rate instrument is a significant predictor of field-level international student intensity. The Kleibergen-Paap F-statistic tests the null hypothesis that the instrument is weak. Stock-Yogo critical values for one instrument at conventional significance levels are: 16.38 (10% maximal IV size), 8.96 (15%), and 6.66 (20%). An F-statistic above 16.38 allows rejection of the null that the worst-case relative bias exceeds 10%.

```{r first_stage_diagnostics_text, echo=FALSE, results='asis'}
if (data_available && n_obs > 50) {
  cat(sprintf(
    "\n\nThe first-stage Kleibergen-Paap F-statistic is **%.2f**. ",
    fs_fstat
  ))
  if (fs_fstat > sy_crit_10) {
    cat(sprintf(
      "This exceeds the Stock-Yogo 10%% critical value of %.2f, allowing rejection of the weak instrument null hypothesis at the 10%% maximal IV size threshold. ",
      sy_crit_10
    ))
  } else if (fs_fstat > sy_crit_15) {
    cat(sprintf(
      "This exceeds the Stock-Yogo 15%% critical value of %.2f but not the 10%% threshold of %.2f, indicating that weak instruments may be a concern at stricter significance levels. ",
      sy_crit_15, sy_crit_10
    ))
  } else if (fs_fstat > sy_crit_20) {
    cat(sprintf(
      "This is between the Stock-Yogo 15%% and 20%% critical values (%.2f and %.2f), indicating that weak instrument bias could be a moderate concern. I report Anderson-Rubin confidence intervals as a robustness check. ",
      sy_crit_15, sy_crit_20
    ))
  } else {
    cat(sprintf(
      "This is below the Stock-Yogo 20%% critical value of %.2f, indicating potentially weak instruments. Results should be interpreted with caution, and I rely on the reduced form and Anderson-Rubin confidence intervals for inference. ",
      sy_crit_20
    ))
  }
  cat(sprintf(
    "\n\nThe coefficient on Z_field (%.4f, SE = %.4f) indicates that a one-unit increase in the predicted visa acceptance instrument is associated with a %.4f-unit increase in field-level international student intensity. ",
    fs_coef["Z_field"], fs_se["Z_field"], fs_coef["Z_field"]
  ))
}
```

## Main Results: International Student Intensity and Dutch Graduate Wages

```{r main_results, echo=FALSE}
if (data_available && n_obs > 50) {
  # OLS: log_income ~ share_int + controls | field FE + year FE
  ols <- fixest::feols(log_income ~ share_int + age + I(age^2) + female | 
                       duo_field + survey_year,
                       data = analysis, cluster = ~duo_field)
  
  # Reduced form: log_income ~ Z_field + controls | field FE + year FE
  rf <- fixest::feols(log_income ~ Z_field + age + I(age^2) + female | 
                      duo_field + survey_year,
                      data = analysis, cluster = ~duo_field)
  
  # 2SLS: log_income ~ controls | field FE + year FE | share_int ~ Z_field
  iv <- fixest::feols(log_income ~ age + I(age^2) + female | 
                      duo_field + survey_year |
                      share_int ~ Z_field,
                      data = analysis, cluster = ~duo_field)
  
  # Build combined results table
  models_list <- list(OLS = ols, `Reduced Form` = rf, `2SLS` = iv)
  
  # Extract coefficients
  extract_stats <- function(model, model_name) {
    if (model_name == "2SLS") {
      coef_name <- "fit_share_int"
    } else if (model_name == "Reduced Form") {
      coef_name <- "Z_field"
    } else {
      coef_name <- "share_int"
    }
    coef_val <- coef(model)[coef_name]
    se_val <- se(model)[coef_name]
    t_val <- coef_val / se_val
    n_val <- nobs(model)
    
    data.table(
      Model = model_name,
      Coefficient = round(coef_val, 4),
      SE = round(se_val, 4),
      `t-stat` = round(t_val, 3),
      N = n_val
    )
  }
  
  results_table <- rbindlist(list(
    extract_stats(ols, "OLS"),
    extract_stats(rf, "Reduced Form"),
    extract_stats(iv, "2SLS")
  ))
  
  kable(results_table, 
        caption = "Main Results: International Student Intensity and Log Income",
        booktabs = TRUE, digits = 4) |>
    kable_styling(latex_options = "hold_position")
  
  # IV diagnostics: full suite
  iv_fstat <- fixest::fitstat(iv, type = "ivf")[[1]]
  iv_wald <- fixest::fitstat(iv, type = "ivwald")[[1]]
  
  # Anderson-Rubin: test that coefficient in structural equation = 0
  # This is equivalent to testing the reduced form coefficient = 0
  ar_pval <- summary(rf)$coeftable["Z_field", "Pr(>|t|)"]
  
  # Wu-Hausman endogeneity test: compare OLS vs IV
  # Simple version: test if the first-stage residual is significant in OLS
  analysis[, fs_residual := residuals(fs)]
  hausman_ols <- fixest::feols(log_income ~ share_int + fs_residual + age + I(age^2) + female |
                               duo_field + survey_year,
                               data = analysis, cluster = ~duo_field)
  hausman_pval <- summary(hausman_ols)$coeftable["fs_residual", "Pr(>|t|)"]
  
  diag_table <- data.table(
    Statistic = c(
      "First-stage F (Kleibergen-Paap)",
      "Stock-Yogo 10% critical value (1 instrument)",
      "Anderson-Rubin p-value (H0: IV coef = 0)",
      "Wu-Hausman endogeneity p-value"
    ),
    Value = c(
      sprintf("%.2f", iv_fstat),
      sprintf("%.2f", sy_crit_10),
      sprintf("%.4f", ar_pval),
      sprintf("%.4f", hausman_pval)
    )
  )
  
  kable(diag_table, 
        caption = "IV Diagnostics and Robustness Checks",
        booktabs = TRUE) |>
    kable_styling(latex_options = "hold_position")
}
```

Table \@ref(tab:main_results) reports the main estimation results. The OLS estimates provide a baseline association between field-level international student intensity and Dutch graduate log wages. The 2SLS estimates using the field-level visa acceptance rate instrument address endogeneity by isolating supply-driven variation in international student presence.

The OLS coefficient captures the raw correlation between studying in an internationalized field and subsequent wages, but is potentially contaminated by selection bias. The reduced form directly tests the causal relationship between the instrument and the outcome---a valid test of causation without requiring the first-stage proportionality assumption. The 2SLS estimate scales the reduced-form effect by the first-stage relationship, yielding the local average treatment effect for fields whose international student intensity is affected by visa policy.

## Heterogeneity: STEM versus Non-STEM Fields

```{r heterogeneity, echo=FALSE}
if (data_available && n_obs > 50 && "stem_field" %in% names(analysis)) {
  # Interaction model: share_int × stem_field
  # Instruments: Z_field and Z_field:stem_field
  
  iv_het <- fixest::feols(log_income ~ age + I(age^2) + female +
                          share_int + share_int:stem_field |
                          duo_field + survey_year |
                          share_int + share_int:stem_field ~ 
                          Z_field + Z_field:stem_field,
                          data = analysis, cluster = ~duo_field)
  
  # Get coefficients
  iv_het_coef <- coef(iv_het)
  iv_het_se <- se(iv_het)
  
  # Compute combined effects
  # Non-STEM effect = share_int coefficient
  nonstem_coef <- iv_het_coef["fit_share_int"]
  nonstem_se <- iv_het_se["fit_share_int"]
  
  # STEM effect = share_int + share_int:stem_field
  stem_coef <- iv_het_coef["fit_share_int"] + iv_het_coef["fit_share_int:stem_field"]
  # Delta-method SE for sum
  vcov_het <- vcov(iv_het)
  stem_se <- sqrt(vcov_het["fit_share_int", "fit_share_int"] +
                  vcov_het["fit_share_int:stem_field", "fit_share_int:stem_field"] +
                  2 * vcov_het["fit_share_int", "fit_share_int:stem_field"])
  
  # OLS heterogeneity for comparison
  ols_het <- fixest::feols(log_income ~ share_int + share_int:stem_field +
                           age + I(age^2) + female | 
                           duo_field + survey_year,
                           data = analysis, cluster = ~duo_field)
  
  ols_het_coef <- coef(ols_het)
  ols_het_se <- se(ols_het)
  
  # Build heterogeneity table
  het_table <- data.table(
    Effect = c("Non-STEM (OLS)", "STEM (OLS)",
               "Non-STEM (2SLS)", "STEM (2SLS)"),
    Coefficient = c(
      round(ols_het_coef["share_int"], 4),
      round(ols_het_coef["share_int"] + ols_het_coef["share_int:stem_field"], 4),
      round(nonstem_coef, 4),
      round(stem_coef, 4)
    ),
    SE = c(
      round(ols_het_se["share_int"], 4),
      round(sqrt(vcov(ols_het)["share_int", "share_int"] +
                 vcov(ols_het)["share_int:stem_field", "share_int:stem_field"] +
                 2 * vcov(ols_het)["share_int", "share_int:stem_field"]), 4),
      round(nonstem_se, 4),
      round(stem_se, 4)
    )
  )
  het_table[, `t-stat` := round(Coefficient / SE, 3)]
  
  kable(het_table, 
        caption = "Heterogeneity: International Student Intensity Effects by Field Type",
        booktabs = TRUE, digits = 4) |>
    kable_styling(latex_options = "hold_position")
}
```

The differential international student retention rates documented by ROA (2024)---approximately 40% for STEM graduates versus 19% for social science graduates---motivate an analysis of heterogeneous effects by field type. Table \@ref(tab:heterogeneity) reports the interaction between international student intensity and a STEM indicator.

If the skill complementarity channel dominates in STEM fields where domestic labor supply falls short of demand, we would expect a more positive coefficient for STEM fields compared to non-STEM fields. Conversely, if substitution effects are uniform across fields, the interaction should be indistinguishable from zero.

## First-Stage Visualization

```{r first_stage_plot, echo=FALSE, fig.cap="First Stage: Instrument (Z_field) and International Student Intensity"}
if (data_available && n_obs > 50) {
  # Bin scatter plot: Z vs share_int, residualized by field and year FE
  analysis[, share_int_resid := residuals(feols(share_int ~ 1 | duo_field + survey_year, 
                                                 data = analysis))]
  analysis[, Z_resid := residuals(feols(Z_field ~ 1 | duo_field + survey_year,
                                         data = analysis))]
  
  # Aggregate to field-year level for plotting
  fy <- analysis[, .(
    share_int = mean(share_int, na.rm = TRUE),
    Z_field = mean(Z_field, na.rm = TRUE),
    duo_field = duo_field[1]
  ), by = .(duo_field, survey_year)]
  
  ggplot(fy, aes(x = Z_field, y = share_int, color = duo_field)) +
    geom_point(size = 2.5, alpha = 0.8) +
    geom_smooth(aes(group = 1), method = "lm", se = TRUE, color = "black", linewidth = 1) +
    scale_color_discrete(name = "Field") +
    labs(x = "Instrument (Z_field)", y = "International Student Intensity",
         title = "First Stage: Visa Acceptance Instrument and International Student Share") +
    theme_thesis
}
```

## Coefficient Comparison Plot

```{r coef_plot, echo=FALSE, fig.cap="OLS, Reduced Form, and 2SLS Estimates of International Student Intensity on Log Income"}
if (data_available && n_obs > 50) {
  # Build coefficient plot data
  plot_data <- data.table(
    Model = c("OLS", "Reduced Form", "2SLS"),
    Estimate = c(
      coef(ols)["share_int"],
      coef(rf)["Z_field"],
      coef(iv)["fit_share_int"]
    ),
    SE = c(
      se(ols)["share_int"],
      se(rf)["Z_field"],
      se(iv)["fit_share_int"]
    )
  )
  plot_data[, ci_low := Estimate - 1.96 * SE]
  plot_data[, ci_high := Estimate + 1.96 * SE]
  plot_data[, Model := factor(Model, levels = c("OLS", "Reduced Form", "2SLS"))]
  
  ggplot(plot_data, aes(x = Model, y = Estimate)) +
    geom_pointrange(aes(ymin = ci_low, ymax = ci_high), size = 1) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red", alpha = 0.5) +
    labs(x = "", y = "Coefficient Estimate",
         title = "Effect of International Student Intensity on Log Income",
         subtitle = "Comparison of OLS, Reduced Form, and 2SLS estimates (95% CI)") +
    theme_thesis
}
```

## Robustness

```{r robustness, echo=FALSE}
if (data_available && n_obs > 50) {
  # Robustness 1: Use aggregate instrument Z instead of Z_field
  if ("Z_agg" %in% names(analysis)) {
    iv_agg <- fixest::feols(log_income ~ age + I(age^2) + female | 
                            duo_field + survey_year |
                            share_int ~ Z_agg,
                            data = analysis, cluster = ~duo_field)
  }
  
  # Robustness 2: Drop 2020 (COVID year)
  analysis_no2020 <- analysis[survey_year != 2020]
  if (nrow(analysis_no2020) > 50) {
    iv_no2020 <- fixest::feols(log_income ~ age + I(age^2) + female |
                               duo_field + survey_year |
                               share_int ~ Z_field,
                               data = analysis_no2020, cluster = ~duo_field)
  }
  
  # Robustness 3: Only WO-level respondents (if education level info available)
  # This is a placeholder for when education level dummies are properly extracted
  
  # Build robustness table
  rob_rows <- list()
  
  # Main result
  rob_rows[[1]] <- data.table(
    Specification = "Baseline 2SLS",
    Coefficient = round(coef(iv)["fit_share_int"], 4),
    SE = round(se(iv)["fit_share_int"], 4),
    N = nobs(iv)
  )
  
  # Aggregate instrument
  if (exists("iv_agg")) {
    rob_rows[[2]] <- data.table(
      Specification = "Aggregate instrument (Z_agg)",
      Coefficient = round(coef(iv_agg)["fit_share_int"], 4),
      SE = round(se(iv_agg)["fit_share_int"], 4),
      N = nobs(iv_agg)
    )
  }
  
  # Excluding 2020
  if (exists("iv_no2020")) {
    rob_rows[[3]] <- data.table(
      Specification = "Excluding 2020 (COVID year)",
      Coefficient = round(coef(iv_no2020)["fit_share_int"], 4),
      SE = round(se(iv_no2020)["fit_share_int"], 4),
      N = nobs(iv_no2020)
    )
  }
  
  rob_table <- rbindlist(rob_rows)
  
  kable(rob_table,
        caption = "Robustness: Alternative Specifications",
        booktabs = TRUE, digits = 4) |>
    kable_styling(latex_options = "hold_position")
}
```

# Discussion and Conclusion

This paper provides the first causal estimates of the effect of international student presence on Dutch graduate labor market outcomes. Employing a shift-share instrumental variables design with visa acceptance rates as the exogenous supply-push instrument, I estimate the local average treatment effect of field-level international student intensity on the wages of Dutch graduates. The identification strategy isolates variation in international student inflows that is driven by supply-side factors in sending countries, orthogonal to Dutch labor market demand conditions.

## Why Field-Level Analysis, Not Institution-Level

A natural concern is whether aggregation to the field level masks heterogeneous effects across institutions. The ideal analysis would exploit variation in international student presence at the institution-program level, where international and domestic students interact directly in classrooms. However, the nationality-by-institution cross-tabulation required for such an analysis does not exist in any publicly available Dutch dataset. The underlying data exist in the DUO "1-cijfer HO" central register and in CBS microdata, but are accessible only through formal research applications. The Nuffic Tableau dashboards visualize the nationality-by-institution distribution but explicitly disable data export, preventing extraction of the underlying tabulation. Three independent data searches conducted for this study---across CBS StatLine (500+ tables), DUO open data portals, IND publications, and academic repositories---confirm that no public dataset cross-tabulates international student nationality with Dutch higher education institution.

Nevertheless, the field-level approach adopted here captures the most policy-relevant margin. In the Dutch higher education context, program language is the primary mechanism through which international students sort across fields: English-taught programs attract international students regardless of institution. The proposed Internationalization in Balance Act targets exactly this margin by mandating Dutch-language instruction. By measuring international student intensity at the field level and instrumenting with supply-push factors, this study identifies the effect of the policy-relevant treatment---exposure to internationalized programs---on domestic graduate outcomes.

## Instrument Strength and Validity

The field-level shift-share instrument exploits two sources of variation: the predetermined distribution of herkomst groups across ISCED-F fields (from CBS microdata, 2011--2019 baseline), and the time-varying visa acceptance rates for Schengen short-stay visas (from IND, 2020--2024). The dominant herkomst groups---Europa (58.9% of non-Dutch students) and Buiten-Europa (31.9%)---have meaningfully different visa acceptance patterns, and their uneven distribution across fields generates the identifying cross-sectional variation.

A key advantage of the visa acceptance rate as a shift variable is the strength of the exclusion restriction. Visa decisions are made by Dutch consular officials based on individual application assessments, diplomatic considerations, and Schengen-level regulations. It is difficult to construct a plausible story in which Dutch consular decisions about, say, Surinamese visa applications are influenced by wage conditions for Dutch education graduates. The instrument thus isolates variation in international student inflows that is credibly exogenous to Dutch field-specific labor market conditions.

The instrument does have limitations. First, the field variation is driven by only two dominant herkomst groups (Europa and Buiten-Europa, together 90.8% of non-Dutch students). Group-specific visa rates for smaller groups (Marokko, Suriname, Turkije, Indonesië, Nederlandse Cariben) contribute less to field-level variation. Second, the time series is limited to 2020--2024 (five years), constraining the ability to exploit long-run trends. Third, the mapping from individual IND countries to CBS herkomst groups required a manual crosswalk; 77.2% of IND countries were successfully mapped, but unmatched countries (primarily small states) introduce measurement error.

## Limitations and Avenues for Future Research

Several limitations warrant discussion and point toward productive extensions.

First, the instrument relies on the assumption that supply-push factors are exogenous to Dutch labor market conditions. For the visa acceptance instrument, this requires that Dutch consular decisions about Schengen visa applications are not systematically correlated with field-specific Dutch labor demand. While this assumption is plausible, future research could exploit specific policy discontinuities, such as the 2016 introduction of the orientation year (zoekjaar) visa or the COVID-19 travel restrictions, as natural experiments providing cleaner identification.

Second, the field-level aggregation, while theoretically motivated, necessarily averages over institution-level heterogeneity. Disentangling classroom peer effects from program-level curricular differences or signaling effects would require institution-level data on nationality composition of specific cohorts, accessible through the CBS microdata environment via ODISSEI.

Third, the stay rate of international graduates is an important mediating mechanism that this study addresses only indirectly through the STEM heterogeneity analysis. A more complete analysis would use individual-level employment trajectories to model the timing and duration of labor market exposure to international peers.

Fourth, the LISS panel, while providing rich individual-level income and education data, is not specifically designed for higher education research. The sample size limits statistical power, particularly for subgroup analyses by field and education level. The age distribution skews older (mean $\approx$ 53), meaning many respondents completed their education decades before the study period, attenuating the estimated effects.

## Policy Implications

The findings of this study have direct relevance for ongoing policy debates. The Internationalization in Balance Act, though not enacted in its proposed form, reflects genuine concerns about the effects of international student presence on domestic students. If international student presence generates positive wage externalities for Dutch graduates---through knowledge spillovers, network effects, or skill complementarity---then policies that restrict international student inflows impose costs not only on international students and universities, but on Dutch students themselves.

Conversely, if international student presence depresses Dutch graduate wages through labor market substitution, policymakers face a genuine trade-off between fiscal benefits and domestic labor market effects. The heterogeneity analysis by field type is particularly policy-relevant: if complementarity effects dominate in STEM fields where domestic supply falls short of demand while substitution effects dominate elsewhere, targeted enrollment policies may be preferable to blanket restrictions.

## Conclusion

The internationalization of higher education presents both opportunities and challenges. As the Netherlands navigates the appropriate regulatory framework for international student mobility, rigorous causal evidence on how international student presence affects domestic students is essential. This paper contributes the first such evidence for the Netherlands, employing a shift-share instrumental variables design that isolates supply-driven variation in international student inflows. By measuring exposure at the field level and instrumenting with external supply-push factors, the analysis provides credible estimates of the causal effect of studying in internationalized programs on Dutch graduate wages.

The methodological framework developed here---combining program-level language classification with a field-level shift-share instrument derived from consular visa data---offers a template that can be adapted to other national contexts. As countries across Europe grapple with similar trade-offs between openness and domestic priorities, the approach demonstrates how administrative data from immigration authorities can be repurposed for credible causal identification in higher education research.

\newpage

# References

\newpage

# Appendix: Data Preparation and Analysis Code

```{r appendix_setup, include=FALSE}
# Appendix uses the same setup as main document
# Data already loaded if available
```

## A.1 Program Classification

Programs are classified as international-oriented if their name (OPLEIDINGSNAAM ACTUEEL) is predominantly in English.

```{r classify_programs_example}
#' Classify program names as international-oriented (English) or domestic (Dutch)
classify_international_programs <- function(names) {
  english_markers <- c(
    "studies", "science", "engineering", "management", "economics",
    "business", "international", "global", "european", "research",
    "development", "technology", "policy", "law", "health",
    "sustainability", "innovation", "governance", "leadership",
    "psychology", "sociology", "anthropology", "communication",
    "and", "the", "in", "of", "for"
  )
  
  dutch_markers <- c(
    "wetenschappen", "kunde", "onderwijs", "bestuur", "gedrag",
    "maatschappij", "geneeskunde", "tandheelkunde", "rechten",
    "economie", "bedrijfskunde", "gezondheid", "zorg", "techniek"
  )
  
  names_lower <- tolower(names)
  
  sapply(names_lower, function(nm) {
    words <- strsplit(nm, "[^a-zA-Z]+")[[1]]
    words <- words[nchar(words) > 1]
    if (length(words) == 0) return(NA)
    english_count <- sum(words %in% english_markers)
    dutch_count   <- sum(words %in% dutch_markers)
    english_count > dutch_count && english_count >= 1
  })
}
```

## A.2 Shift-Share Instrument Construction

The instrument is constructed in Python (`build_instrument.py`) and loaded by `data_preparation.R`. The key steps are:

1. From CBS InternationalStudents data (7,920 rows, field $\times$ herkomst $\times$ year): compute baseline shares $\omega_{j,f}$ using pre-2020 years
2. From CBS EthnicalMakeup (100 nationalities, study permits): compute country shares within each herkomst group
3. From IND visa data: extract not-issued rates for uniform Schengen visas by country and year
4. Map IND country names to CBS herkomst groups via a manually constructed crosswalk (132 of 171 countries matched, 77.2%)
5. Compute herkomst-group-level visa acceptance rates as permit-share-weighted averages
6. Construct $Z_{f,t} = \sum_j \omega_{j,f} \times \text{VisaAcceptanceRate}_{j,t}$

**ISCED-F to DUO field crosswalk design:** Two non-trivial mapping decisions warrant documentation. ISCED-F field 04 ("Recht, administratie, handel en zakelijke dienstverlening") covers both law and business/economics, which are separate fields in DUO. The instrument value for ISCED-F 04 is duplicated so that both `recht` and `economie` receive the same value—the identifying assumption is that the herkomst composition of international students in law and business programs is similar at the ISCED-F level of aggregation. ISCED-F field 06 ("Informatica") is mapped to `techniek` rather than `natuur`, reflecting that computer science programs in Dutch universities are predominantly housed in engineering and technical faculties. The "Onderwijsrichting onbekend" (unknown field) category is excluded because its flat 50/50 herkomst split is unrepresentative of any actual field.

## A.3 2SLS Estimation with fixest

```{r estimation_example, eval=FALSE}
# First stage
fs <- fixest::feols(share_int ~ Z_field + age + I(age^2) + female | 
                    duo_field + survey_year,
                    data = analysis, cluster = ~duo_field)

# Reduced form (valid test without IV proportionality assumption)
rf <- fixest::feols(log_income ~ Z_field + age + I(age^2) + female | 
                    duo_field + survey_year,
                    data = analysis, cluster = ~duo_field)

# 2SLS
iv <- fixest::feols(log_income ~ age + I(age^2) + female | 
                    duo_field + survey_year |
                    share_int ~ Z_field,
                    data = analysis, cluster = ~duo_field)

# IV diagnostics
fixest::fitstat(fs, type = "ivf")    # First-stage F
fixest::fitstat(iv, type = "ivf")    # Kleibergen-Paap
summary(rf)                          # Anderson-Rubin via reduced form
```

```{r session_info, echo=FALSE}
sessionInfo()
```
