---
title: "International Student Presence and Dutch Graduate Labor Market Outcomes"
subtitle: "A Shift-Share Instrumental Variables Approach"
author: "Redon Ismailaga"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    number_sections: true
bibliography: references.bib
csl: apa.csl
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{float}
---

```{r setup, include=FALSE}
# -- Global knitr options ------------------------------------------------------
# echo = FALSE:   don't print the R code itself into the PDF --- only show outputs
# warning = FALSE: suppress R warnings (they'd clutter the PDF with irrelevant noise)
# message = FALSE: suppress package-loading messages (e.g. "Attaching package: dplyr")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r packages}
pacman::p_load(`pacman`)
p_load(tidyverse); p_load(readxl); p_load(janitor); p_load(fixest)
p_load(modelsummary); p_load(kableExtra); p_load(WDI)
p_load(rt); p_load(hunspell)
```

# 1. Data Loading

## 1.1 LISS Panel

```{r liss-load}
# -- LISS Work & Schooling module (2020-2024) ----------------------------------
# The LISS panel is a longitudinal Dutch household survey. The "Work and Schooling"
# module records each respondent's occupation, field of study, education level,
# and working hours. Each yearly file has its own column naming convention.
# Early waves (2020, 2021) use comma-separated format; later waves (2022-2024)
# use semicolon-separated format. check.names = FALSE prevents R from auto-converting
# column names (e.g. spaces to dots).

# 2020 wave: comma-separated (sep = "," is default), 1.0p = version 1.0 public
LISS2020WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw20m_EN_1.0p.csv", check.names = FALSE)
# 2021 wave: also comma-separated
LISS2021WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw21n_EN_1.0p.csv", check.names = FALSE)
# 2022 wave: semicolon-separated (sep = ";")
LISS2022WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw22o_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2023 wave: semicolon-separated
LISS2023WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw23p_EN_2.0p.csv", sep = ";", check.names = FALSE)
# 2024 wave: semicolon-separated
LISS2024WS <- read.csv("../../../Data/LISS Work and Schooling/Extracted/cw24q_EN_1.0p.csv", sep = ";", check.names = FALSE)

# -- LISS Economic Situation / Income module (2020-2024) -----------------------
# The "Economic Situation: Income" module records gross monthly wages, partner status,
# and other income-related variables. Same sep pattern as the Work & Schooling module.

# 2020 income data: comma-separated
LISS2020ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci20m_EN_1.0p.csv", check.names = FALSE)
# 2021 income data: comma-separated
LISS2021ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci21n_EN_1.0p.csv", check.names = FALSE)
# 2022 income data: semicolon-separated
LISS2022ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci22o_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2023 income data: semicolon-separated
LISS2023ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci23p_EN_1.0p.csv", sep = ";", check.names = FALSE)
# 2024 income data: semicolon-separated
LISS2024ES <- read.csv("../../../Data/LISS Economic Situation, Income/Extracted/ci24q_EN_1.0p.csv", sep = ";", check.names = FALSE)
```

## 1.2 IND Visa

```{r visa-load}
# -- IND Schengen visa statistics (2020-2024) ----------------------------------
# The Immigration and Naturalisation Service (IND) publishes yearly Schengen visa
# statistics by consulate/country. Each file is an Excel workbook (.xlsx) with
# a worksheet named "Data for consulates" containing acceptance/rejection rates
# by applicant nationality.
visa_files <- c(
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/Visa-statistics-for-consulates-2020_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2021-consulates-schengen-visa-stats.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/Visa statistics for consulates in 2022_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2023-schengen-visa-statistics-consulates_en.xlsx",
  "../../../Data/IND Short-Term Schengen Visa Acceptance Rates/2024-schengen-visa-statistics-consulates_en.xlsx"
)
```

## 1.4 CBS Nationality Data

```{r cbs-load}
# -- CBS temporary residence permits by nationality (2010) ---------------------
# CBS (Centraal Bureau voor de Statistiek) publishes counts of temporary residence
# permits (verblijfsvergunningen voor bepaalde tijd) by nationality and year.
# This data is used to construct the baseline nationality shares for the shift-share
# instrument. The file name includes a timestamp from when it was downloaded.
EthnicalMakeup <- read.csv("../../../Data/CBS Incoming Student Ethnical Makeup/Verblijfsvergunningen_voor_bepaalde_tijd_19052026_175309.csv", sep = ";")
```

# 2. Process IND Visa

```{r process-visa}
# -- IND Schengen visa acceptance rates (2020-2024) ---------------------------
# Load each year's Excel file, keep only Netherlands consulates, and extract
# acceptance rates by applicant nationality. Column names are selected after
# inspecting the raw datasets.

years <- 2020:2024
visa_list <- list()

for (i in seq_along(visa_files)) {
  if (!file.exists(visa_files[i])) next

  df <- read_excel(visa_files[i], sheet = "Data for consulates") %>%
    clean_names() %>%
    mutate(year = years[i])

  # Filter to Netherlands consulates and select key columns
  # (adjust the exact column names below after viewing the data in R)
  nl <- df %>%
    filter(schengen_state == "Netherlands") %>%
    select(
      country = country_where_consulate_is_located,
      not_issued_rate = not_issued_rate_for_uniform_visas,
      year
    ) %>%
    mutate(
      acceptance_rate = 1 - not_issued_rate,
      country = str_to_upper(str_trim(country)),
      # Turkey changed its English name to "Türkiye" in 2022.
      # The IND files use "TURKEY" (2020-2021) and "TÜRKIYE" (2022-2024).
      # Normalize to "TURKEY" so the inner_join with CBS country_std works.
      country = str_replace(country, "TÜRKIYE", "TURKEY")
    )

  if (nrow(nl) > 0) visa_list[[i]] <- nl
}

Consulates <- bind_rows(visa_list) %>%
  # Average across consulates within each country (e.g. US has DC + NY + Chicago)
  group_by(country, year) %>%
  summarise(acceptance_rate = mean(acceptance_rate, na.rm = TRUE), .groups = "drop")
```

# 3. Field-Level International Share (CBS)
```{r process-cbs-share}
# -- CBS international-background senior share (parents born outside NL) -----
# Uses the subgroups file with granular 3-digit CBS sub-field codes, senior year.
# All CBS sub-fields map to one of the 11 field keys (education through services).
# Field unknown set to NA (drops from merge — no bias from misassignment).

base <- "../../../Data/CBS Incoming Student Ethnical Makeup"

cbs_raw <- read_delim(file.path(base, "Born in or out NL, FoS, gender, subgroups, Senior year.csv"),
                      delim = ";", show_col_types = FALSE) %>%
  clean_names() %>%
  mutate(
    students = as.numeric(str_trim(ouderejaarsstudenten_aantal)),
    year = as.integer(str_sub(perioden, 1, 4))
  )

# Total seniors per field-year: sum across genders and onderwijssoort
total_seniors <- cbs_raw %>%
  filter(!is.na(students), students > 0) %>%
  group_by(studierichting, onderwijssoort, year) %>%
  summarise(total_seniors = sum(students, na.rm = TRUE), .groups = "drop")

# International-background seniors (parents born outside NL)
# Note: this includes Nederlandse Cariben students; they are proportionally
# removed below using the Internationals file's herkomstland breakdown.
intl_seniors <- cbs_raw %>%
  filter(geboorteland_ouders == "Geboren buiten Nederland",
         !is.na(students), students > 0) %>%
  group_by(studierichting, onderwijssoort, year) %>%
  summarise(intl_seniors = sum(students, na.rm = TRUE), .groups = "drop")

cbs_share <- total_seniors %>%
  left_join(intl_seniors, by = c("studierichting", "onderwijssoort", "year")) %>%
  mutate(
    share_int = ifelse(total_seniors > 0,
                       coalesce(intl_seniors, 0) / total_seniors, NA_real_),
    share_int = pmin(share_int, 1)
  )

# Map CBS 3-digit sub-field codes to CBS field groups.
# Skips aggregate codes (z.n.d. = zonder nadere detaillering) since
# the sub-fields already cover them.
cbs_field_map <- tibble(
  studierichting = c(
    "011 Onderwijskunde, lerarenopleiding...",
    "021 Kunst",
    "022 Geesteswetenschappen",
    "023 Talen",
    "031 Gedrag en maatschappij",
    "032 Journalistiek, informatievoorziening",
    "041 Bedrijfskunde en administratie",
    "042 Recht",
    "051 Biologie en biochemie",
    "052 Milieu en natuurbeheer",
    "053 Natuurwetenschappen",
    "054 Wiskunde en statistiek",
    "061 Informatica",
    "071 Techniek en technische dienstverlening",
    "072 Levensmiddelentechnologie",
    "073 Architectuur en bouwkunde",
    "081 Landbouw",
    "082 Bosbouw",
    "083 Visserij",
    "084 Diergeneeskunde en -verzorging",
    "091 Gezondheidszorg",
    "092 Welzijn",
    "101 Persoonlijke dienstverlening",
    "102 Hygiëne en arbeidsomstandigheden",
    "103 Openbare veiligheid",
    "104 Transport en logistiek",
    "Onderwijsrichting onbekend"
  ),
  field = c(
    "education",
    "arts_languages",
    "arts_languages",
    "arts_languages",
    "behavior_society",
    "behavior_society",
    "business_admin",
    "law",
    "natural_sciences",
    "natural_sciences",
    "natural_sciences",
    "natural_sciences",
    "engineering",
    "engineering",
    "engineering",
    "engineering",
    "agriculture",
    "agriculture",
    "agriculture",
    "agriculture",
    "healthcare",
    "healthcare",
    "services",
    "services",
    "services",
    "services",
    NA_character_
  )
)

# Map aggregate CBS fields to CBS field groups (used by the Internationals file)
cbs_field_map_instr <- tibble(
  studierichting = c("01 Onderwijs",
                     "02 Vormgeving, kunst, talen en gesch... ",
                     "03 Journalistiek, gedrag en maatschappij",
                     "041 Bedrijfskunde en administratie",
                     "042 Recht",
                     "05 Wiskunde, natuurwetenschappen",
                     "06 Informatica",
                     "07 Techniek, industrie en bouwkunde",
                     "08 Landbouw, diergeneeskunde en -verz...",
                     "09 Gezondheidszorg en welzijn",
                     "10 Dienstverlening",
                     "Onderwijsrichting onbekend"),
  field = c("education", "arts_languages", "behavior_society",
                "business_admin", "law",
                "natural_sciences", "engineering", "engineering",
                "agriculture", "healthcare",
                "services", NA_character_)
)

# Map CBS shares to CBS field groups (aggregates sub-fields to CBS field level)
field_shares <- cbs_share %>%
  inner_join(cbs_field_map, by = "studierichting") %>%
  group_by(field, onderwijssoort, year) %>%
  summarise(share_int = weighted.mean(share_int, total_seniors, na.rm = TRUE),
            .groups = "drop")

# Aggregate share_int for the joint (pooled) specification
field_shares_agg <- cbs_share %>%
  inner_join(cbs_field_map, by = "studierichting") %>%
  group_by(field, year) %>%
  summarise(share_int = weighted.mean(share_int, total_seniors, na.rm = TRUE),
            .groups = "drop")

# -- Proportional Cariben correction -------------------------------------------
# Caribbean students are Kingdom citizens who don't need Schengen visas.
# The instrument cannot shift their enrollment, so remove their proportional
# contribution from share_int using the Internationals file's breakdown.
# This preserves the subgroups file's foreign counts (which give the correct
# F-stat) while proportionally removing only the Cariben portion.

# Cariben foreign students per aggregate field from the Internationals file
cariben_raw <- read_delim(file.path(base, "Students by homeland, gender, FoS, Senior year, Subgroupped.csv"),
                          delim = ";", show_col_types = FALSE) %>%
  clean_names() %>%
  filter(herkomstland == "Nederlandse Cariben",
         geboorteland_ouders == "Geboren buiten Nederland") %>%
  mutate(
    count = as.numeric(str_trim(ouderejaarsstudenten_aantal)),
    year = as.integer(str_sub(perioden, 1, 4))
  ) %>%
  filter(!is.na(count), count > 0) %>%
  group_by(studierichting, onderwijssoort, year) %>%
  summarise(cariben = sum(count, na.rm = TRUE), .groups = "drop")

# Total foreign students per aggregate field from the Internationals file
intl_total_raw <- read_delim(file.path(base, "Students by homeland, gender, FoS, Senior year, Subgroupped.csv"),
                             delim = ";", show_col_types = FALSE) %>%
  clean_names() %>%
  filter(geboorteland_ouders == "Geboren buiten Nederland") %>%
  mutate(
    count = as.numeric(str_trim(ouderejaarsstudenten_aantal)),
    year = as.integer(str_sub(perioden, 1, 4))
  ) %>%
  filter(!is.na(count), count > 0) %>%
  group_by(studierichting, onderwijssoort, year) %>%
  summarise(intl_total = sum(count, na.rm = TRUE), .groups = "drop")

# Map both to CBS fields
cariben_field <- cariben_raw %>%
  inner_join(cbs_field_map_instr, by = "studierichting") %>%
  filter(!is.na(field)) %>%
  group_by(field, onderwijssoort, year) %>%
  summarise(cariben = sum(cariben, na.rm = TRUE), .groups = "drop")

intl_total_field <- intl_total_raw %>%
  inner_join(cbs_field_map_instr, by = "studierichting") %>%
  filter(!is.na(field)) %>%
  group_by(field, onderwijssoort, year) %>%
  summarise(intl_total = sum(intl_total, na.rm = TRUE), .groups = "drop")

# Cariben share of foreign students per CBS field
cariben_share <- cariben_field %>%
  left_join(intl_total_field, by = c("field", "onderwijssoort", "year")) %>%
  mutate(cariben_share = ifelse(intl_total > 0, cariben / intl_total, 0)) %>%
  select(field, onderwijssoort, year, cariben_share)

cariben_share_agg <- cariben_field %>%
  left_join(intl_total_field, by = c("field", "onderwijssoort", "year")) %>%
  group_by(field, year) %>%
  summarise(
    cariben = sum(cariben, na.rm = TRUE),
    intl_total = sum(intl_total, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(cariben_share = ifelse(intl_total > 0, cariben / intl_total, 0)) %>%
  select(field, year, cariben_share)

# Apply correction: share_int = share_int * (1 - cariben_share)
field_shares <- field_shares %>%
  left_join(cariben_share, by = c("field", "onderwijssoort", "year")) %>%
  mutate(
    cariben_share = coalesce(cariben_share, 0),
    share_int = share_int * (1 - cariben_share)
  ) %>%
  select(field, onderwijssoort, year, share_int)

field_shares_agg <- field_shares_agg %>%
  left_join(cariben_share_agg, by = c("field", "year")) %>%
  mutate(
    cariben_share = coalesce(cariben_share, 0),
    share_int = share_int * (1 - cariben_share)
  ) %>%
  select(field, year, share_int)
```

# 4. Process CBS -> Nationality Shares (2010, per AEA 10-year lag)

```{r process-cbs}
# -- Extract pre-period nationality shares -------------------------------------
# The shift-share instrument uses pre-period (2010) nationality shares as weights.
# We use year 2010 --- a full decade before the outcome period (2020-2024) --- to
# ensure the shares are exogenous to contemporaneous shocks (AEA convention).
# -- Clean and parse directly from EthnicalMakeup -----------------------------
cbs_shares <- EthnicalMakeup %>%
  mutate(
    nationality = str_to_title(str_trim(Nationaliteit)),
    permits = as.numeric(str_trim(Verblijfsvergunning.regulier.naar.motief.Studie..aantal.)),
    year = as.integer(str_extract(as.character(Perioden), "^[0-9]{4}"))
  ) %>%
  filter(!is.na(permits), permits > 0, !is.na(year))

# -- Aggregate to 2010 nationality shares -------------------------------------
# Keep only 2010, sum permits per nationality, compute share of total
nat_shares <- cbs_shares %>%
  filter(year == 2010) %>%                                   # Pre-period: 2010 only
  group_by(nationality) %>%                                  # Group by nationality name
  summarise(permits = sum(permits, na.rm = TRUE), .groups = "drop") %>%  # Total permits per nationality
  mutate(share = permits / sum(permits)) %>%                   # Each nationality's share of ALL permits
  arrange(desc(share))                                         # Sort by share (largest first)

# -- Map Dutch nationality adjectives to English country names -----------------
# CBS uses Dutch nationality adjectives (e.g. "Chinees", "Indiaas"), while IND
# visa data uses English country names. Complete named-vector lookup covering
# all 80 nationalities in the CBS permit data.

# Manual overrides for Dutch adjectives that CLDR regex can't resolve cleanly
nl_adj_to_en <- c(
  "Chinees"        = "CHINA",
  "Indiaas"        = "INDIA",
  "Turks"          = "TURKEY",
  "Turkse"         = "TURKEY",
  "Amerikaans"     = "UNITED STATES",
  "Brits"          = "UNITED KINGDOM",
  "Duits"          = "GERMANY",
  "Frans"          = "FRANCE",
  "Italiaans"      = "ITALY",
  "Spaans"         = "SPAIN",
  "Pools"          = "POLAND",
  "Roemeens"       = "ROMANIA",
  "Bulgaars"       = "BULGARIA",
  "Grieks"         = "GREECE",
  "Portugees"      = "PORTUGAL",
  "Hongaars"       = "HUNGARY",
  "Litouws"        = "LITHUANIA",
  "Russisch"       = "RUSSIA",
  "Braziliaans"    = "BRAZIL",
  "Japans"         = "JAPAN",
  "Mexicaans"      = "MEXICO",
  "Koreaans"       = "SOUTH KOREA",
  "Zuid-Koreaans"  = "SOUTH KOREA",
  "Indonesisch"    = "INDONESIA",
  "Iraans"         = "IRAN",
  "Taiwanees"      = "TAIWAN",
  "Maledivisch"    = "MALDIVES",
  "Surinaams"      = "SURINAME",
  "Oekraïens"      = "UKRAINE",
  "Bengalees"      = "BANGLADESH",
  "Filippijns"     = "PHILIPPINES",
  "Burger Servië"  = "SERBIA",
  "Ugandees"       = "UGANDA",
  "Argentijns"     = "ARGENTINA",
  "Armeens"        = "ARMENIA",
  "Mongools"       = "MONGOLIA",
  "Venezolaans"    = "VENEZUELA",
  "Wit-Russisch"   = "BELARUS",
  "Kosovaars"      = "KOSOVO",
  "Oezbeeks"       = "UZBEKISTAN",
  "Algerijns"      = "ALGERIA",
  "Jemenitisch"    = "YEMEN",
  "Canadees"       = "CANADA",
  "Australisch"    = "AUSTRALIA",
  "Singaporees"    = "SINGAPORE",
  "Nepalees"       = "NEPAL",
  "Marokkaans"     = "MOROCCO",
  "Nieuw-Zeelands" = "NEW ZEALAND",
  "Zuid-Afrikaans" = "SOUTH AFRICA",
  "Sierra Leoons"  = "SIERRA LEONE",
  "Salvadoraans"   = "EL SALVADOR",
  "Omanitisch"     = "OMAN",
  "Myanmarees"     = "MYANMAR",
  "Chileens"       = "CHILE",
  "Bhutaans"       = "BHUTAN",
  "Kazaks"         = "KAZAKHSTAN",
  "Sri Lankaans"   = "SRI LANKA",
  "Ecuadoraans"    = "ECUADOR",
  "Costa Ricaans"  = "COSTA RICA",
  "Macedonisch"    = "NORTH MACEDONIA",
  "Boliviaans"     = "BOLIVIA",
  "Moldavisch"     = "MOLDOVA",
  "Eritrees"       = "ERITREA",
  "Cubaans"        = "CUBA",
  "Jordaans"       = "JORDAN",
  "Liberiaans"     = "LIBERIA",
  "Syrisch"        = "SYRIA",
  "Tunesisch"      = "TUNISIA",
  "Albanees"       = "ALBANIA",
  "Kroatisch"      = "CROATIA",
  "Maleisisch"     = "MALAYSIA",
  "Bosnisch"       = "BOSNIA AND HERZEGOVINA",
  "Libanees"       = "LEBANON",
  "Namibisch"      = "NAMIBIA",
  "Soedanees"      = "SUDAN",
  "Azerbeidzjaans" = "AZERBAIJAN",
  "Georgisch"      = "GEORGIA",
  "Colombiaans"    = "COLOMBIA",
  "Ethiopisch"     = "ETHIOPIA",
  "Vietnamees"     = "VIETNAM",
  "Nigeriaans"     = "NIGERIA",
  "Ghanees"        = "GHANA",
  "Keniaans"       = "KENYA",
  "Peruaans"       = "PERU",
  "Tanzaniaans"    = "TANZANIA",
  "Rwandees"       = "RWANDA",
  "Zimbabwaans"    = "ZIMBABWE",
  "Thais"          = "THAILAND",
  "Egyptisch"      = "EGYPT",
  "Israëlisch"     = "ISRAEL",
  "Kameroens"      = "CAMEROON",
  "Pakistaans"     = "PAKISTAN",
  "Zambiaans"      = "ZAMBIA",
  "Afghaans"       = "AFGHANISTAN"
)

nat_shares <- nat_shares %>%
  mutate(country_std = coalesce(nl_adj_to_en[nationality],
                                str_to_upper(nationality)))
```

# 5. Field × Nationality Matrix (for shift-share instrument)
#
# Use the CBS Internationals file (Herkomstland × Studierichting) to get
# field-specific nationality composition of international students.
# For the "Buiten-Europa" bucket, split into individual nationalities using
# the residence permit shares from section 4.

```{r field-nat-matrix}
base <- "../../../Data/CBS Incoming Student Ethnical Makeup"

# -- Read Internationals file --------------------------------------------------
intl_field <- read_delim(file.path(base, "Students by homeland, gender, FoS, Senior year, Subgroupped.csv"),
                         delim = ";", show_col_types = FALSE) %>%
  clean_names() %>%
  mutate(
    count = as.numeric(str_trim(ouderejaarsstudenten_aantal)),
    year = as.integer(str_sub(perioden, 1, 4))
  ) %>%
  # Keep onderwijssoort (HBO/WO) separate for level-specific instruments
  filter(!is.na(count), count > 0) %>%
  group_by(studierichting, herkomstland, onderwijssoort, year) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop")

# -- Aggregate across onderwijssoort for the joint instrument -----------------
intl_field_agg <- intl_field %>%
  group_by(studierichting, herkomstland, year) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop")

# -- Map CBS Herkomstland to English country_std used in visa data --------------
# Keep the line count from each origin group, then split Buiten-Europa.
# The 5 named groups are already separate from the Buiten-Europa bucket.
herkomst_map <- tibble(
  herkomstland = c("Nederland",
                   "Europa (exclusief Nederland)",
                   "Turkije", "Marokko", "Suriname",
                   "Nederlandse Cariben", "Indonesië",
                   "Buiten-Europa (excl. 5 grote herkomst..."),
  country_std = c(NA_character_,       # exclude Dutch
                  "EUROPE",            # keep as aggregate
                  "TURKEY", "MOROCCO", "SURINAME",
                  "NEDERLANDSE CARIBEN", "INDONESIA",
                  "BUITEN_EUROPA")     # to be split
)

intl_origin <- intl_field %>%
  inner_join(herkomst_map, by = "herkomstland") %>%
  filter(!is.na(country_std))  # drop Netherlands

intl_origin_agg <- intl_field_agg %>%
  inner_join(herkomst_map, by = "herkomstland") %>%
  filter(!is.na(country_std))

# Non-EU/EEA nationalities from the permit data that are NOT among the 5
# named groups (Turkey, Morocco, Suriname, Dutch Caribbean, Indonesia).
# CBS codes "Europa (exclusief Nederland)" as EU/EEA aggregate, so we
# need to exclude EU + EEA member states from the Buiten-Europa split.
# UK included too — was in the EU at baseline (2011) when shift-share
# weights are computed. UK Schengen visa rates also irrelevant (visa-free).
eu_eea_countries <- c("EUROPE",
  "AUSTRIA", "BELGIUM", "BULGARIA", "CROATIA", "CYPRUS",
  "CZECH REPUBLIC", "DENMARK", "ESTONIA", "FINLAND", "FRANCE",
  "GERMANY", "GREECE", "HUNGARY", "ICELAND", "IRELAND", "ITALY",
  "LATVIA", "LIECHTENSTEIN", "LITHUANIA", "LUXEMBOURG", "MALTA",
  "NETHERLANDS", "NORWAY", "POLAND", "PORTUGAL", "ROMANIA",
  "SLOVAKIA", "SLOVENIA", "SPAIN", "SWEDEN", "SWITZERLAND",
  "UNITED KINGDOM")

five_named_groups <- c("TURKEY", "MOROCCO", "SURINAME",
                       "NEDERLANDSE CARIBEN", "INDONESIA")

non_eu_countries <- nat_shares %>%
  filter(!is.na(country_std)) %>%
  filter(!country_std %in% c(eu_eea_countries, five_named_groups, "NETHERLANDS", "NEDERLAND")) %>%
  # Re-normalize shares to sum to 1 within this non-EU group
  mutate(share_non_eu = share / sum(share)) %>%
  select(country_std, share_non_eu)

# For each (studierichting, year), split the Buiten-Europa count
# across non-EU countries proportional to their permit share
bu_count <- intl_origin %>%
  filter(country_std == "BUITEN_EUROPA") %>%
  select(studierichting, onderwijssoort, year, bu_count = count)

bu_count_agg <- intl_origin_agg %>%
  filter(country_std == "BUITEN_EUROPA") %>%
  select(studierichting, year, bu_count = count)

bu_split <- bu_count %>%
  cross_join(non_eu_countries) %>%
  mutate(count = bu_count * share_non_eu) %>%
  select(studierichting, onderwijssoort, year, country_std, count)

bu_split_agg <- bu_count_agg %>%
  cross_join(non_eu_countries) %>%
  mutate(count = bu_count * share_non_eu) %>%
  select(studierichting, year, country_std, count)

# -- Combine all origin groups --------------------------------------------------
intl_detailed <- intl_origin %>%
  filter(country_std != "BUITEN_EUROPA") %>%
  bind_rows(bu_split)

intl_detailed_agg <- intl_origin_agg %>%
  filter(country_std != "BUITEN_EUROPA") %>%
  bind_rows(bu_split_agg)

# -- Map CBS Studierichting to CBS field groups ----------------------------------
cbs_field_map_instr <- tibble(
  studierichting = c("01 Onderwijs",
                     "02 Vormgeving, kunst, talen en gesch... ",
                     "03 Journalistiek, gedrag en maatschappij",
                     "041 Bedrijfskunde en administratie",
                     "042 Recht",
                     "05 Wiskunde, natuurwetenschappen",
                     "06 Informatica",
                     "07 Techniek, industrie en bouwkunde",
                     "08 Landbouw, diergeneeskunde en -verz...",
                     "09 Gezondheidszorg en welzijn",
                     "10 Dienstverlening",
                     "Onderwijsrichting onbekend"),
  field = c("education", "arts_languages", "behavior_society",
                "business_admin", "law",
                "natural_sciences", "engineering", "engineering",
                "agriculture", "healthcare",
                "services", NA_character_)
)

# -- Compute nationality shares per field-year ----------------------------------
# Baseline year = earliest year in data (2011)
baseline_year <- min(intl_detailed$year)

field_nat_shares <- intl_detailed %>%
  inner_join(cbs_field_map_instr, by = "studierichting") %>%
  group_by(field, onderwijssoort, year, country_std) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop") %>%
  group_by(field, onderwijssoort, year) %>%
  mutate(share_in_field = count / sum(count)) %>%
  ungroup()

field_nat_shares_agg <- intl_detailed_agg %>%
  inner_join(cbs_field_map_instr, by = "studierichting") %>%
  group_by(field, year, country_std) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop") %>%
  group_by(field, year) %>%
  mutate(share_in_field = count / sum(count)) %>%
  ungroup()

# Baseline nationality shares per field (time-invariant, from earliest year)
field_nat_bl <- field_nat_shares %>%
  filter(year == baseline_year) %>%
  select(field, onderwijssoort, country_std, share_jf = share_in_field)

field_nat_bl_agg <- field_nat_shares_agg %>%
  filter(year == baseline_year) %>%
  select(field, country_std, share_jf = share_in_field)

# -- Construct field-specific instrument: Z_ft = sum_j(share_jf × accept_rate_jt)
Z_visa_field_year <- field_nat_bl %>%
  inner_join(Consulates, by = c("country_std" = "country")) %>%
  filter(!is.na(acceptance_rate), !is.na(share_jf)) %>%
  group_by(field, onderwijssoort, year) %>%
  summarise(Z_visa_field = sum(share_jf * acceptance_rate, na.rm = TRUE),
            .groups = "drop")

# Joint instrument: aggregate across onderwijssoort for the pooled specification
Z_visa_field_agg <- field_nat_bl_agg %>%
  inner_join(Consulates, by = c("country_std" = "country")) %>%
  filter(!is.na(acceptance_rate), !is.na(share_jf)) %>%
  group_by(field, year) %>%
  summarise(Z_visa_field = sum(share_jf * acceptance_rate, na.rm = TRUE),
            .groups = "drop")
```


# 6. LISS Panel Data
#
# The LISS panel provides individual-level survey data on Dutch households.
# We use two modules: "Work and Schooling" (field of study, education level,
# working hours) and "Economic Situation: Income" (gross monthly wages,
# partner status). The survey covers five years (2020-2024), with a change
# in column naming convention between early (2020-2021) and late (2022-2024)
# waves.

## 6.1 Field-of-Study Code Map

```{r liss-field-map}
# -- LISS field-of-study code -> CBS field group --------------------------------
# LISS codes 011-027 map to 11 CBS field groups. Codes 018, 019, 020 are STEM.
field_map <- tribble(
  ~code, ~field,             ~stem,
  "011", NA,                 FALSE,   # General / no specific field
  "012", "education",        FALSE,   # Teacher training
  "013", "arts_languages",   FALSE,   # Art
  "014", "arts_languages",   FALSE,   # Humanities (languages, history, theology)
  "015", "behavior_society", FALSE,   # Social and behavioral studies
  "016", "business_admin",   FALSE,   # Economics, management, business
  "017", "law",              FALSE,   # Law, public administration
  "018", "natural_sciences", TRUE,    # Mathematics, physics, IT
  "019", "engineering",      TRUE,    # Technology, architecture, crafts
  "020", "agriculture",      TRUE,    # Agriculture, forestry, environment
  "021", "healthcare",       FALSE,   # Medical, health services, nursing
  "022", "services",         FALSE,   # Personal care services
  "023", "services",         FALSE,   # Catering, recreation
  "024", "services",         FALSE,   # Transport, logistics
  "025", "services",         FALSE,   # Telecommunication
  "026", "services",         FALSE,   # Public order and safety (police, army)
  "027", NA,                 FALSE    # Other area
)
```

## 6.2 Work and Schooling Module

```{r liss-parse-work}
# Survey year identifier
survey_years <- c(2020, 2021, 2022, 2023, 2024)

# -- Extract and harmonize fields across all waves -----------------------------
# Each LISS wave records the same information (field of study, age, education,
# hours worked, tenure), but column names change between early (2020-2021) and
# late (2022-2024) waves. We process each year independently, rename to a
# consistent schema, then bind the results into a single panel.

work_list <- list()

for (yr in survey_years) {

  if (yr <= 2021) {
    # -- Early waves (2020-2021): descriptive column names --------------------
    df <- get(paste0("LISS", yr, "WS"))
    names(df)[1] <- "nomem_encr"

    # Field of study: columns 12-28 map to codes 011-027
    # Data stores "Yes"/"No" strings (not numeric codes) — match "Yes" directly
    field_codes <- sprintf("%03d", 11:27)   
    mat <- as.data.frame(lapply(df[, 12:28], function(x) tolower(trimws(as.character(x))) == "yes"))
    liss_code <- apply(mat, 1, function(r) {
      idx <- which(r)
      if (length(idx) == 0) NA_character_ else field_codes[idx[1]]
    })

    # Age: column 6
    age <- as.numeric(df[[6]])

    # Education
    edu_col <- grep("highest level.*education.*completed.*diploma",
                    names(df), value = TRUE, ignore.case = TRUE)[1]
    edu_text <- as.character(df[[edu_col]])

    # Hours worked
    hrs_col <- grep("actually work", names(df), value = TRUE, ignore.case = TRUE)[1]
    hours_worked <- if (!is.na(hrs_col)) as.numeric(df[[hrs_col]]) else NA_real_

    # Tenure
    ten_col <- grep("year.*(current employer|enter into employment|start.*employment)",
                    names(df), value = TRUE, ignore.case = TRUE)[1]
    start_year <- if (!is.na(ten_col)) as.numeric(df[[ten_col]]) else NA_real_
    tenure_years <- ifelse(!is.na(start_year) & start_year > 1900 & start_year < yr,
                           yr - start_year, NA_real_)

  } else {
    # -- Late waves (2022-2024): prefixed column codes -----------------------
    df <- get(paste0("LISS", yr, "WS"))
    nm <- names(df)

    # Detect wave letter from the age column (ends in "003")
    age_idx <- grep("003$", nm)[1]
    wave_letter <- substr(nm[age_idx], 5, 5)
    prefix <- paste0("cw", substr(yr, 3, 4), wave_letter)

    # Field of study: columns cwXX011 through cwXX027
    fcols <- nm[grep(paste0("^", prefix, "0[12][0-9]$"), nm)]
    fcodes <- sub(".*([0-9]{3})$", "\\1", fcols)

    nr <- nrow(df)
    liss_code <- rep(NA_character_, nr)
    for (i in seq_len(nr)) {
      vals <- as.integer(df[i, fcols])
      idx <- which(vals %in% c(1, 2))
      if (length(idx) > 0) liss_code[i] <- fcodes[idx[1]]
    }

    # Age
    age <- as.numeric(df[[paste0(prefix, "003")]])

    # Education: standardized code cwXX525
    edu_col <- paste0(prefix, "525")
    if (edu_col %in% nm) {
      edu_code <- as.numeric(df[[edu_col]])
      # Map code 1-13 to years of schooling (Leuven et al. 2013 convention)
      edu_yrs_vec <- c(0, 6, 10, 11, 12, 13, 16, 15, 17, 21, 21, 21, 21)
      education_years <- edu_yrs_vec[edu_code]
    } else {
      education_years <- rep(NA_real_, nr)
    }

    # Hours worked: cwXX126
    hrs_col <- paste0(prefix, "126")
    hours_worked <- if (hrs_col %in% nm) as.numeric(df[[hrs_col]]) else rep(NA_real_, nr)

    # Tenure: cwXX149
    ten_col <- paste0(prefix, "149")
    start_year <- if (ten_col %in% nm) as.numeric(df[[ten_col]]) else rep(NA_real_, nr)
    tenure_years <- ifelse(!is.na(start_year) & start_year > 1900 & start_year < yr,
                           yr - start_year, NA_real_)

    # Keep education_text as NA for late waves (standardized codes used instead)
    edu_text <- rep(NA_character_, nr)
  }

  # -- Education years lookup (early waves only) --------------------------------
  if (yr <= 2021) {
    edu_lookup <- c(
      "did not complete any education"                                               = "0",
      "did not complete primary school"                                              = "0",
      "primary school"                                                               = "6",
      "vmbo theoretical or combined program (preparatory intermediate vocational school)" = "10",
      "vmbo vocational training program (preparatory intermediate vocational school)"    = "10",
      "mulo, ulo, mavo (lower/intermediate secondary education; US: junior high school)" = "10",
      "lower technical school, household school"                                     = "10",
      "lbo (lower professional education)"                                           = "10",
      "vglo (continued lower education)"                                             = "10",
      "lower and continued special education"                                        = "10",
      "havo (higher general secondary education; US: junior high school)"            = "11",
      "hbs (former pre-university education, US: senior high school)"               = "11",
      "mms (intermediate girls' school)"                                             = "11",
      "vwo (pre-university education, US: senior high school)"                       = "12",
      "gymnasium, atheneum, lyceum (types of pre-university education programs)"     = "12",
      "mbo professional training program (intermediate professional education (bbl))" = "13",
      "mbo professional training program (intermediate professional education (bol))" = "13",
      "kmbo (short intermediate professional education), vhbo (preparatory higher professional education)" = "13",
      "mbo-plus to access hbo, short hbo education (less than two years) (higher professional education)" = "14",
      "hbo (higher professional education), institutes of higher education, new style" = "16",
      "teacher training school"                                                      = "16",
      "academic education, bachelor"                                                 = "15",
      "academic education (including technical and economic colleges, former style) bachelor's degree (kandidaats)" = "15",
      "conservatory and art academy"                                                 = "15",
      "academic education, master"                                                   = "17",
      "academic education (including technical and economic colleges, former style) master's degree (doctoraal)" = "17",
      "doctor's degree (Ph.D, including doctoral research program to obtain Ph.D)"   = "21",
      "other"                                                                        = NA_character_
    )
    education_years <- as.numeric(edu_lookup[edu_text])
  }

  # -- Education level (HBO vs WO) ----------------------------------------------
  onderwijssoort <- rep(NA_character_, length(education_years))
  onderwijssoort[education_years == 16] <- "Hoger beroepsonderwijs"
  onderwijssoort[education_years %in% c(15, 17, 21)] <- "Wetenschappelijk onderwijs"

  work_list[[as.character(yr)]] <- data.frame(
    nomem_encr     = as.character(df[[1]]),
    survey_year    = yr,
    liss_code      = liss_code,
    age            = age,
    education_years = education_years,
    onderwijssoort = onderwijssoort,
    hours_worked   = hours_worked,
    tenure_years   = tenure_years,
    stringsAsFactors = FALSE
  )
}

work_all <- bind_rows(work_list)
```

## 6.3 Income Module

```{r liss-parse-income}
# -- Parse gross monthly income and partner status -----------------------------
# Early waves use descriptive column names; late waves use prefixed codes.
parse_income <- function(df, yr) {

  if (yr <= 2021) {
    # -- Early wave ------------------------------------------------------------
    names(df)[1] <- "nomem_encr"
    gc <- grep("gross wages in total", names(df), value = TRUE, ignore.case = TRUE)[1]
    partner_col <- grep("household head lives together with a partner",
                        names(df), value = TRUE, ignore.case = TRUE)[1]

    data.frame(nomem_encr = as.character(df$nomem_encr),
               survey_year = yr,
               gross_income = if (!is.na(gc)) as.numeric(df[[gc]]) else NA_real_,
               partner = if (!is.na(partner_col)) tolower(trimws(as.character(df[[partner_col]]))) == "yes" else NA,
               stringsAsFactors = FALSE)

  } else {
    # -- Late wave (prefixed columns) ------------------------------------------
    nm <- names(df)
    wave_letter <- substr(nm[grep("372$", nm)[1]], 5, 5)
    prefix <- paste0("ci", substr(as.character(yr), 3, 4), wave_letter)
    inc_col <- paste0(prefix, "372")
    partner_col <- paste0(prefix, "003")

    data.frame(nomem_encr = as.character(df[[1]]),
               survey_year = yr,
               gross_income = if (inc_col %in% nm) as.numeric(df[[inc_col]]) else NA_real_,
               partner = if (partner_col %in% nm) as.numeric(df[[partner_col]]) == 1 else NA,
               stringsAsFactors = FALSE)
  }
}

income_all <- bind_rows(
  parse_income(LISS2020ES, 2020),
  parse_income(LISS2021ES, 2021),
  parse_income(LISS2022ES, 2022),
  parse_income(LISS2023ES, 2023),
  parse_income(LISS2024ES, 2024)
)
```

## 6.4 Merge Work and Income

```{r liss-merge}
# -- Combine work + income data into a single panel ----------------------------
LISS_panel <- work_all %>%
  inner_join(income_all, by = c("nomem_encr", "survey_year")) %>%
  left_join(field_map, by = c("liss_code" = "code")) %>%
  filter(!is.na(field),
         !is.na(gross_income),
         gross_income > 0) %>%
  mutate(
    log_income    = log(gross_income),
    age_sq        = age^2,
    hours_sq      = hours_worked^2,
    partner       = as.logical(partner),
    education_years = ifelse(is.na(education_years),
                             median(education_years, na.rm = TRUE),
                             education_years)
  )
```

```{r liss-merge}
# -- Merge work data + income data + field map into a single panel ------------
LISS_panel <- work_all %>%
  # Inner join: keep only respondents present in BOTH work and income modules
  inner_join(income_all, by = c("nomem_encr", "survey_year")) %>%
  # Map LISS numeric field code -> CBS field group + STEM indicator
  left_join(field_map, by = c("liss_code" = "code")) %>%
  # Drop rows missing essential variables (can't estimate without these)
  filter(!is.na(field), !is.na(gross_income), gross_income > 0) %>%
  mutate(
    log_income = log(gross_income),           # Log transform for wage equation (Mincer)
    age_sq = age^2,                                     # Quadratic age term (non-linear experience effects)
    hours_sq = hours_worked^2,                          # Quadratic hours term (non-linear labour supply)
    partner = as.logical(partner),                      # Convert partner indicator to logical TRUE/FALSE
    # Impute missing education years with the sample median (attenuates toward zero -> conservative)
    education_years = ifelse(is.na(education_years), median(education_years, na.rm = TRUE),
                             education_years)
  )
```

## 6.5 Alternative LISS Parsing (Codebook-Driven)
#
# This section reimplements the LISS parsing using exact codebook column names
# instead of grep-based column detection. The approach is:
#   1. Convert late-wave (2022-2024) coded values to match early-wave text
#   2. Rename late-wave coded columns to early-wave descriptive names
#   3. Select the same columns across all years, then bind_rows
#   4. Compare results with the original implementation

### 6.5.1 Codebook Lookup Tables

```{r alt-liss-lookups}
# -- Education code 1-28 -> text (from cw22o codebook p.9-10) -----------------
# The late-wave LISS stores education as codes 1-28, while early waves store
# the equivalent free-text strings. Map codes to text so all years align.
edu_code_to_text <- c(
  "1"  = "did not complete any education",
  "2"  = "did not complete primary school",
  "3"  = "primary school",
  "4"  = "lower and continued special education",
  "5"  = "vglo (continued lower education)",
  "6"  = "lbo (lower professional education)",
  "7"  = "lower technical school, household school",
  "8"  = "mulo, ulo, mavo (lower/intermediate secondary education; US: junior high school)",
  "9"  = "vmbo vocational training program (preparatory intermediate vocational school)",
  "10" = "vmbo theoretical or combined program (preparatory intermediate vocational school)",
  "11" = "mms (intermediate girls' school)",
  "12" = "hbs (former pre-university education, US: senior high school)",
  "13" = "havo (higher general secondary education; US: junior high school)",
  "14" = "vwo (pre-university education, US: senior high school)",
  "15" = "gymnasium, atheneum, lyceum (types of pre-university education programs)",
  "16" = "kmbo (short intermediate professional education), vhbo (preparatory higher professional education)",
  "17" = "mbo professional training program (intermediate professional education) (BOL)",
  "18" = "mbo professional training program (intermediate professional education) (BBL)",
  "19" = "mbo-plus to access hbo, short hbo education (less than two years) (higher professional education)",
  "20" = "hbo (higher professional education), institutes of higher education, new style",
  "21" = "teacher training school",
  "22" = "conservatory and art academy",
  "23" = "academic education (including technical and economic colleges, former style) bachelor's degree (kandidaats)",
  "24" = "academic education (including technical and economic colleges, former style) master's degree (doctoraal)",
  "25" = "academic education, bachelor",
  "26" = "academic education, master",
  "27" = "doctor's degree (Ph.D, including doctoral research program to obtain Ph.D)",
  "28" = "other"
)

# -- Late-wave coded column -> early-wave descriptive column name -------------
# Field-of-study column names (shared across all years, from codebook p.12-13)
field_prefix <- "In what field did you complete your highest level of education? - "

field_code_to_name <- c(
  "011" = paste0(field_prefix, "General or no specific field"),
  "012" = paste0(field_prefix, "Teacher training or education"),
  "013" = paste0(field_prefix, "Art"),
  "014" = paste0(field_prefix, "Humanities (modern or classical languages, history, theology, etc.)"),
  "015" = paste0(field_prefix, "Social and behavioral studies (including organization studies, media, culture, sports and leisure studies, etc.)"),
  "016" = paste0(field_prefix, "Economics, management, business administration, accountancy"),
  "017" = paste0(field_prefix, "Law, public administration"),
  "018" = paste0(field_prefix, "Mathematics, physics, IT"),
  "019" = paste0(field_prefix, "Technology (including architecture, industry, crafts, construction, etc.)"),
  "020" = paste0(field_prefix, "Agriculture, forestry, environment"),
  "021" = paste0(field_prefix, "Medical, health services, nursing, etc."),
  "022" = paste0(field_prefix, "Personal care services (home economics, hair dressing school, etc.)"),
  "023" = paste0(field_prefix, "Catering, recreation"),
  "024" = paste0(field_prefix, "Transport, logistics"),
  "025" = paste0(field_prefix, "Telecommunication"),
  "026" = paste0(field_prefix, "Public order and safety (police, army, fire brigade, etc.)"),
  "027" = paste0(field_prefix, "Other area")
)

# Work module columns to keep (exact names from 2020/2021 data)
work_cols_keep <- c(
  "Number of household member encrypted",
  "Respondent's age",
  "What is the highest level of education that you have completed with diploma or certificate?",
  field_code_to_name,           # all 17 field columns
  "How many hours per week do (did) you actually work on average in your (last) job?",
  "In which year did you enter into employment with your current employer?"
)

# Income module columns (partner name is same for both early waves)
partner_col_name <- "The household head lives together with a partner (wedded or unwedded)."
income_cols_keep <- c("Number of household member encrypted",
                      partner_col_name,
                      "gross_income")
```

### 6.5.2 Standardise Work Module

```{r alt-parse-work}
# -- Convert late-wave coded values to match early-wave text, rename columns --
# Then bind_rows all five years into a single work panel.

alt_work_list <- list()

for (yr in 2020:2024) {

  df <- get(paste0("LISS", yr, "WS"))
  nm <- names(df)

  if (yr <= 2021) {
    # -- Early waves: already have descriptive column names -------------------
    # Just select the columns we need (by name)
    sel <- intersect(work_cols_keep, nm)
    out <- df[, sel, drop = FALSE]
    out$survey_year <- yr
    # Ensure person ID is character (matches income module type)
    out[[1]] <- as.character(out[[1]])

  } else {
    # -- Late waves: rename coded columns -> descriptive, convert values ------
    yr_sfx <- substr(yr, 3, 4)
    wave_letter <- substr(nm[grep("003$", nm)[1]], 5, 5)
    prefix <- paste0("cw", yr_sfx, wave_letter)

    # Build rename mapping: coded name -> descriptive name
    rename_map <- list()

    # Age: cwXX003 -> "Respondent's age"
    rename_map[[paste0(prefix, "003")]] <- "Respondent's age"

    # Education: cwXX005 -> descriptive name
    rename_map[[paste0(prefix, "005")]] <-
      "What is the highest level of education that you have completed with diploma or certificate?"

    # Field columns: cwXX011-027 -> descriptive names
    for (code_num in 11:27) {
      code_str <- sprintf("%03d", code_num)
      old_name <- paste0(prefix, code_str)
      new_name <- field_code_to_name[[code_str]]
      if (old_name %in% nm) {
        rename_map[[old_name]] <- new_name
      }
    }

    # Hours: cwXX127 (actual hours, from codebook p.29) -> descriptive
    rename_map[[paste0(prefix, "127")]] <-
      "How many hours per week do (did) you actually work on average in your (last) job?"

    # Tenure: cwXX134 (start year, from codebook p.32) -> descriptive
    rename_map[[paste0(prefix, "134")]] <-
      "In which year did you enter into employment with your current employer?"

    # Also rename nomem_encr to match early-wave name
    if ("nomem_encr" %in% nm) {
      rename_map[["nomem_encr"]] <- "Number of household member encrypted"
    }

    # Apply rename
    for (old_n in names(rename_map)) {
      if (old_n %in% nm) {
        names(df)[names(df) == old_n] <- rename_map[[old_n]]
      }
    }

    # Convert field column values: 1 -> "Yes", 0/-9 -> "No", NA -> NA
    for (code_num in 11:27) {
      col_name <- field_code_to_name[[sprintf("%03d", code_num)]]
      if (col_name %in% names(df)) {
        vals <- df[[col_name]]
        new_vals <- rep(NA_character_, length(vals))
        new_vals[!is.na(vals) & vals == 1] <- "Yes"
        new_vals[!is.na(vals) & vals == 0] <- "No"
        new_vals[!is.na(vals) & vals == -9] <- "No"
        df[[col_name]] <- new_vals
      }
    }

    # Convert education code 1-28 -> text
    edu_name <- "What is the highest level of education that you have completed with diploma or certificate?"
    if (edu_name %in% names(df)) {
      codes <- as.character(df[[edu_name]])
      df[[edu_name]] <- ifelse(codes %in% names(edu_code_to_text),
                                edu_code_to_text[codes],
                                NA_character_)
    }

    # Select columns matching our keep list
    sel <- intersect(work_cols_keep, names(df))
    out <- df[, sel, drop = FALSE]
    out$survey_year <- yr
    # Ensure person ID is character (matches income module type)
    out[[1]] <- as.character(out[[1]])
  }

  alt_work_list[[as.character(yr)]] <- out
}

alt_work_all <- bind_rows(alt_work_list)
```

### 6.5.3 Standardise Income Module

```{r alt-parse-income}
# -- Same approach for income: rename late-wave coded columns, convert values --

alt_income_list <- list()

for (yr in 2020:2024) {

  df <- get(paste0("LISS", yr, "ES"))
  nm <- names(df)

  if (yr <= 2021) {
    # -- Early waves: select partner column by name, rename income column -------
    partner_col <- partner_col_name
    # Find the income column (name differs by year: "... in 2019 ...", "... in 2020 ...")
    # Hardcoded by year to avoid grep:
    inc_name <- if (yr == 2020) {
      "How much were your gross wages in total in 2019 [at your employer /at all of your employers]?"
    } else {
      "How much were your gross wages in total in 2020 [at your employer / at all of your employers]?"
    }

    out <- data.frame(
      nomem_encr = as.character(df[[1]]),
      partner = tolower(trimws(as.character(df[[partner_col]]))) == "yes",
      gross_income = as.numeric(df[[inc_name]]),
      survey_year = yr,
      stringsAsFactors = FALSE
    )

  } else {
    # -- Late waves: rename coded columns, convert values --------------------
    yr_sfx <- substr(yr, 3, 4)
    wave_letter <- substr(nm[grep("372$", nm)[1]], 5, 5)
    prefix <- paste0("ci", yr_sfx, wave_letter)

    partner_code <- paste0(prefix, "003")
    income_code <- paste0(prefix, "372")

    partner_vals <- if (partner_code %in% nm) {
      as.numeric(df[[partner_code]]) == 1
    } else {
      rep(NA, nrow(df))
    }

    inc_vals <- if (income_code %in% nm) {
      v <- as.numeric(df[[income_code]])
      v[v %in% c(-9, -8)] <- NA
      v
    } else {
      rep(NA_real_, nrow(df))
    }

    out <- data.frame(
      nomem_encr = as.character(df[[1]]),
      partner = partner_vals,
      gross_income = inc_vals,
      survey_year = yr,
      stringsAsFactors = FALSE
    )
  }

  alt_income_list[[as.character(yr)]] <- out
}

alt_income_all <- bind_rows(alt_income_list)
```

### 6.5.4 Merge and Map Fields

```{r alt-merge}
# -- Merge work + income, map LISS codes to field groups, log income ----------
alt_LISS_panel <- alt_work_all %>%
  left_join(alt_income_all, by = c("Number of household member encrypted" = "nomem_encr",
                                    "survey_year" = "survey_year")) %>%
  rename(nomem_encr = "Number of household member encrypted")

# Detect field of study from the 17 field columns
# (same logic as the original, applied to the renamed columns)
field_codes <- sprintf("%03d", 11:27)
field_mat <- as.data.frame(lapply(field_code_to_name, function(nm) {
  if (nm %in% names(alt_LISS_panel)) {
    tolower(trimws(as.character(alt_LISS_panel[[nm]]))) == "yes"
  } else {
    rep(FALSE, nrow(alt_LISS_panel))
  }
}))
names(field_mat) <- field_codes

alt_LISS_panel$liss_code <- apply(field_mat, 1, function(r) {
  idx <- which(r)
  if (length(idx) == 0) NA_character_ else field_codes[idx[1]]
})

# Map LISS code -> field group + STEM
# (reusing the same field_map from the original section 6.1)
alt_LISS_panel <- alt_LISS_panel %>%
  left_join(field_map, by = c("liss_code" = "code")) %>%
  filter(!is.na(field), !is.na(gross_income), gross_income > 0) %>%
  mutate(
    survey_year  = survey_year,
    age          = as.numeric(`Respondent's age`),
    edu_text     = `What is the highest level of education that you have completed with diploma or certificate?`,
    hours_worked = as.numeric(`How many hours per week do (did) you actually work on average in your (last) job?`),
    start_year   = as.numeric(`In which year did you enter into employment with your current employer?`),
    tenure_years = ifelse(!is.na(start_year) & start_year > 1900 & start_year < survey_year,
                           survey_year - start_year, NA_real_),
    log_income   = log(gross_income),
    age_sq       = age^2,
    hours_sq     = hours_worked^2,
    partner      = as.logical(partner),
    education_years = as.numeric(edu_lookup[edu_text]),
    education_years = ifelse(is.na(education_years),
                              median(education_years, na.rm = TRUE),
                              education_years)
  ) %>%
  select(nomem_encr, survey_year, liss_code, field, stem, age, education_years,
         hours_worked, hours_sq, tenure_years, log_income, gross_income,
         age_sq, partner)
```

### 6.5.5 Comparison with Original Implementation

```{r alt-compare}
# -- Compare row counts and key distributions --------------------------------
cat("============ Comparison: Original vs Codebook-Driven ============\
")
cat(sprintf("Original LISS_panel: %d rows\
", nrow(LISS_panel)))
cat(sprintf("Alternative LISS_panel: %d rows\
", nrow(alt_LISS_panel)))

# Compare survey year distributions
cat("\
--- Survey year distribution (original) ---\
")
print(table(LISS_panel$survey_year))
cat("\
--- Survey year distribution (alternative) ---\
")
print(table(alt_LISS_panel$survey_year))

# Compare field distributions
cat("\
--- Field distribution (original) ---\
")
print(table(LISS_panel$field, useNA = "ifany"))
cat("\
--- Field distribution (alternative) ---\
")
print(table(alt_LISS_panel$field, useNA = "ifany"))

# Compare education years
cat(sprintf("\
Education years mean (orig): %.2f\
",
            mean(LISS_panel$education_years, na.rm = TRUE)))
cat(sprintf("Education years mean (alt):  %.2f\
",
            mean(alt_LISS_panel$education_years, na.rm = TRUE)))

# Compare log income
cat(sprintf("\
Log income mean (orig): %.3f\
",
            mean(LISS_panel$log_income, na.rm = TRUE)))
cat(sprintf("Log income mean (alt):  %.3f\
",
            mean(alt_LISS_panel$log_income, na.rm = TRUE)))

# Identify rows unique to each
orig_ids <- paste(LISS_panel$nomem_encr, LISS_panel$survey_year)
alt_ids  <- paste(alt_LISS_panel$nomem_encr, alt_LISS_panel$survey_year)

only_orig <- setdiff(orig_ids, alt_ids)
only_alt  <- setdiff(alt_ids, orig_ids)
common    <- intersect(orig_ids, alt_ids)

cat(sprintf("\
Rows only in original: %d\
", length(only_orig)))
cat(sprintf("Rows only in alternative: %d\
", length(only_alt)))
cat(sprintf("Rows in both: %d\
", length(common)))

if (length(only_orig) > 0) {
  cat("\
Sample of rows only in ORIGINAL:\
")
  print(head(only_orig))
}
if (length(only_alt) > 0) {
  cat("\
Sample of rows only in ALTERNATIVE:\
")
  print(head(only_alt))
}

# Summary verdict
if (identical(orig_ids, alt_ids)) {
  cat("\
*** VERDICT: IDENTICAL (same person-years in both) ***\
")
} else if (length(common) == min(length(orig_ids), length(alt_ids))) {
  cat("\
*** VERDICT: SUBSET (one is subset of the other) ***\
")
} else {
  cat("\
*** VERDICT: DIFFERENT (partial overlap) ***\
")
}
```

# 7. Merge All Data

```{r merge-all}
# -- Build the final estimation dataset ---------------------------------------
est_data <- LISS_panel %>%
  # Merge field-level international share (endogenous variable)
  # Merge level-specific share_int (matches by field x onderwijssoort x year)
  # Merge level-specific share_int (matches by field x onderwijssoort x year)
  left_join(field_shares, by = c("field", "onderwijssoort", "survey_year" = "year")) %>%
  # Merge aggregate share_int for the joint specification
  left_join(field_shares_agg %>% rename(share_int_agg = share_int),
            by = c("field", "survey_year" = "year")) %>%
  # Merge level-specific instrument
  left_join(Z_visa_field_year, by = c("field", "onderwijssoort", "survey_year" = "year")) %>%
  # Merge aggregate instrument for the joint specification
  left_join(Z_visa_field_agg %>% rename(Z_visa_field_agg = Z_visa_field),
            by = c("field", "survey_year" = "year"))

est_data <- est_data %>%
  # Drop any row missing a key variable (complete-case analysis)
  # Filter on aggregate vars for the main joint specification;
  # level-specific (share_int, Z_visa_field) may have NAs for non-matched onderwijssoort
  filter(!is.na(share_int_agg), !is.na(Z_visa_field_agg), !is.na(log_income),
         !is.na(education_years), !is.na(hours_worked), !is.na(partner))
```

```{r merge-all-alt}
# -- Build alternative estimation dataset (from codebook-driven LISS parse) ----
# Same merge logic as merge-all, but sourced from alt_LISS_panel instead.
alt_est_data <- alt_LISS_panel %>%
  left_join(field_shares, by = c("field", "onderwijssoort", "survey_year" = "year")) %>%
  left_join(field_shares_agg %>% rename(share_int_agg = share_int),
            by = c("field", "survey_year" = "year")) %>%
  left_join(Z_visa_field_year, by = c("field", "onderwijssoort", "survey_year" = "year")) %>%
  left_join(Z_visa_field_agg %>% rename(Z_visa_field_agg = Z_visa_field),
            by = c("field", "survey_year" = "year")) %>%
  filter(!is.na(share_int_agg), !is.na(Z_visa_field_agg), !is.na(log_income),
         !is.na(education_years), !is.na(hours_worked), !is.na(partner))
```

# 8. Descriptive Statistics

```{r descriptives}
# -- Summary statistics table --------------------------------------------------
# Mean (SD) shown in each row.
est_data %>%
  summarise(N = sprintf("%d", n()),
            `Log Income` = sprintf("%.3f (%.3f)", mean(log_income), sd(log_income)),
            `Age` = sprintf("%.1f (%.1f)", mean(age), sd(age)),
            `Education` = sprintf("%.1f (%.1f)", mean(education_years), sd(education_years)),
            `Hours` = sprintf("%.1f (%.1f)", mean(hours_worked, na.rm = TRUE), sd(hours_worked, na.rm = TRUE)),
            `Partner` = sprintf("%.2f (%.2f)", mean(partner, na.rm = TRUE), sd(partner, na.rm = TRUE)),
            `Share Intl` = sprintf("%.3f (%.3f)", mean(share_int_agg), sd(share_int_agg)),
            `Z_visa_field` = sprintf("%.4f (%.4f)", mean(Z_visa_field_agg, na.rm = TRUE), sd(Z_visa_field_agg, na.rm = TRUE)),
            `STEM` = sprintf("%.2f (%.2f)", mean(stem), sd(stem))) %>%
  # Reshape to vertical format (one row per statistic)
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>%
  # Render as a publication-quality LaTeX table
  kable(caption = "Descriptive Statistics", booktabs = TRUE) %>%
  kable_styling(latex_options = "hold_position")
```

```{r save-data}
# -- Save estimation dataset for analysis Rmd ----------------------------
# Run analysis.Rmd to load this and run regressions.
write.csv(est_data, './est_data.csv', row.names = FALSE)
cat(sprintf('est_data saved: %d rows
', nrow(est_data)))
```

# ---- Run analysis.Rmd for regressions --------------------------------------
# The analysis file loads est_data.csv and runs all models.
